---
title: "Temporal network analysis (LA Methods, ch. 17) with Dynet"
author: "After Saqr, M. — *Temporal network analysis: Introduction, methods and analysis with R*"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE, fig.align = "center",
                      fig.width = 7, fig.height = 4.5, dpi = 120)
```

## What this is

Chapter 17 of *Learning Analytics Methods and Tutorials* builds a temporal
network from a MOOC discussion forum using `networkDynamic`, `tsna` and `ndtv`.
This is the same analysis in **Dynet**.

It is not a line-by-line transcription. Two things in the chapter's code are
defects, and reproducing them would mean writing worse code to get worse
numbers; both are named and measured in [what differs](#what-differs).

```{r libs}
library(Dynet)
library(dplyr)
library(rio)
```

## 1. Data

```{r import}
net_edges <- import("https://raw.githubusercontent.com/lamethods/data/main/6_snaMOOC/DLT1%20Edgelist.csv") |>
  janitor::clean_names()
net_nodes <- import("https://raw.githubusercontent.com/lamethods/data/main/6_snaMOOC/DLT1%20Nodes.csv")
```

Recode expertise and name the vertex key.

```{r nodes}
net_nodes <- net_nodes |>
  rename(name = UID) |>
  mutate(name = as.character(name),
         expert_level = case_match(experience,
                                   1 ~ "Expert", 2 ~ "Student", 3 ~ "Teacher"))
```

Drop self-loops and keep only discussions that had an exchange. Timestamps stay
as they are: `dynet()` converts them.

```{r prep}
posts <- net_edges |>
  filter(sender != receiver) |>
  mutate(sender = as.character(sender), receiver = as.character(receiver)) |>
  group_by(discussion_title) |>
  filter(n() > 1) |>
  ungroup()

nrow(posts)
n_distinct(posts$discussion_title)
```

## 2. The temporal network

Forum data is threaded: a tie is live from its own post until the thread falls
silent (Saqr & Nouri, 2020). Naming `thread` selects that.

```{r build}
dn_full <- dynet(posts, from = "sender", to = "receiver", time = "timestamp",
                 thread = "discussion_title", nodes = net_nodes,
                 time_unit = "days", directed = TRUE, loops = FALSE)

summary(dn_full)
```

## 3. The active subnetwork

The chapter keeps vertices with more than 20 ties.

```{r active}
dn <- induce_subgraph(dn_full, degree > 20)
dn
```

## 4. Visualisation

```{r net-plot, fig.width = 7, fig.height = 6}
plot(dn, type = "network")
```

Weeks 1 to 4 — the chapter's four `network.extract()` panels.

```{r snapshots, fig.width = 12, fig.height = 3.6}
plot(dn, type = "snapshots", panels = 4)
```

What those panels hold:

```{r snapshot-counts}
summary(snapshots(dn, start = 1, end = 22, step = 7, window = 7))
```

Activity over time.

```{r timeline, fig.height = 5}
plot(dn, type = "timeline", top = 25)
```

Tie formation and dissolution.

```{r turnover}
plot(events(dn, measure = c("formation", "dissolution"),
            start = 0, end = 72, step = 1, window = 1))
```

Proximity timeline.

```{r proximity, fig.width = 9, fig.height = 7}
plot(dn, type = "proximity", slices = 20)
```

## 5. Graph-level measures

A seven-day rolling density.

```{r density}
plot(metrics(dn, measure = "density", start = 14, end = 60, step = 1, window = 7))
```

Density over the whole period, and the time-integrated density.

```{r density-scalars}
metrics(dn, measure = c("density", "temporal_density", "edges"), window = "all")
```

Reciprocity, on the full network as in the chapter.

```{r recip}
plot(metrics(dn_full, measure = "reciprocity",
             start = 1, end = 73, step = 1, window = 1))
```

The dyad census. `window = 0` samples at each instant, which is what
`tSnaStats()` does by default.

```{r dyads}
plot(metrics(dn, measure = c("mutual", "asymmetric", "null"),
             start = 0, end = 72, step = 1, window = 0))
```

Degree centralization.

```{r centralization}
plot(metrics(dn, measure = "centralization_degree",
             start = 1, end = 73, step = 1, window = 1))
```

## 6. Node-level measures

Degree, in-degree and out-degree in one call.

```{r degree}
degree_series <- dyn_centrality(dn, measure = "degree",
                                mode = c("all", "in", "out"),
                                start = 1, end = 73, step = 1, window = 1)
summary(degree_series)
```

```{r degree-plot, fig.height = 4.5}
plot(dyn_centrality(dn, measure = "degree", start = 1, end = 73,
                    step = 1, window = 1))
```

Closeness, betweenness and eigenvector.

```{r other-centrality}
summary(dyn_centrality(dn, measure = c("closeness", "betweenness", "eigenvector"),
                       start = 1, end = 73, step = 1, window = 1))
```

```{r betweenness-plot, fig.height = 4.5}
plot(dyn_centrality(dn, measure = "betweenness",
                    start = 1, end = 73, step = 1, window = 1))
```

Aggregate centrality sits on the vertex table, so it can be read or filtered
without a second call.

```{r node-table}
head(as.data.frame(dn, what = "nodes",
                   measure = c("degree", "betweenness", "eigenvector")))
```

## 7. Reachability

The chapter seeds on the 44th vertex of its induced subgraph, which is
participant 444. Dynet addresses vertices by name.

```{r paths}
fwd <- paths(dn, from = "444", direction = "forward")
fwd
```

```{r path-plot, fig.width = 9, fig.height = 6}
plot(fwd)
```

The chapter's `transmissionTimeline()` becomes a trajectory tree, which carries
route counts and branching probabilities the original does not.

```{r trajectories, fig.width = 10, fig.height = 6}
plot_path_trajectories(path_trajectories(fwd), measure = "time")
```

## 8. Mixing between expertise levels

```{r mixing}
mix <- mixing(dn, attribute = "expert_level",
              start = 1, end = 73, step = 1, window = 1)
summary(mix)
```

```{r mixing-plot, fig.height = 4.5}
plot(mix)
```

`nodemix` drops one cell as the `ergm` base category, so the chapter plots eight
of the nine group pairs. `mixing()` reports all nine.

## What differs {#what-differs}

Two differences from the published figures, both deliberate.

**The spell rule.** The chapter sets every tie in a discussion to start at the
discussion's *first* post. Saqr and Nouri (2020), which `thread =` implements,
starts each tie at *its own* post. Only 302 of the surviving posts are opening
posts, so the chapter's spells are about a quarter longer on average, and its
time-integrated density is correspondingly higher.

**29 ties that never happened.** The chapter builds its static edge list before
dropping single-post discussions, so 29 pairs enter the network with no
activity at all. `networkDynamic()` then gives each of them activity across the
entire observation window, flagged censored at both ends — permanently-on ties
that no message ever created. Five of them fall inside the active subnetwork,
which is why the chapter reports 433 ties where this reports 428, and why its
day-73 statistics are non-empty for a day past the end of observation.

```{r differences}
metrics(dn, measure = c("edges", "temporal_density"), window = "all")
```

## What Dynet does not do

Four chapter outputs have no counterpart, all animation or 3-D rendering:
`compute.animation()`, `timePrism()`, `render.d3movie()`, and the Graphviz
layout used to redraw the path tree. Dynet draws static figures only.

## Do the two implementations agree?

The differences above are definitional, not numerical. Given the *same*
network, Dynet and `tsna`/`sna` return the same values. The check below rebuilds
the chapter's own construction and compares; the bracket indexing is there
because `tsna` returns `mts` matrices.

```{r agree, eval = requireNamespace("tsna", quietly = TRUE)}
library(tsna); library(networkDynamic); library(sna)

chapter_spells <- posts |>
  mutate(day = round(as.numeric(difftime(
    lubridate::parse_date_time(timestamp, "mdy HM"),
    min(lubridate::parse_date_time(timestamp, "mdy HM")), units = "days")), 2)) |>
  group_by(discussion_title) |>
  mutate(onset = min(day), terminus = max(day)) |>
  ungroup()

nd <- networkDynamic(edge.spells = data.frame(
  onset = chapter_spells$onset, terminus = chapter_spells$terminus,
  tail = as.integer(chapter_spells$sender),
  head = as.integer(chapter_spells$receiver)))
same <- dynet(data.frame(from = chapter_spells$sender, to = chapter_spells$receiver,
                         start = chapter_spells$onset, end = chapter_spells$terminus),
              directed = TRUE, loops = FALSE)

active_nd <- get.inducedSubgraph(nd, v = which(sna::degree(nd) > 20))
active_dn <- induce_subgraph(same, degree > 20)

data.frame(
  quantity = c("aggregate density", "mutual dyads per day",
               "degree per node per day", "earliest arrival time"),
  max_abs_diff = c(
    abs(as.data.frame(metrics(active_dn, "density", window = "all"))$value -
        gden(active_nd)),
    max(abs(as.data.frame(metrics(active_dn, "mutual", start = 1, end = 73,
                                  step = 1, window = 1))$value -
            as.numeric(tSnaStats(active_nd, "mutuality", start = 1, end = 73,
                                 time.interval = 1, aggregate.dur = 1)))),
    {
      ref <- tSnaStats(active_nd, "degree", start = 1, end = 73,
                       time.interval = 1, aggregate.dur = 1, cmode = "freeman")
      got <- as.data.frame(dyn_centrality(active_dn, "degree", start = 1,
                                          end = 73, step = 1, window = 1))
      m <- matrix(NA_real_, 73, ncol(ref), dimnames = list(NULL, colnames(ref)))
      m[cbind(match(got$time, sort(unique(got$time))),
              match(got$node, colnames(ref)))] <- got$value
      max(abs(m[seq_len(72), ] - as.matrix(ref)[seq_len(72), ]))
    },
    {
      tp <- tPath(active_nd, v = 44, direction = "fwd")
      dp <- as.data.frame(paths(active_dn, from = "444"))
      a <- setNames(tp$tdist, as.character(network.vertex.names(active_nd)))
      b <- setNames(dp$latency, dp$node)[names(a)]
      ok <- is.finite(a) & is.finite(b)
      max(abs(a[ok] - b[ok]))
    }))
```

```{r session}
sessionInfo()
```
