---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 120,
  out.width = "100%"
)
```

# Dynet

Dynet provides functions for constructing, analysing, and visualising temporal networks in R. The constructor accepts interval, contact, threaded, and co-presence data. Analyses include graph-level structure, vertex-level centrality, relational spell onsets and terminations, duration, burstiness, group mixing, and time-respecting paths. Measurement functions return tidy data frames identified by vertex names, temporal intervals, and measures, as applicable. Core calculations are implemented in base R.

## Temporal networks

Temporal networks map the dynamics of relationships as they occur over time, preserving when interactions occur, their duration, and their temporal order. Two common representations are contact sequences and interval networks (Holme and Saramäki, 2012). In a contact sequence, interactions are recorded as timestamps and treated as instantaneous because their duration is negligible for the analysis. In an interval network, relationships are represented as relational spells with onset and termination times. Both representations accommodate repeated interactions between the same vertices.

In temporal networks, interactions unfold as time progresses, and paths must follow their chronological order. For example, if A shares information with B on Monday and B interacts with C on Tuesday, the information can potentially reach C through B. If B’s interaction with C occurred on Sunday, that interaction could not carry information received on Monday. A sequence that follows this temporal order is called a *time-respecting path*. Unlike static networks, which connect interactions regardless of when they occurred, temporal networks permit only paths that follow the order and availability of interactions. [Saqr and Peeters (2022)](https://doi.org/10.1111/bjet.13187).

Temporal networks also account for when relationships begin and end. A connection may be active during one period and absent during another, changing the network’s structure and the paths available between vertices. Combining all interactions into a static network obscures these changes: the resulting network may appear densely connected even when few connections were active at the same time. [Saqr and Peeters (2022)](https://doi.org/10.1111/bjet.13187).

Temporal networks can be analysed at the graph, vertex, and edge levels. Graph-level measures, such as density and reciprocity, describe the overall network structure. Vertex-level measures, such as degree, closeness, and betweenness, describe the positions of individual vertices. Edge-level measures describe relationships between pairs of vertices, including their onset, termination, duration, and recurrence. Examining these measures over time reveals changes in the network, its vertices, and their relationships. Time-respecting paths provide additional information about reachability, the time required to reach other vertices, and the intermediate vertices involved.

The construction of a temporal network depends on how relationships and their timing are recorded. Dynet supports four data formats. **Interval data** explicitly record when each relationship begins and ends; a duration may be supplied instead of a termination time. **Contact data** record interactions at specific timestamps, with their duration treated as negligible. **Threaded data**, such as discussion replies, include timestamps and discussion identifiers. For these data, the constructor assigns each reply relationship an onset at its posting time and a termination at the last recorded event in its discussion. **Co-presence data** record participation in shared events, such as meetings or seminars; relationships between participants are derived from their attendance at the same event.

## Installation


The development version can be installed from the author’s R-universe
repository, which also provides the `cograph` dependency, or from GitHub.
The installation commands are not evaluated when this page is rendered.

```{r install, eval = FALSE}
install.packages("Dynet",
                 repos = c("https://mohsaqr.r-universe.dev",
                           "https://cloud.r-project.org"))

# or from GitHub
remotes::install_github("mohsaqr/Dynet")
```

## Data


The examples use bundled datasets in tidy format. `school_contacts` is a
simulated interval dataset containing 240 face-to-face contacts among
fourteen students over approximately 21.5 arbitrary time units.
`forum_posts` contains 241 simulated posts by twenty participants over eight
weeks; `forum_people` supplies participant attributes, including role.
`seminar_attendance` records simulated attendance at weekly seminars. The
package also includes `mooc_posts`, containing 2,529 posts from the MOOC
forum analysed by Saqr (2024), and `thought_chains`, a synthetic reply
dataset based on the study by Saqr, López-Pernas, and Törmänen (2026).

```{r data}
library(Dynet)
head(school_contacts)
```

In `school_contacts`, `from` and `to` identify the relational endpoints,
while `start` and `end` specify contact onset and termination.

## Building a network

`dynet()` recognises common column names automatically, using case-insensitive matching. Relational endpoints may be named `from`/`to`, `sender`/`receiver`, or `source`/`target`; interval boundaries may be named `start`/`end` or `onset`/`terminus`. For contact data, recognised timestamp names include `time`, `timestamp`, `date`, and `datetime`. Explicit column specification is needed only when names do not match recognised aliases or when their intended interpretation is ambiguous. The `format` argument can explicitly select the input format; for example, `format = "contact"` specifies instantaneous interactions.

`dynet()` represents relational data in tidy format, retaining the endpoints and recording the onset, termination, duration, and multiplicity of each relational spell. For interval data, termination is supplied through `end` or calculated as `start + duration`. Duration is recorded as `duration = end - start`. Instantaneous contacts have equal onset and termination times and therefore zero duration. Multiplicity is recorded in `weight`, which defaults to 1 when no multiplicity variable is supplied or recognised.

The following example constructs a temporal network from `school_contacts`, a simulated dataset of 240 contacts among fourteen students. The supplied variables `from` and `to` identify the relational endpoints, while `start` and `end` record contact onset and termination in arbitrary time units. These names are recognised automatically, so no column arguments are required.

```{r build}
library(Dynet)
dn <- dynet(school_contacts)
dn
```

`summary()` describes the network over its observation period, including the number of vertices, relational spells, and distinct pairs of connected vertices. It also reports mean snapshot density: the average proportion of possible vertex pairs connected within each measurement interval. This distinguishes the relationships accumulated over the full observation period from the connectivity observed within individual intervals.

```{r summary}
summary(dn)
```

The same constructor can represent timestamped replies as either instantaneous contacts or relationships that remain active within a discussion. In the first call, `time` explicitly identifies the timestamp column, and the constructor infers contact data. In the second, `thread` selects threaded construction: each reply relationship remains active until the last recorded event in its discussion. The `nodes` argument supplies participant attributes from `forum_people`.

```{r shapes}
contacts <- dynet(forum_posts, time = "timestamp")
forum <- dynet(forum_posts, thread = "thread", nodes = forum_people)
```

Co-presence data record participation in shared events rather than direct interactions between participants. The `actor` argument identifies the participant column, and `group` identifies the shared event. The constructor derives an undirected relationship between each pair of participants attending the same event, with onset and termination determined by the event’s recorded temporal extent.

```{r copresence}
seminars <- dynet(seminar_attendance, actor = "student", group = "seminar")
```

These ties represent shared attendance; they do not establish that participants interacted directly.

Dynet accepts numeric times, dates, and date-times. Numeric times retain their supplied scale. Dates and date-times are converted to elapsed time, with the unit selected automatically when `time_unit = "auto"`, the default. The unit can also be specified explicitly through `time_unit`, for example as `"hours"` or `"days"`. In the threaded example, time is expressed in days.

The timeline plot displays when each pair of vertices is connected during the observation period. For interval data, colour indicates the proportion of each time bin during which at least one relational spell is active. Overlapping spells between the same endpoints are counted once, so the plot represents the duration of connectivity rather than the number of interactions.

```{r plot-spells}
plot(dn)
```

## Measuring in windows

Window-based analyses use four arguments to define when measurements are taken and which interactions they include. `start` and `end` delimit the measurement period. `step` specifies the interval between measurements, while `window` specifies the duration covered by each measurement. Setting `window` equal to `step` produces non-overlapping intervals; a larger `window` produces overlapping intervals. For snapshot measures, `window = 0` evaluates the network at individual time points.

`metrics()` computes graph-level measures within the specified temporal windows. These measures describe the structure of the network as a whole. For example, density is the proportion of possible ties present within a window, while reciprocity describes the extent to which directed ties are reciprocated. Other measures characterise connected components, local configurations of ties, and the distribution of centrality across vertices.

The following call computes density at one-unit intervals, using a seven-unit window for each measurement. Because `window` exceeds `step`, successive measurements include overlapping periods.

```{r metrics}
density <- metrics(dn, measure = "density", step = 1, window = 7)
density
```

The result is a tidy data frame identifying the measurement time, measure, and value. Multiple statistics can be requested through `measure`; their results are distinguished by the `measure` column. Available measures include density, reciprocity, transitivity, dyad and triad censuses, component counts, centralisation, and Krackhardt’s indices. Their definitions and calculation conventions are documented in `?metrics`.

`summary()` summarises the resulting time series, including its mean, dispersion, range, and peak time. `plot()` displays how the measure changes over the observation period.

```{r metrics-summary}
summary(density)
plot(density)
```

## Vertex measures

`centrality_series()` computes measures that describe how a vertex is connected to the rest of the network. Each centrality captures a different aspect of structural position. **Degree** counts direct connections, indicating the extent of a vertex’s immediate neighbourhood. In directed networks, indegree counts incoming connections and outdegree counts outgoing connections. **Strength** accounts for the weights of those connections, distinguishing the number of connected neighbours from the multiplicity of their relationships.

**Closeness centrality** is based on the shortest-path distances from a vertex to other vertices. Higher closeness indicates shorter average distances, meaning that fewer steps are required to reach other vertices. In a directed network, these paths follow tie direction.

**Betweenness** measures the extent to which a vertex lies on shortest paths between other vertices. It therefore identifies potential intermediary positions, although occupying such a position does not establish that information actually passed through that vertex.

Other measures incorporate the positions of a vertex’s neighbours. **Eigenvector centrality** assigns higher scores to vertices connected to other highly central vertices. **PageRank** and **hub and authority scores** also use neighbouring vertices’ scores but apply different rules for directed relationships. **Prestige** measures characterise a vertex through its incoming connections, including, for some variants, the vertices that can reach it indirectly. The choice of measure should therefore follow the substantive question and the meaning assigned to tie direction.

By default, these measures are calculated within successive temporal windows. This produces a centrality trajectory for each vertex, showing when it becomes more or less connected, accessible, or prominent as an intermediary. The following call computes degree using the default measurement intervals:

```{r centrality}
degree <- centrality_series(dn, measure = "degree")
degree
```

The result is a tidy data frame containing vertex names, measurement times, measures, and values. `summary()` summarises each trajectory; `peak_time` identifies when the corresponding centrality attained its maximum.

```{r centrality-summary}
summary(degree)
```

`path_centrality()` instead computes centrality from time-respecting paths across the specified observation period. **Temporal closeness** measures how quickly a vertex can reach other reachable vertices, using the inverse of their mean arrival latency. **Temporal betweenness** measures a vertex’s contribution as an intermediary on optimal temporal paths between other vertices. These measures account for the order and availability of interactions, whereas window-based centralities describe the connections present within each window.

```{r temporal-closeness}
closeness <- path_centrality(dn, measure = "closeness")
closeness
```

## Time-respecting paths

`paths()` examines how a source vertex can reach other vertices through a sequence of interactions over time. A path may involve a direct interaction or several intermediate vertices. Each interaction must occur when its tie is available, and a vertex cannot pass something onward before receiving it. Waiting between interactions is permitted: if A contacts B on Monday and B next contacts C on Wednesday, a path from A to C can include the intervening wait.

The timing of these interactions determines which path arrives earliest. Suppose A can contact C directly on Friday, but can reach C through B on Wednesday. The two-hop path through B arrives earlier than the direct path. `paths()` therefore minimises arrival time first, rather than the number of hops. If several paths arrive equally early, it retains those with the fewest hops.

The following call searches from `Ana`, beginning at the start of the observation period in this example:

```{r paths}
from_ana <- paths(dn, from = "Ana")
from_ana
```

The result describes the optimal paths to each destination. `arrival_time` records the earliest time at which the destination can be reached. `latency` is the elapsed time between the search origin and arrival, including any waiting for subsequent interactions. `n_hops` records how many ties the path traverses, while `n_paths` records how many distinct paths achieve both the earliest arrival and the minimum hop count. A destination with `n_paths = 0` is unreachable under the specified temporal constraints.

By default, traversing a tie takes zero time, although a path may still require waiting for a later interaction. A positive `traversal_time` assigns a duration to each traversal, expressed in the network’s time unit. This is an analytical assumption that affects arrival times and potentially reachability; it should be selected according to the process being represented.

```{r paths-summary}
summary(from_ana)
```

The summary describes the reachable set and its path characteristics. These results concern possible routes through the recorded relationships. They do not establish that information, resources, or other content actually travelled along those routes.

Reachability also depends on when the search begins. Starting later excludes earlier contacts that may have connected the source to other vertices. The following call starts at time 18:

```{r paths-late}
from_ana_late <- paths(dn, from = "Ana", start = 18)
summary(from_ana_late)
```

Comparing the two results shows how the timing of participation changes the destinations Ana can reach and the routes available to her. This distinction is lost in a static network, where all recorded connections are available without regard to their timing.

`pathways()` provides the corresponding vertex sequences, such as A → B → C, and counts the optimal paths following each sequence. The path plot displays the routes and their temporal progression.

```{r pathways, fig.height = 5}
pathways(dn, from = "Ana")
plot(from_ana)
```

## Temporal reachability

`reachability()` summarises which vertices can be connected through time-respecting paths. **Forward reachability** identifies the vertices that a source can reach during the specified observation period. **Backward reachability** identifies the vertices from which a focal vertex can be reached by the end of that period. These sets need not be identical: the order of interactions may permit a path from A to B without permitting a return path from B to A.

The `direction` argument selects `"forward"`, `"backward"`, or `"both"`. Setting `measure = "reach_count"` returns the number of reachable vertices other than the focal vertex. Setting `measure = "reach"` expresses this count as a proportion of all other vertices in the network.

```{r reachability}
reach <- reachability(
  dn,
  direction = "both",
  measure = c("reach", "reach_count")
)
reach
```

A forward reach of 1 indicates that the source can reach every other vertex within the specified period. A value of 0 indicates that it cannot reach any other vertex. Intermediate values describe partial reachability. These measures include indirect connections through intermediate vertices, so they capture more than the immediate neighbourhood measured by degree.

Reachability describes whether a temporal connection is possible; it does not describe how quickly the destination can be reached. Two vertices can therefore have equal reachability but different temporal closeness because their arrival latencies differ. Examining reachability alongside path timing distinguishes the extent of possible contact from the time required to achieve it.

## Tie dynamics and duration

Relational spells have an onset, a termination, and a duration. Examining these properties describes when relationships become active, when they cease, and how long they persist. In datasets with repeated interactions, several spells may connect the same pair of vertices. Counts of spell onsets and terminations therefore differ from counts of distinct connected pairs.

`events()` summarises spell onsets and terminations within temporal windows through the measures `"formation"` and `"dissolution"`, which are requested by default. The following call uses the network’s default measurement intervals:

```{r events}
turnover <- events(dn)
summary(turnover)
```

These counts describe changes in relational activity. A spell onset does not necessarily introduce a previously unconnected pair: it may represent another interaction between endpoints that already have an active spell. Similarly, the termination of one spell does not necessarily disconnect its endpoints if another spell remains active.

`durations()` summarises the duration of relationships. By default, results are grouped by endpoint pair. Setting `measure = "mean"` returns the mean spell duration for each pair:

```{r durations}
lengths <- durations(dn, measure = "mean")
lengths
```

Different duration summaries answer different questions. **Summed duration** adds the durations of individual spells, including overlapping spells separately. **Union duration** measures the total time during which at least one spell connects the endpoints, counting overlap once. For example, if two spells connect the same pair during hours 1–3 and 2–4, their summed duration is four hours, while their union duration is three hours. Mean duration describes the average length of an individual spell.

The `unit` argument determines whether durations are summarised for endpoint pairs, individual spells, vertex activity, or ties incident to a vertex. Vertex activity describes when a vertex is present or eligible to participate; incident-tie duration describes its recorded relationships. These quantities should be interpreted separately.

## Burstiness and memory

`burstiness()` describes how interactions are distributed over time using the intervals between successive events. Regular activity produces intervals of similar length, whereas clustered activity produces short intervals within bursts and longer intervals between them. The burstiness coefficient summarises this variation: −1 corresponds to equal intervals, 0 to the exponential inter-event distribution associated with a Poisson process, and values approaching 1 indicate increasingly heterogeneous intervals (Goh and Barabási, 2008).

```{r burstiness}
rhythm <- burstiness(dn)
summary(rhythm)
```

Burstiness and memory describe different properties of event timing. **Burstiness** measures variation in interval lengths, without considering their order. **Memory** measures the correlation between consecutive intervals. Positive memory indicates that short intervals tend to follow short intervals and long intervals tend to follow long intervals. Negative memory indicates a tendency for short and long intervals to alternate. A value near zero indicates little linear association between successive intervals.

A vertex can therefore exhibit bursty activity without substantial memory: its inter-event intervals may vary considerably while consecutive interval lengths remain weakly correlated. Interpreting both measures helps distinguish variability in activity timing from dependence between successive intervals.

## Mixing between groups


`mixing()` counts distinct active vertex pairs within and between groups
defined by a vertex attribute. Repeated spells and their weights do not
multiply these binary-dyad counts. The forum network inherits `role` from
`forum_people`. Setting `step = 60`, which exceeds this network’s observation
period, produces a single interval for the comparison.

```{r mixing}
mixing(forum, attribute = "role", step = 60)
```

## Network aggregation, subsetting, and snapshots

`collapse_network()` summarises a temporal network as a weighted static network. Each edge represents a pair of vertices connected during the observation period, and its weight summarises their relational spells according to the selected rule. This provides an overall description of relationships while removing their temporal order.

Setting `weight = "union_duration"` assigns each edge the total time during which at least one spell connects its endpoints. Overlapping spells are counted once. For example, spells spanning hours 1–3 and 2–4 produce an edge weight of three hours.

```{r collapse}
flat <- collapse_network(dn, weight = "union_duration")
flat
```

`induce_subgraph()` selects part of the network for further analysis. Selection may be based on vertex attributes or structural measures. In the following example, degree is calculated over the full observation period, and vertices with degree greater than 16 are retained together with the relational spells connecting them.

```{r induce}
core <- induce_subgraph(dn, degree > 16)
core
```

`snapshots()` extracts connections within specified temporal windows. The following call returns the connections in the default measurement window containing time 3:

```{r snapshots}
snapshots(dn, at = 3)
```

A snapshot summarises connectivity within its window; it does not retain the ordering of interactions within that interval.

## Editing a temporal network

Dynet provides functions for modifying vertices, relational spells, and observation settings. Each editing function returns a modified network without changing the input object. This allows alternative network specifications to be constructed while retaining the original for comparison.

`add_nodes()` adds vertices and their attributes. `add_ties()` adds relational spells using their endpoints, onset, and termination. The following example adds a vertex named `Omar`, then adds a directed relationship from `Ana` to `Omar`, active from time 4 to time 6:

```{r edit}
dn2 <- add_nodes(dn, data.frame(name = "Omar"))
dn2 <- add_ties(dn2, data.frame(from = "Ana", to = "Omar", start = 4, end = 6))
dn2
```

Related functions support removal, revision, and relabelling. `remove_nodes()` and `remove_ties()` remove selected vertices or spells; `update_nodes()` and `update_ties()` revise their attributes or values; and `rename_nodes()` changes vertex names while maintaining their references in the network.

`set_vertex_spells()` specifies periods during which vertices are present or eligible to participate. `set_observations()` specifies the periods over which the network is observed. These declarations serve different purposes: vertex activity constrains participation, whereas observation periods constrain measurement. Changing observation settings does not alter the original relational spell endpoints.

## Visualising temporal networks

Temporal network visualisations describe both the structure of relationships and their development over time. Different views address different questions: which vertices are connected during a particular period, how connectivity changes between periods, and when individual relationships are active.

Setting `type = "network"` displays connections within a selected measurement window. The `at` argument identifies the time of interest. Setting `type = "snapshots"` displays several temporal slices using a shared layout, making changes in connections easier to compare.

```{r network-views}
plot(dn, type = "network", at = 3)
plot(dn, type = "snapshots", panels = 6)
```

A timeline places relational activity along a time axis. For interval data, colour represents the proportion of each time bin during which an endpoint pair is connected. This view shows periods of activity and inactivity that may be difficult to distinguish in successive network diagrams.

```{r timeline}
plot(dn, type = "timeline")
```

A proximity timeline follows the changing relationships among vertices. Network distances are calculated within temporal slices and represented as positions along a single axis. Lines connect each vertex’s positions across slices, showing how its proximity to other vertices changes. Nearby lines indicate shorter network distances in the corresponding slice; they do not necessarily indicate a direct tie.

```{r proximity, fig.height = 7}
plot(dn, type = "proximity")
```

These views complement one another: network diagrams show connections, activity timelines show their timing, and proximity timelines summarise changes in network distances.

## Animating temporal networks

`animate()` displays a sequence of network snapshots over the observation period. Vertices retain a shared layout across frames, allowing changes in connections to be followed without changes in position obscuring the comparison.

The temporal settings follow the same conventions as window-based measurement. `start` and `end` delimit the period, `step` determines the time between successive frames, and `window` determines the period represented in each frame. A positive window displays connections active within that interval; it does not imply that all displayed connections were active simultaneously.

The output filename determines the format. GIF output requires the `gifski` package, while MP4 and WebM output require `av`. The animation tutorial provides examples of temporal settings, visual options, and export procedures.

## Further documentation

The package documentation provides additional examples and methodological details:

- `vignette("dynet")` presents an analysis from network construction through measurement and interpretation.
- `vignette("building-networks")` explains input formats, column recognition, vertex attributes, observation periods, sessions, and censoring.
- `vignette("ch17-temporal-networks")` reproduces the temporal network analysis chapter of *Learning Analytics Methods and Tutorials* using the bundled MOOC data.
- The [package website](https://pak.dynasite.org/Dynet/) provides a function catalogue, animation examples, dataset tutorials, and a case study using `thought_chains`.

Function help pages document arguments, defaults, return values, and calculation conventions. For example, `?dynet`, `?metrics`, and `?paths` describe network construction, graph-level measurement, and temporal path analysis, respectively.

## References

Butts, C. T. (2008). A relational event framework for social action.
*Sociological Methodology*, 38(1), 155–200.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Holme, P., & Saramäki, J. (2012). Temporal networks. *Physics Reports*, 519(3),
97–125.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.

Masuda, N., & Lambiotte, R. (2016). *A guide to temporal networks*. World
Scientific.

Moody, J. (2002). The importance of relationship timing for diffusion. *Social
Forces*, 81(1), 25–56.

Nicosia, V., Tang, J., Mascolo, C., Musolesi, M., Russo, G., & Latora, V.
(2013). Graph metrics for temporal networks. In P. Holme & J. Saramäki (Eds.),
*Temporal networks* (pp. 15–40). Springer.

Pan, R. K., & Saramäki, J. (2011). Path lengths, correlations, and centrality
in temporal networks. *Physical Review E*, 84(1), 016105.

Saqr, M. (2024). Temporal network analysis: Introduction, methods and analysis
with R. In M. Saqr & S. López-Pernas (Eds.), *Learning analytics methods and
tutorials: A practical guide using R*. Springer.

Saqr, M., López-Pernas, S., & Törmänen, T. (2026). A temporal network approach to reveal the longitudinal dynamics of CSCL group regulation and productive collaboration. *International Journal of Computer-Supported Collaborative Learning*, 21, 237–270. https://doi.org/10.1007/s11412-025-09464-5

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319). ACM.

Saqr, M., & Peeters, W. (2022). Temporal networks in collaborative learning: A case study. *British Journal of Educational Technology*, 53(5), 1283–1303. https://doi.org/10.1111/bjet.13187
