---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 120,
  out.width = "100%"
)
```

# Dynet

The Dynet R package provides functions for constructing, measuring and
visualising temporal networks, that is, networks whose ties are located in
time. A network is built with one constructor from any of four kinds of
relational log: interval logs with a start and an end per tie, contact logs of
instantaneous events, threaded logs such as forum or chat data, and
co-presence logs where actors sharing a group become connected. Measurement
covers time-varying centrality, graph-level structure, tie formation and
dissolution, tie duration, burstiness, group mixing, time-respecting paths,
reachability and temporal centrality. Every function returns a data frame with
one row per observation, addressed by vertex name, so results are read, plotted
and modelled without further manipulation. The implementation is base R; the
values are checked against the `igraph`, `sna` and `tsna` packages in the test
suite. Drawing is delegated to `cograph`, of which every Dynet network is an
instance.

## Temporal networks

A social network is usually recorded as a set of ties between actors, and the
ties are then analysed as if they had all existed at once. Most relational data
are not like that. Conversations, messages, meetings and collaborations happen
at particular moments and last for particular lengths of time. A temporal
network keeps this information: each tie is a *spell*, a pair of actors together
with the time at which the tie began and the time at which it ended, or a single
instant when the tie has no duration (Holme and Saramäki, 2012). The same pair of
actors may have many spells.

Aggregating the spells into one static network loses two things. The first is
duration and intensity: a pair that spoke once for a minute and a pair that
worked together every day become the same edge. The second, and more
consequential, is order. In a static network, a tie from A to B and a tie from B
to C imply a path from A to C. In a temporal network they imply a path only if
the tie from B to C is still active at, or begins after, the moment the tie from
A to B delivered something to B. A path that respects this constraint is called
a *time-respecting path* (Kempe, Kleinberg and Kumar, 2002). Because the
constraint is one-directional, temporal reachability is not symmetric even in an
undirected network, and a static network systematically overstates how many
actors can reach one another (Moody, 2002). Measures built on time-respecting
paths, such as temporal closeness and temporal betweenness, describe the network
as a medium through which something can travel (Pan and Saramäki, 2011;
Nicosia et al., 2013). Measures built on the timing of a single actor's events,
such as burstiness and memory, describe the rhythm of activity itself (Goh and
Barabási, 2008).

Measures of a temporal network are computed in one of two ways. Density,
reciprocity, transitivity, centrality and the counts of ties formed and
dissolved are computed within windows of time: the observation period is
divided into windows, which may tile it or overlap, each window is measured as
a static network, and the result is a time series whose resolution is set by
the width of the window. Reachability, latency, temporal closeness and temporal
betweenness are computed on the spells directly, along time-respecting paths,
and have no static counterpart: they report the earliest time at which one
actor can reach another, how many distinct routes achieve it, and how much of
the network is reachable from a given moment. In educational research these
methods have been used to study how collaboration unfolds within a course and
which students act as bridges over time (Saqr and Nouri, 2020; Saqr, 2024).

Relational logs arrive in different shapes, and the shape determines how spells
are defined. An *interval* log records a start and an end for every row. A
*contact* log records only a timestamp, so every spell is a point. A *threaded*
log, typical of discussion forums, records a timestamp and the thread a message
belongs to; following Saqr and Nouri (2020), a message is taken to be active
from the moment it is posted until the thread falls silent, so a message that
provoked a long discussion stays live longer than one that did not. A
*co-presence* log records which actors attended which occasion; every pair of
attendees is connected for the span of the occasion. Dynet reads all four with
one function and infers the shape from the arguments given to it.

## Installation

Dynet is not on CRAN. The development version is installed from the author's
R-universe repository, which also resolves its `cograph` dependency, or from
GitHub. These commands are not evaluated when this page is rendered.

```{r install, eval = FALSE}
install.packages("Dynet",
                 repos = c("https://mohsaqr.r-universe.dev",
                           "https://cloud.r-project.org"))

# or from GitHub
remotes::install_github("mohsaqr/Dynet")
```

## Data

Four bundled logs are used below. `school_contacts` is a simulated interval log
of 240 face-to-face contacts among fourteen named students, with time in
arbitrary units over about 21.5 units. `forum_posts` is a simulated threaded log
of 241 posts by twenty participants over eight weeks, with a companion table
`forum_people` that gives each participant a role. `seminar_attendance` is a
simulated co-presence log of which students attended which weekly seminar. The
package also ships `mooc_posts`, the 2,529 posts of a real MOOC discussion forum
analysed in Saqr (2024), and `thought_chains`, the reply table of the *Trees of
Thought* study.

```{r data}
library(Dynet)
head(school_contacts)
```

Each row is one contact from one student to another, with the time it began and
the time it ended.

## Building a network

To build a temporal network from an interval log, we call `dynet()` with the
log. The columns `from`, `to`, `start` and `end` are recognised by name;
`sender` and `receiver`, `source` and `target`, `onset` and `terminus`, and
several other spellings are accepted as well, and times may be numeric, dates or
date-times. Printing the network describes its format and shows the first
spells.

```{r build}
dn <- dynet(school_contacts)
dn
```

To see the network's dimensions in one table, we call `summary()`. The mean
snapshot density is the average over one-unit bins of the share of pairs that
are connected in the bin.

```{r summary}
summary(dn)
```

The other three log shapes are built with the same function. For a contact log,
we name the timestamp column in `time`. For a threaded log, we name the thread
column in `thread`, and pass the participant table to `nodes` so that its
attributes travel with the network. For a co-presence log, we name the actor
column in `actor` and the occasion column in `group`.

```{r shapes}
contacts <- dynet(forum_posts, time = "timestamp")
forum <- dynet(forum_posts, thread = "thread", nodes = forum_people)
seminars <- dynet(seminar_attendance, actor = "student", group = "seminar")
forum
```

The threaded network reports its time in days, because the log carried
date-times; the unit is chosen to suit the span and is stated in every result.

To draw the spells over time, we call `plot()` on the network. Each horizontal
segment is one spell.

```{r plot-spells}
plot(dn)
```

## Measuring in windows

Every measuring function that returns a time series takes the same four
arguments. `start` and `end` bound the period, `step` is the interval between
measurements, and `window` is the length of time each measurement covers. When
`window` equals `step`, the windows tile the period; when `window` is larger,
they overlap and the series is a rolling one; when `window` is zero, the network
is sampled at each instant. The defaults tile the observed period into unit
bins.

To obtain a graph-level measure in each window, we call `metrics()` with the
network and `measure` for the statistic. Here density is measured every unit
over a rolling window of seven units.

```{r metrics}
density <- metrics(dn, measure = "density", step = 1, window = 7)
density
```

To summarise a series, we call `summary()` on it. The table reports one row per
measure with its mean, spread, range and the time at which it peaked.

```{r metrics-summary}
summary(density)
```

To draw the series, we call `plot()` on it.

```{r metrics-plot}
plot(density)
```

Several measures may be requested in one call; they arrive stacked in one data
frame with a `measure` column. The graph-level catalogue includes density,
reciprocity, transitivity, the dyad and triad censuses, component counts,
centralisation and the Krackhardt indices, and is listed in `?metrics`.

## Vertex measures

To obtain a centrality for every vertex in every window, we call
`dyn_centrality()` with the network and `measure` for the index. The result has
one row per vertex per time point.

```{r centrality}
degree <- dyn_centrality(dn, measure = "degree")
degree
```

To reduce the series to one row per vertex, we call `summary()` on it. The
`peak_time` column gives the time at which each vertex was most central.

```{r centrality-summary}
summary(degree)
```

The measures above are computed on the network as it stands in each window.
To compute a centrality on time-respecting paths across the whole period
instead, we set `scope` to `"temporal"`. Temporal closeness is the inverse of
the mean time a vertex needs to reach the others; it has one value per vertex,
not one per window.

```{r temporal-closeness}
closeness <- dyn_centrality(dn, measure = "closeness", scope = "temporal")
closeness
```

On a directed network, `mode` selects whether incoming, outgoing or all ties are
counted. The vertex-level catalogue includes degree, strength, closeness,
betweenness, eigenvector centrality, PageRank, hub and authority scores,
coreness, Burt's constraint and several prestige indices, and is listed in
`?dyn_centrality`.

## Time-respecting paths

To find the earliest time-respecting path from one vertex to every other, we
call `paths()` with the network and `from` for the source. The result has one
row per vertex. `arrival_time` is the earliest moment the vertex can be
reached, `latency` is that moment minus the time the source started, `n_hops`
is the number of ties the earliest route uses, and `n_paths` is the number of
distinct routes that arrive equally early.

```{r paths}
from_ana <- paths(dn, from = "Ana")
from_ana
```

To summarise the reachable set, we call `summary()` on the result.

```{r paths-summary}
summary(from_ana)
```

Ana reaches every other student within four hops. Three of them, Ben, Eve and
Finn, never meet her directly and are joined to her only through intermediaries
whose contacts happened in the right order. The same question
asked from a later starting point has a different answer, because the contacts
that carried the earlier routes have already passed. To start the search later,
we set `start`.

```{r paths-late}
from_ana_late <- paths(dn, from = "Ana", start = 18)
summary(from_ana_late)
```

To list the routes themselves, we call `pathways()` with the network and
`from` for the source. Each row is one distinct sequence of vertices, with the
number of earliest routes that follow it.

```{r pathways}
pathways(dn, from = "Ana")
```

To draw the routes as a tree, we call `plot()` on the paths result. A vertex
appears more than once when it is entered at different times, because each entry
time leaves a different set of onward contacts available.

```{r paths-plot, fig.height = 5}
plot(from_ana)
```

To obtain the share of the network each vertex can reach, or be reached from,
we call `dyn_reachability()` with the network; `direction` selects forward,
backward or both.

## Tie dynamics and timing

To count the ties that form and dissolve in each window, we call `events()`
with the network. The default measures are formation and dissolution.

```{r events}
turnover <- events(dn)
summary(turnover)
```

To measure how long ties last, we call `durations()` with the network and
`measure` for the statistic. With `measure = "mean"` the result is the mean
spell length of each pair; `unit` moves the question from pairs to individual
spells or to vertex activity.

```{r durations}
lengths <- durations(dn, measure = "mean")
lengths
```

To measure whether a vertex's contacts are regular or clustered in time, we call
`burstiness()` with the network. Burstiness compares the dispersion of the gaps
between a vertex's events with the exponential reference: 1 is the bursty limit,
0 the Poisson reference and -1 perfectly regular activity. Memory is the
correlation between consecutive gaps.

```{r burstiness}
rhythm <- burstiness(dn)
summary(rhythm)
```

## Mixing between groups

To count ties within and between groups of vertices, we call `mixing()` with
the network and `attribute` for the vertex attribute that defines the groups.
The forum network carries a `role` attribute from `forum_people`. A `step`
wider than the observation period collapses the counts into one table; the
default reports them per window.

```{r mixing}
mixing(forum, attribute = "role", step = 60)
```

## Static views of a temporal network

To reduce a temporal network to a weighted static network, we call
`collapse_network()` with the network and `weight` for the aggregation rule.
The result is a `cograph` network and can be drawn and measured as one.

```{r collapse}
flat <- collapse_network(dn, weight = "union_duration")
flat
```

To restrict a network to some of its vertices, we call `induce_subgraph()` with
the network and a condition on the vertex table in `nodes`. A centrality named
in the condition is computed over the whole observed period.

```{r induce}
core <- induce_subgraph(dn, degree > 16)
core
```

To list the ties active in one window, we call `snapshots()` with the network
and `at` for the time.

```{r snapshots}
snapshots(dn, at = 3)
```

## Editing a network

Editing functions return a new network and leave the input unchanged. To add a
vertex, we call `add_nodes()` with a table of names and attributes. To add a
tie, we call `add_ties()` with a table of spells. `remove_nodes()`,
`remove_ties()`, `update_nodes()`, `update_ties()`, `rename_nodes()`,
`set_vertex_spells()` and `set_observations()` complete the set.

```{r edit}
dn2 <- add_nodes(dn, data.frame(name = "Omar"))
dn2 <- add_ties(dn2, data.frame(from = "Ana", to = "Omar", start = 4, end = 6))
dn2
```

## Drawing and animating

A Dynet network is a `cograph` network, so every `cograph` drawing function
applies to it directly. `plot()` adds the views that concern time. To draw the
network as it stands in one window, we set `type` to `"network"` and give the
time in `at`. To draw a sequence of windows as small multiples with a shared
layout, we set `type` to `"snapshots"`. To draw the proximity timeline, in which
vertices that interact are placed near each other and their positions are
followed through time, we set `type` to `"proximity"`. The proximity timeline
is described in the article on drawing.

To animate the measurement windows, we call `animate()` with the same four grid
arguments as the measuring functions and a file name whose extension selects the
encoder; `.gif` needs the `gifski` package and `.mp4` or `.webm` the `av`
package. The article on animating shows the result.

## Documentation

- `vignette("dynet")` walks through one analysis from log to result.
- `vignette("building-networks")` covers the four log shapes, vertex
  attributes, observation windows, sessions and censoring.
- `vignette("ch17-temporal-networks")` re-runs the temporal network chapter of
  *Learning Analytics Methods and Tutorials* on the bundled MOOC forum.
- The package website at <https://pak.dynasite.org/Dynet/> adds a function
  catalogue, tutorials on animation and on the MOOC data, and a full
  reproduction of the *Trees of Thought* study.

## References

Butts, C. T. (2008). A relational event framework for social action.
*Sociological Methodology*, 38(1), 155–200.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Holme, P., & Saramäki, J. (2012). Temporal networks. *Physics Reports*, 519(3),
97–125.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.

Masuda, N., & Lambiotte, R. (2016). *A guide to temporal networks*. World
Scientific.

Moody, J. (2002). The importance of relationship timing for diffusion. *Social
Forces*, 81(1), 25–56.

Nicosia, V., Tang, J., Mascolo, C., Musolesi, M., Russo, G., & Latora, V.
(2013). Graph metrics for temporal networks. In P. Holme & J. Saramäki (Eds.),
*Temporal networks* (pp. 15–40). Springer.

Pan, R. K., & Saramäki, J. (2011). Path lengths, correlations, and centrality
in temporal networks. *Physical Review E*, 84(1), 016105.

Saqr, M. (2024). Temporal network analysis: Introduction, methods and analysis
with R. In M. Saqr & S. López-Pernas (Eds.), *Learning analytics methods and
tutorials: A practical guide using R*. Springer.

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319). ACM.
