---
title: "Building and inspecting a temporal network"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building and inspecting a temporal network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(Dynet)
```

A temporal network is a relational log plus a clock. Everything Dynet measures
depends on that clock being right, so this vignette is about the first half of
the work: getting a log into a `dynet` object, declaring what the clock means,
and reading back what you actually built before you measure anything.

There is one constructor, `dynet()`. It covers the four shapes relational logs
arrive in, and the shape is inferred from the arguments you name.

## The four log formats

### Interval log: each row carries its own start and end

This is the richest shape. Duration is recorded, so two students who met once
at length are distinguishable from two who met briefly many times. Face-to-face
sensor data, class attendance with sign-in and sign-out, tutoring sessions and
collaboration episodes all arrive this way.

`school_contacts` is fourteen students over three weeks:

```{r}
head(school_contacts, 4)
```

```{r}
school <- dynet(school_contacts)
school
```

No column names had to be spelled out. `from`, `to`, `start` and `end` were
matched from a table of aliases, case-insensitively and after stripping
punctuation, so `Sender`/`Receiver`, `source`/`target` and `onset`/`terminus`
also work untouched. A `duration` column is accepted in place of `end`.

### Contact log: instantaneous events with no duration

A message, a click, a citation, a card swipe. There is a time and nothing else;
the tie exists at an instant and has zero length. `forum_posts` can be read this
way, because a post has a timestamp and no end:

```{r}
head(forum_posts, 3)
```

```{r}
clicks <- dynet(forum_posts, time = "timestamp")
clicks
```

The timestamps are `POSIXct`, and Dynet converted them to elapsed time in a unit
it chose to suit the span — days, here, reported in the header. Numeric times are
left alone and the unit is reported as `"step"`.

Every spell has `duration` zero. That is the honest reading of a contact log, but
it has a consequence worth knowing before you measure: quantities that integrate
occupied time over the clock are exactly zero on point contacts, because points
occupy no time.

```{r}
summary(clicks)
```

`temporal density` is `0`, while `mean snapshot density` is not. The snapshot
measures ask whether a tie was present anywhere inside a bin; the integrated
measures ask how much of the bin it occupied. On instantaneous data those are
genuinely different questions with genuinely different answers.

### Threaded log: an edge stays active until its thread falls silent

Forum, chat and email data have a timestamp and no end, but treating them as
points throws away the fact that a post which provoked a long argument stayed
relevant longer than one that fell flat. Dynet applies the rule from Saqr and
Nouri (2020): a post is active from the moment it appears until the last post in
the same thread. Name the `thread` argument and you get that derivation.

```{r}
forum <- dynet(forum_posts, thread = "thread", nodes = forum_people)
forum
```

Same 241 rows as the contact reading above, same 20 vertices, but the spells now
have length — and the network has an integrated density that is not zero:

```{r}
summary(forum)
```

### Co-presence log: actors sharing a group become connected

Two-mode attendance data. Students are not linked to each other in the source at
all, only to the seminars they turned up to; the ties have to be projected. Name
both `actor` and `group`:

```{r}
head(seminar_attendance, 3)
```

```{r}
seminars <- dynet(seminar_attendance, actor = "student", group = "seminar")
seminars
```

Every pair of students who shared a room becomes a spell, tagged with the group
that created it. Co-presence networks are always undirected — `directed = TRUE`
is silently overridden, because "A attended with B" has no direction. Note again
that `seminar_attendance` carries a `date` and no end time, so each seminar
contributes point contacts; give the source a genuine start and end column if
you know how long the meetings ran.

### `format = "auto"`, and when to say it out loud

The default `format = "auto"` infers the shape from the arguments you name and
the columns present. Naming `thread` selects threaded; naming `actor` and `group`
selects co-presence; an end or duration column gives interval; a bare time column
gives contact.

The inference is deliberately driven by what you *name*, not by what happens to
be in the data. `forum_posts` has a `thread` column, but:

```{r}
dynet(forum_posts)
```

That is a contact network, not a threaded one. Auto-detection will not silently
apply the thread-persistence rule on your behalf — it is a modelling decision,
so you have to ask for it. State `format` explicitly whenever the same table
could plausibly be read two ways, or when you are building inside a function and
want the shape pinned:

```{r}
dynet(school_contacts, format = "interval")
```

## Inspecting what you built

`summary()` on a `dynet` returns a tidy two-column description, one row per
property. Read it before you measure anything.

```{r}
summary(school)
```

Row by row:

* **format** — which of the four readings was used. Check it matches what you
  intended.
* **directed** — whether `from -> to` is meaningful.
* **vertices** — the fixed vertex universe. Every result keeps these rows, so a
  vertex that never sends a tie still appears with a zero.
* **edge spells** — one per derived spell, which for interval input is one per
  source row.
* **distinct pairs** — how many different dyads those spells fall on. The gap
  between 240 spells and 110 pairs is repetition: these students met each other
  again and again.
* **time unit** — `step` for numeric input; `seconds` to `weeks` for calendar
  input, chosen to suit the span.
* **observed from / observed to / span** — the measurement calendar. Without an
  explicit declaration these are the first and last event times.
* **bin width** — the default measurement grid, set by the `interval` argument.
* **time bins** — how many bins of that width fit the observed period.
* **mean snapshot density** — averaged over the bins, the share of possible
  edges present *anywhere* inside each bin.
* **temporal density** — occupied pair-time divided by eligible pair-time over
  the whole history. This one integrates; it is not an average of the row above.
* **sessions** — how many session labels, if the network has any.
* **vertex attributes** — which node columns came along.

The spell table itself, and the vertex table, come back through
`as.data.frame()`:

```{r}
head(as.data.frame(school), 4)
```

```{r}
head(as.data.frame(forum, what = "nodes"), 4)
```

`what` also takes `"bins"` for the measurement grid, `"network"` for the
aggregate edge list, `"observations"` for the observation calendar,
`"observed_edges"` for spells intersected with it, and `"vertex_spells"` for
declared vertex activity.

```{r}
head(as.data.frame(school, what = "bins"), 4)
```

```{r}
head(as.data.frame(school, what = "network"), 4)
```

## Directedness, loops, weights, node attributes

These are all `dynet()` arguments. A small hand-written log makes each one
visible:

```{r}
tiny <- data.frame(
  from  = c("A", "B", "A", "C", "A"),
  to    = c("B", "A", "C", "A", "A"),
  start = c(0, 1, 2, 3, 4),
  end   = c(2, 3, 5, 4, 6),
  posts = c(3, 1, 2, 5, 1)
)
```

`directed = FALSE` folds `A -> B` and `B -> A` onto one dyad, and `weight`
names a column of event multiplicity:

```{r}
dynet(tiny, directed = FALSE, weight = "posts")
```

Five source rows became four spells and two distinct pairs. The fifth row is a
self-loop, `A -> A`, and self-loops are dropped by default with a message —
almost always what a relational log needs. Keep them if the loop means
something in your design:

```{r}
dynet(tiny, loops = TRUE)
```

`interval` sets the default bin width used by every measurement verb, in the
network's own time unit:

```{r}
summary(dynet(tiny, interval = 2))
```

Vertex attributes come in through `nodes`. The vertex key column is
auto-detected, and `groups` names one attribute to use as the vertex partition —
written into the places cograph looks for it, so `cograph::splot()` colours and
groups by it without further argument.

```{r}
roles <- dynet(forum_posts, thread = "thread",
               nodes = forum_people, groups = "role")
head(as.data.frame(roles, what = "nodes"), 4)
```

Those attributes are what `mixing()` partitions on:

```{r}
head(mixing(roles, attribute = "role"), 4)
```

## Sessions

A session is a wall in time. Terms, courses, class periods, hospital admissions:
periods that a time-respecting path should not be allowed to walk across, even
though the clock runs continuously through them. Name a column with `session`.

`school_contacts` has no session column, so derive one — three weeks of contact
give three weekly sessions:

```{r}
weekly <- transform(
  school_contacts,
  week = paste0("week_", floor(start / 7) + 1)
)
head(weekly, 3)
```

```{r}
sessioned <- dynet(weekly, session = "week")
summary(sessioned)
```

Every verb that takes time then takes a `sessions` argument with three settings,
and the choice changes the answer rather than the formatting. Temporal paths show
it most sharply:

```{r}
summary(paths(sessioned, from = "Ana", sessions = "collapse"))
```

`"collapse"` ignores the labels entirely and treats the history as one
continuous stream. Ana reaches all thirteen other students, at a median latency
of 7.51 time steps.

```{r}
summary(paths(sessioned, from = "Ana", sessions = "bounded"))
```

`"bounded"` keeps every path inside a single session. Ana still reaches
everyone, but she has to do it by routes that never cross a week boundary, so
the journeys get longer — median latency rises to 8.21 and the longest journey
grows from four hops to five.

```{r}
summary(paths(sessioned, from = "Ana", sessions = "separate"))
```

`"separate"` reports each session on its own rows instead of pooling them. Now
the weeks are visibly different: within week 1 Ana reaches 6 of the other 13
students, within week 2 all 13, within week 3 only 5.

`"bounded"` is the default everywhere. On a network with no sessions there is
nothing to bound, so `"bounded"` and `"collapse"` give the same answer and the
default costs nothing when you have not declared any sessions. `"separate"` is
the exception: with no session column there is nothing to report separately,
and it raises an error rather than silently pooling.

## Observation windows

The observed range defaults to the first and last event in the log. That is a
guess, and it is usually wrong: a study that ran for four weeks but saw no
contact in the last three days is not a study that ran for 25 days. Since the
observed range is the denominator of every rate and the horizon of every path,
declare it when you know it.

```{r}
bounded <- dynet(school_contacts, observation_start = 0, observation_end = 14)
summary(bounded)
```

The span is now 14 rather than 21.52, and both densities changed because the
exposure they divide by changed. This is an administrative limit, not a filter:
the raw spell boundaries are untouched, and `as.data.frame()` still returns them
exactly as supplied. Positive spells contribute their half-open intersection with
the window; a genuine point contact sitting on either limit is kept.

For interrupted observation — a holiday, a sensor outage, a gap between waves —
give the observed periods directly. Overlapping and adjacent spells are merged
into canonical components.

```{r}
gapped <- dynet(
  school_contacts,
  observation_spells = data.frame(start = c(0, 12), end = c(8, 21))
)
as.data.frame(gapped, what = "observations")
```

```{r}
summary(gapped)
```

Seventeen bins, not 21: the grid restarts inside each component and never
creates a bin that lies entirely in the gap. Exposure sums the component
durations rather than taking the hull, so events falling in the gap are excluded
from the denominator instead of quietly inflating it.

The same calendar is editable after the fact, and `clear_observations()` puts
back the implicit continuous window:

```{r}
as.data.frame(set_observations(school, start = 2, end = 10),
              what = "observations")
```

```{r}
as.data.frame(clear_observations(gapped), what = "observations")
```

## Vertex activity spells

Observation windows say when *the study* was running. Vertex spells say when a
*vertex* was eligible to have ties at all — a student who enrolled late, a nurse
on a rota, a user who deleted their account. Supply a tidy table with `node`,
`start` and `end`.

A vertex with no row in the table stays eligible at all times, so you only
declare the ones that change.

```{r}
arrivals <- data.frame(
  node  = c("Ana", "Ben"),
  start = c(0, 7),
  end   = c(21.52, 21.52)
)
scheduled <- dynet(school_contacts, vertex_spells = arrivals)
as.data.frame(scheduled, what = "vertex_spells")
```

Ben becomes eligible at time 7. Every measurement now runs on the eligible population
rather than the fixed one. The fixed vertex universe is unchanged and Ben keeps
a row in every bin, but the bins before he arrives report `NA` rather than a
degree of zero — the difference between "had no contacts" and "was not there to
have any":

```{r}
head(as.data.frame(dyn_centrality(school, measure = "degree")), 4)
```

```{r}
head(as.data.frame(dyn_centrality(scheduled, measure = "degree")), 4)
```

That distinction moves the summaries, because the empty bins are no longer
averaged in as zeroes:

```{r}
head(summary(dyn_centrality(school, measure = "degree")), 3)
```

```{r}
head(summary(dyn_centrality(scheduled, measure = "degree")), 3)
```

Ana is untouched, at mean degree 2.182 either way. Ben's mean rises from 2.000
to 2.133: the same contacts, divided by the fifteen bins in which he was
actually eligible rather than all twenty-two.

The same declaration is editable. `set_vertex_spells()` replaces the table,
`add_vertex_spells()` adds to it:

```{r}
activity <- set_vertex_spells(school, arrivals)
as.data.frame(add_vertex_spells(activity,
                                data.frame(node = "Cara",
                                           start = 3, end = 12)),
              what = "vertex_spells")
```

## Editing without breaking time

Every editing verb returns a **new** network. Nothing is modified in place, so a
chain of edits is a chain of objects and the source stays exactly as it was.
Each edit rebuilds the canonical spell identities and the cograph projection
together, which is why the temporal object must be edited through these verbs
rather than through cograph's static setters.

Add an isolate before you refer to it as a tie endpoint:

```{r}
step1 <- add_nodes(school, data.frame(name = "Nova", role = "exchange"))
step2 <- add_ties(step1, data.frame(
  from = "Ana", to = "Nova", start = 4, end = 6
))
summary(step2)
```

Fifteen vertices, 241 spells, 111 distinct pairs, and a `role` attribute that
did not exist a moment ago. The original is untouched:

```{r}
summary(school)
```

Removal addresses a tie by name and time, and renaming takes a named vector of
old-to-new:

```{r}
summary(remove_ties(step2, from = "Ana", to = "Nova", start = 4))
```

```{r}
tail(as.data.frame(rename_nodes(step2, c(Nova = "Nova B.")), what = "nodes"), 3)
```

`induce_subgraph()` cuts a temporal subgraph by vertex or by an edge-level
condition, keeping the time structure of what survives:

```{r}
summary(induce_subgraph(school, nodes = c("Ana", "Ben", "Cara", "Dan", "Eve")))
```

Note that the observed range narrowed too, from `0`–`21.52` to `3.17`–`21.33`:
these five students' first contact with one another was not on day zero.

## Basic summaries

With the network built and its calendar declared, the descriptive verbs all
follow the same shape — one call, named arguments, a tidy frame back.

`metrics()` measures graph-level structure in each bin. Ask for several at once
and they arrive stacked with a `measure` column:

```{r}
head(metrics(school, measure = c("density", "edges", "active_nodes")), 6)
```

`summary()` collapses a series to one row per measure, with the time at which it
peaked:

```{r}
summary(metrics(school, measure = c("density", "edges", "active_nodes",
                                    "components", "transitivity",
                                    "reciprocity")))
```

`step` and `window` are separate arguments and that separation is the point.
`step` is how often you look; `window` is how much of the timeline each look
takes in. Equal values tile the period into disjoint bins, which is the default;
a larger window slides an overlapping one:

```{r}
head(metrics(school, measure = "density", step = 1, window = 7), 4)
```

```{r}
plot(metrics(school, measure = "density"))
```

The `temporal_density` and `onset_intensity` selectors are the integrated
counterparts — occupied pair-time and known spell starts per unit of eligible
pair-time, rather than a presence indicator:

```{r}
metrics(school, measure = c("temporal_density", "onset_intensity"),
        step = 7, window = 7)
```

`snapshots()` returns the edges themselves rather than a statistic — the network
as it stood, one row per edge per time point. Give `at` for a single instant, or
leave it out for the whole grid:

```{r}
head(snapshots(school, at = 5), 5)
```

`events()` counts what changed rather than what was present: how many ties formed
and dissolved in each bin.

```{r}
head(events(school), 6)
```

`durations()` summarises how long relationships lasted. `unit` decides what a
duration belongs to — `"pair"` for a dyad's whole history, `"spell"` for the raw
episodes, `"vertex_activity"` and `"vertex_spell"` for vertex presence,
`"node_ties"` for incident tie time.

```{r}
head(durations(school), 6)
```

```{r}
head(durations(school, unit = "spell"), 4)
```

```{r}
head(durations(school, unit = "node_ties", mode = "all"), 4)
```

## Collapsing to a static network

Sometimes the last step is a static picture — a figure, a handover to a
static-network tool, a sanity check. `collapse_network()` flattens any range and
returns a cograph network carrying every weighting it can compute, so you choose
the one you want by name rather than recomputing it.

```{r}
flat <- collapse_network(school, start = 0, end = 7)
flat
```

```{r}
head(as.data.frame(collapse_network(school, start = 0, end = 7,
                                    weight = "union_duration")), 4)
```

`weight` selects which column drives the static edge weight: `"binary"` for
presence, `"union_duration"` for calendar time with at least one spell,
`"total_duration"` for the sum with repetition, `"spell_count"`,
`"weight_sum"`, `"duration_fraction"`, `"weighted_duration"` or
`"latest_weight"`. The result is a `cograph_network`, so `cograph::splot()`
draws it directly.

Collapsing is the end of the analysis, not the start of it. Everything the
static picture cannot say — the order events happened in, how long each tie
lasted, whether a path could have run at all — is in the object you built at the
top of this vignette.

## References

Holme, P., & Saramaki, J. (2012). Temporal networks. *Physics Reports*, 519(3),
97–125.

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge*, 314–319.
