---
title: "Temporal network analysis: the LA Methods chapter with Dynet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Temporal network analysis: the LA Methods chapter with Dynet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.align = "center",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(Dynet)
has_tsna <- requireNamespace("tsna", quietly = TRUE) &&
  requireNamespace("networkDynamic", quietly = TRUE) &&
  requireNamespace("network", quietly = TRUE) &&
  requireNamespace("sna", quietly = TRUE)
```

Chapter 17 of *Learning Analytics Methods and Tutorials* builds a temporal
network from a MOOC discussion forum and takes it through visualisation,
graph-level measures, node-level measures, reachability and mixing, using
`networkDynamic`, `tsna`, `sna` and `ndtv`. This vignette is the same
analysis, in the same order, with Dynet's verbs and the chapter's own data,
bundled here as `mooc_posts` and `mooc_people`.

It is not a line-by-line transcription. Two things in the chapter's code
change the numbers, and reproducing them would mean writing worse code to get
worse numbers; both are named and measured in
[what differs](#what-differs).

## 1. Data

The log is one row per post: who wrote it, who it answers, when, and the
thread it sits in.

```{r data}
head(mooc_posts)
```

```{r people}
head(mooc_people)
```

The chapter recodes the experience level and uses it as the mixing attribute.

```{r recode}
people <- transform(
  mooc_people,
  expert_level = as.character(factor(experience, levels = c(1L, 2L, 3L),
                                     labels = c("Expert", "Student", "Teacher")))
)
with(people, table(expert_level))
```

Self-replies carry no tie, and a thread with a single post never became an
exchange. The chapter drops both.

```{r prepare}
replies <- subset(mooc_posts, sender != receiver)
busy_threads <- with(replies, names(which(table(discussion) > 1)))
exchanges <- subset(replies, discussion %in% busy_threads)
nrow(exchanges)
with(exchanges, length(unique(discussion)))
```

## 2. The temporal network

Forum data is threaded: a tie is live from its own post until the thread falls
silent (Saqr & Nouri, 2020). Naming `thread` selects that rule, and the
timestamps are converted to elapsed days.

```{r build}
dn_full <- dynet(exchanges, from = "sender", to = "receiver",
                 time = "timestamp", thread = "discussion",
                 nodes = people, time_unit = "days",
                 directed = TRUE, loops = FALSE)
summary(dn_full)
```

## 3. The active subnetwork

The chapter keeps the vertices with more than 20 ties. `induce_subgraph()`
takes the condition directly.

```{r active}
dn <- induce_subgraph(dn_full, degree > 20)
dn
```

## 4. Visualisation

```{r net-plot, fig.width = 7, fig.height = 6}
plot(dn, type = "network")
```

Four windows across the course, the chapter's four `network.extract()`
panels.

```{r snapshots, fig.width = 9, fig.height = 3}
plot(dn, type = "snapshots", panels = 4)
```

What those panels hold, week by week:

```{r snapshot-counts}
weekly <- snapshots(dn, start = 1, end = 22, step = 7, window = 7)
weekly_summary <- summary(weekly)
knitr::kable(weekly_summary, caption = "Ties, active participants and weight per week.")
```

Who was active, and when.

```{r timeline, fig.height = 5}
plot(dn, type = "timeline", top = 25)
```

Ties forming and dissolving, day by day.

```{r turnover}
turnover <- events(dn, measure = c("formation", "dissolution"),
                   start = 0, end = 72, step = 1, window = 1)
plot(turnover)
```

The chapter's proximity timeline.

```{r proximity, fig.width = 8, fig.height = 6}
plot(dn, type = "proximity", slices = 20)
```

## 5. Graph-level measures

A seven-day rolling density.

```{r density}
weekly_density <- metrics(dn, measure = "density",
                          start = 14, end = 60, step = 1, window = 7)
plot(weekly_density)
```

Density over the whole period, and the time-integrated density that accounts
for how long each tie was actually live.

```{r density-scalars}
scalars <- metrics(dn, measure = c("edges", "density", "temporal_density"),
                   window = "all")
scalar_table <- as.data.frame(scalars)
knitr::kable(scalar_table, caption = "Aggregate structure.")
```

Reciprocity, on the full network as in the chapter.

```{r recip}
reciprocity <- metrics(dn_full, measure = "reciprocity",
                       start = 1, end = 73, step = 1, window = 1)
plot(reciprocity)
```

The dyad census. `window = 0` samples at each instant, which is what
`tSnaStats()` does by default.

```{r dyads}
dyad_census <- metrics(dn, measure = c("mutual", "asymmetric", "null"),
                       start = 0, end = 72, step = 1, window = 0)
plot(dyad_census)
```

Degree centralisation.

```{r centralization}
centralisation <- metrics(dn, measure = "centralization_degree",
                          start = 1, end = 73, step = 1, window = 1)
plot(centralisation)
```

## 6. Node-level measures

The chapter runs a `while` loop per centrality. One call takes all three
directions at once.

```{r degree}
degree_series <- dyn_centrality(dn, measure = "degree",
                                mode = c("all", "in", "out"),
                                start = 1, end = 73, step = 1, window = 1)
degree_summary <- summary(degree_series, by = "measure")
knitr::kable(degree_summary, caption = "Degree, in-degree and out-degree per day.")
```

```{r degree-plot, fig.height = 4.5}
degree <- dyn_centrality(dn, measure = "degree",
                         start = 1, end = 73, step = 1, window = 1)
plot(degree, top = 10)
```

Closeness, betweenness and eigenvector, again in one call.

```{r other-centrality}
other_centrality <- dyn_centrality(dn,
                                   measure = c("closeness", "betweenness",
                                               "eigenvector"),
                                   start = 1, end = 73, step = 1, window = 1)
other_summary <- summary(other_centrality, by = "measure")
knitr::kable(other_summary, caption = "Three centralities over the same grid.")
```

```{r betweenness-plot, fig.height = 4.5}
betweenness <- dyn_centrality(dn, measure = "betweenness",
                              start = 1, end = 73, step = 1, window = 1)
plot(betweenness, top = 10)
```

Aggregate centrality can be read straight off the vertex table, so ranking
participants needs no second call.

```{r node-table}
ranked <- as.data.frame(dn, what = "nodes",
                        measure = c("degree", "betweenness", "eigenvector"))
top_ranked <- head(ranked)
knitr::kable(top_ranked, caption = "The vertex table carries aggregate centrality.")
```

## 7. Reachability

The chapter seeds its forward path search on participant 444. Dynet addresses
vertices by name, so the seed is the name.

```{r paths}
forward <- paths(dn, from = "444", direction = "forward")
forward
```

```{r path-plot, fig.width = 8, fig.height = 6}
plot(forward, base_size = 9)
```

The chapter's `transmissionTimeline()` becomes a trajectory tree, which
carries the route counts and branching probabilities the original does not.

```{r trajectories, fig.width = 9, fig.height = 6}
tree <- path_trajectories(forward)
plot_path_trajectories(tree, measure = "time", base_size = 9)
```

## 8. Mixing between experience levels

```{r mixing}
mix <- mixing(dn, attribute = "expert_level",
              start = 1, end = 73, step = 1, window = 1)
mixing_summary <- summary(mix)
mixing_pairs <- head(mixing_summary, 9)
knitr::kable(mixing_pairs,
             caption = "Ties within and between experience levels.")
```

```{r mixing-plot, fig.height = 4.5}
plot(mix)
```

`nodemix` drops one cell as the `ergm` base category, so the chapter plots
eight of the nine group pairs. `mixing()` reports all nine.

## What differs {#what-differs}

Two differences from the published figures, both deliberate.

**The spell rule.** The chapter sets every tie in a discussion to start at the
discussion's *first* post. Saqr and Nouri (2020), which `thread =`
implements, starts each tie at *its own* post. Only 299 of the 2406
surviving posts are opening posts, so the chapter's spells are longer on
average and its time-integrated density is correspondingly higher.

**Ties that never happened.** The chapter builds its static edge list before
dropping single-post discussions, so pairs enter the network with no activity
at all. `networkDynamic()` then gives each of them activity across the entire
observation window, flagged censored at both ends — permanently-on ties that
no message ever created. That is why the chapter reports 433 ties in the
active subnetwork where this reports 428, and why its day-73 statistics are
non-empty for a day past the end of observation.

Neither is a fault in the chapter's argument. Both are worth knowing before
its numbers are reused.

## 9. Do the two implementations agree?

The differences above are definitional, not numerical: they come from how the
network is *built*, not from how it is measured. Given the same network,
Dynet and `tsna`/`sna` should return the same values, and this section checks
that rather than asserting it. The chunks run when the four packages the
chapter uses are installed.

The comparison network is built from Dynet's own spell table, so both sides
measure the same thing. Turning vertex names back into the integers
`networkDynamic` wants is the bookkeeping Dynet hides, and it belongs inside
a function rather than in the analysis.

```{r agree-build, eval = has_tsna, message = FALSE, warning = FALSE}
as_network_dynamic <- function(x) {
  spells <- as.data.frame(x)
  vertices <- as.data.frame(x, what = "nodes")
  ids <- stats::setNames(seq_len(nrow(vertices)), vertices$name)
  base <- network::network.initialize(length(ids), directed = TRUE)
  nd <- networkDynamic::networkDynamic(
    base.net = base,
    edge.spells = data.frame(onset = spells$start, terminus = spells$end,
                             tail = as.integer(ids[spells$from]),
                             head = as.integer(ids[spells$to])),
    verbose = FALSE
  )
  network::set.vertex.attribute(nd, "vertex.names", vertices$name)
  nd
}
nd <- as_network_dynamic(dn)
```

Four quantities, one from each level the chapter works at: a graph-level
scalar, a graph-level series, a node-level series, and a path result. Each
helper reports how many values it lined up as well as the largest difference,
so an empty comparison cannot pass as agreement.

```{r agree-helpers, eval = has_tsna}
graph_agreement <- function(quantity, result, reference) {
  measured <- as.data.frame(result)
  reference <- as.numeric(reference)
  stopifnot("the two series have different lengths" =
              length(reference) == nrow(measured))
  data.frame(quantity = quantity, compared = nrow(measured),
             max_abs_difference = max(abs(measured$value - reference)))
}

node_agreement <- function(quantity, result, reference) {
  measured <- as.data.frame(result)
  reference <- as.matrix(reference)
  flat <- data.frame(
    time = rep(sort(unique(measured$time)), times = ncol(reference)),
    node = rep(colnames(reference), each = nrow(reference)),
    reference = as.numeric(reference), stringsAsFactors = FALSE
  )
  both <- merge(measured, flat, by = c("time", "node"))
  stopifnot("a node-day was lost in the join" = nrow(both) == nrow(measured))
  data.frame(quantity = quantity, compared = nrow(both),
             max_abs_difference = max(abs(both$value - both$reference)))
}

arrival_agreement <- function(quantity, result, reference, vertices) {
  measured <- as.data.frame(result)
  theirs <- stats::setNames(reference$tdist, vertices)
  mine <- stats::setNames(measured$latency, measured$node)
  shared <- intersect(names(theirs)[is.finite(theirs)], names(mine))
  data.frame(quantity = quantity, compared = length(shared),
             max_abs_difference = max(abs(theirs[shared] - mine[shared])))
}
```

```{r agree, eval = has_tsna, message = FALSE, warning = FALSE}
participants <- network::network.vertex.names(nd)

aggregate_density <- metrics(dn, measure = "density", window = "all")
mutual_series <- metrics(dn, measure = "mutual",
                         start = 1, end = 73, step = 1, window = 1)
degree_per_day <- dyn_centrality(dn, measure = "degree",
                                 start = 1, end = 73, step = 1, window = 1)
forward_from_444 <- paths(dn, from = "444")

their_density <- sna::gden(nd)
their_mutual <- tsna::tSnaStats(nd, "mutuality", start = 1, end = 73,
                                time.interval = 1, aggregate.dur = 1)
their_degree <- tsna::tSnaStats(nd, "degree", start = 1, end = 73,
                                time.interval = 1, aggregate.dur = 1,
                                cmode = "freeman")
their_paths <- tsna::tPath(nd, v = match("444", participants), direction = "fwd")

agreement <- rbind(
  graph_agreement("aggregate density", aggregate_density, their_density),
  graph_agreement("mutual dyads per day", mutual_series, their_mutual),
  node_agreement("degree per node per day", degree_per_day, their_degree),
  arrival_agreement("earliest arrival time", forward_from_444, their_paths,
                    participants)
)
knitr::kable(agreement,
             caption = "Dynet against tsna and sna on the same network.")
```

```{r agree-absent, eval = !has_tsna, echo = FALSE, results = "asis"}
cat("*`tsna`, `networkDynamic`, `network` or `sna` is not installed, so the",
    "comparison was skipped.*")
```

## References

Saqr, M. (2024). Temporal network analysis: Introduction, methods and
analysis with R. In M. Saqr & S. López-Pernas (Eds.), *Learning Analytics
Methods and Tutorials*. Springer.
https://doi.org/10.1007/978-3-031-54464-4_17

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge*, 314-319.
