---
title: "Case study: A MOOC discussion forum as a temporal network"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: A MOOC discussion forum as a temporal network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.align = "center",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
```

In this vignette we analyse the discussion forum of a massive open online
course as a temporal network, in the order of the temporal network chapter of
*Learning Analytics Methods and Tutorials* (Saqr, 2024): the network as a
whole, its structure over time, the position of each participant, how far a
post could travel, and how participants of different experience levels
addressed one another.

## Data

`mooc_posts` has one row per post: `sender` and `receiver` are the
participants, `timestamp` the moment of the post and `discussion` the thread.
`mooc_people` has one row per participant with the experience code, 1 for
expert, 2 for student and 3 for teacher, and its label in `expert_level`,
which serves later as the mixing attribute.

```{r data}
library(Dynet)
head(mooc_posts)
head(mooc_people)
```

## The network

Following Saqr and Nouri (2020), a reply is taken to be active from the moment
it is posted until its thread falls silent. Two kinds of rows carry no
interaction: a post that answers its own author, which `dynet()` drops by
default, and a thread that never received a second post, which
`min_thread_posts = 2` drops. To build the network, we call `dynet()` with
`from` and `to` for the two participant columns, `time` for the timestamp,
`thread` for the thread column, which selects the threaded rule, `nodes` for
the participant table so that `expert_level` travels with the network, and
`time_unit = "days"`.

```{r build}
dn_full <- dynet(mooc_posts, from = "sender", to = "receiver",
                 time = "timestamp", thread = "discussion",
                 nodes = mooc_people, time_unit = "days",
                 min_thread_posts = 2)
summary(dn_full)
```

The 86 self-replies and the 37 single-post threads are reported and dropped,
leaving 2,406 posts in 299 threads. The network has 441 vertices and 2,406 spells on 1,907 distinct pairs,
observed from day 0 to day 72.01 in 73 one-day bins. The mean snapshot
density is 0.0035: on an average day about one pair in three hundred is
connected.

Most participants posted a few times. To restrict the analysis to the
participants with more than 20 ties, we call `induce_subgraph()` with a
condition on the vertex table; a centrality named in the condition is
computed over the whole period.

```{r active}
dn <- induce_subgraph(dn_full, degree > 20)
dn
```

The active subnetwork has 45 vertices, 686 spells and 428 pairs, observed
from day 0.11 to day 72.01. The first spell shows the thread rule at work: the
reply from 219 to 444 posted on day 0.11 stays active until day 69.48,
because its thread received posts for that long.

## Visualisation

To draw the aggregate network, we call `plot()` with `type = "network"`.

```{r net-plot, fig.width = 7, fig.height = 6}
plot(dn, type = "network")
```

To draw the network in four one-day bins spaced across the period, with a
shared layout, we set `type` to `"snapshots"` and `panels` to 4.

```{r snapshots, fig.width = 9, fig.height = 3}
plot(dn, type = "snapshots", panels = 4)
```

To count what the first four weeks hold, we call `snapshots()` with a grid of
four one-week windows beginning on day 1, and `summary()` on the result.
`ties` is the number of distinct pairs active in the window, `nodes` the
number of participants with at least one active tie, and `weight` the number
of spells.

```{r snapshot-counts}
weekly <- snapshots(dn, start = 1, end = 22, step = 7, window = 7)
summary(weekly)
```

The week beginning on day 1 holds 58 pairs among 29 participants; the week
beginning on day 22 holds 154 pairs among 44 of the 45. The counts grow
because a reply stays active until its thread closes, so early ties are still
present in later windows.

To see when each participant was active, we set `type` to `"timeline"` and
`top` to the number of participants to show.

```{r timeline, fig.height = 5}
plot(dn, type = "timeline", top = 25)
```

To count the ties that form and dissolve each day, we call `events()` with
`measure` for the two counts and a grid of one-day bins.

```{r turnover}
turnover <- events(dn, measure = c("formation", "dissolution"),
                   start = 0, end = 72, step = 1, window = 1)
summary(turnover)
plot(turnover)
```

On average 9.4 ties form and 9.4 dissolve per day; the means are equal because
every spell begins and ends inside the period. Formation peaks at 40 on day
46. Dissolution peaks at 99 on day 69, when the threads that had stayed open
through the course fall silent together.

The proximity timeline places participants who interact near each other and
follows their positions through time. To draw it, we set `type` to
`"proximity"` and `slices` to the number of time slices.

```{r proximity, fig.width = 8, fig.height = 6}
plot(dn, type = "proximity", slices = 20)
```

## Structure over time

Every measuring function takes `start`, `end`, `step` and `window`. To measure
density over a rolling week, we call `metrics()` with `step = 1` and
`window = 7`, so that each value covers the seven days beginning at its time.

```{r density}
weekly_density <- metrics(dn, measure = "density",
                          start = 14, end = 60, step = 1, window = 7)
summary(weekly_density)
plot(weekly_density)
```

Over the 47 windows the weekly density averages 0.101 and ranges from 0.066
to 0.139, with the peak in the week beginning on day 45.

To obtain the density of the whole period, and the temporal density that
weights each pair by the share of the period during which it was live, we set
`window` to `"all"`.

```{r density-scalars}
scalars <- metrics(dn, measure = c("edges", "density", "temporal_density"),
                   window = "all")
scalars
```

The 428 pairs give an aggregate density of 0.216. The temporal density is
0.063, less than a third of it, because most pairs were live for only part of
the period.

Reciprocity is the share of ties that are returned. To obtain it per day on
the full network, we call `metrics()` with `measure = "reciprocity"`.

```{r recip}
reciprocity <- metrics(dn_full, measure = "reciprocity",
                       start = 1, end = 73, step = 1, window = 1)
summary(reciprocity)
plot(reciprocity)
```

Reciprocity averages 0.151 over the 73 days and peaks at 0.218 on day 47.

The dyad census classifies every pair as mutual, asymmetric or null (Wasserman
and Faust, 1994). To obtain it at each instant of the grid, we name the three
classes in `measure` and set `window` to 0.

```{r dyads}
dyad_census <- metrics(dn, measure = c("mutual", "asymmetric", "null"),
                       start = 0, end = 72, step = 1, window = 0)
summary(dyad_census)
plot(dyad_census)
```

The 45 vertices form 990 dyads, all null at the first instant. Over the 73
instants an average of 21.0 dyads are mutual and 81.5 asymmetric, and the
mutual count peaks at 46 on day 48.

Degree centralisation is 0 when every vertex has the same degree and 1 when
one vertex holds every tie (Freeman, 1979). To obtain it per day, we set
`measure` to `"centralization_degree"`.

```{r centralization}
centralisation <- metrics(dn, measure = "centralization_degree",
                          start = 1, end = 73, step = 1, window = 1)
summary(centralisation)
plot(centralisation)
```

Centralisation averages 0.379 and peaks at 0.508 on day 31, the day on which
one participant's betweenness also peaks.

## Participants

To obtain degree in all three directions for every participant on every day,
we call `dyn_centrality()` with `measure = "degree"` and `mode` for the
directions, then `summary()` with `by = "measure"` for one row per measure.

```{r degree}
degree_series <- dyn_centrality(dn, measure = "degree",
                                mode = c("all", "in", "out"),
                                start = 1, end = 73, step = 1, window = 1)
summary(degree_series, by = "measure")
```

Mean degree is 5.77 per participant per day, with a maximum of 52 on day 49.
In-degree and out-degree share the mean of 2.88, but in-degree is far more
dispersed: its maximum is 35 on day 50 against 19 for out-degree on day 31.
Replies concentrate on a few receivers more than on a few senders.

To draw the trajectories of the ten most central participants, we call
`plot()` with `top`.

```{r degree-plot, fig.height = 4.5}
degree <- dyn_centrality(dn, measure = "degree",
                         start = 1, end = 73, step = 1, window = 1)
plot(degree, top = 10)
```

Closeness, betweenness and eigenvector centrality describe reach, brokerage
and connection to well-connected others (Freeman, 1979; Bonacich, 1987). To
obtain all three on the same grid, we name them in `measure`.

```{r other-centrality}
other_centrality <- dyn_centrality(dn,
                                   measure = c("closeness", "betweenness",
                                               "eigenvector"),
                                   start = 1, end = 73, step = 1, window = 1)
summary(other_centrality, by = "measure")
```

Betweenness averages 34.6 and peaks at 974.1 on day 31. Closeness averages
0.42 and eigenvector centrality 0.24; both reach their bound of 1 on day 1.

```{r betweenness-plot, fig.height = 4.5}
betweenness <- dyn_centrality(dn, measure = "betweenness",
                              start = 1, end = 73, step = 1, window = 1)
plot(betweenness, top = 10)
```

Aggregate centrality over the whole period is carried by the vertex table. To
obtain it, we call `as.data.frame()` with `what = "nodes"` and `measure` for
the centralities to annotate.

```{r node-table}
ranked <- as.data.frame(dn, what = "nodes",
                        measure = c("degree", "betweenness", "eigenvector"))
head(ranked)
```

Among the first six participants, 11, a teacher, has the highest aggregate
degree, 47, the highest betweenness, 180.5, and the highest eigenvector
centrality, 0.896.

## Reachability

A time-respecting path may only use ties in chronological order (Kempe,
Kleinberg and Kumar, 2002), so it shows how far, and how fast, something
posted by one participant could travel through the forum. To find the
earliest path from participant 444 to every other, we call `paths()` with
`from` for the source. `arrival_time` is the earliest moment a participant
can be reached, `latency` that moment minus the source's start, `n_hops` the
number of ties on the earliest route and `n_paths` the number of routes that
arrive equally early.

```{r paths}
forward <- paths(dn, from = "444", direction = "forward")
forward
summary(forward)
```

Participant 444 reaches all 44 others. Participant 19 is reached in one hop
after 2.11 days and participant 11 in one hop after 12.27 days; participant 15
needs five hops but is reached after 6.02 days, earlier than participant 11.
The median latency is 17.6 days and the maximum 50.9; the median route has two
hops and the longest five.

To draw the earliest routes as a tree, we call `plot()` on the result; to draw
them against time, we call `path_trajectories()` on the result and
`plot_path_trajectories()` with `measure = "time"`.

```{r path-plot, fig.width = 8, fig.height = 6}
plot(forward, base_size = 9)
```

```{r trajectories, fig.width = 9, fig.height = 6}
tree <- path_trajectories(forward)
plot_path_trajectories(tree, measure = "time", base_size = 9)
```

## Mixing between experience levels

To count the ties within and between experience levels per day, we call
`mixing()` with `attribute` for the vertex attribute that defines the groups,
then `summary()` for one row per ordered pair of levels. Each measure reads
from the sender's level to the receiver's (Newman, 2003).

```{r mixing}
mix <- mixing(dn, attribute = "expert_level",
              start = 1, end = 73, step = 1, window = 1)
summary(mix)
plot(mix)
```

Teachers writing to teachers is the most common tie, 36.8 active ties per day
on average with a peak of 67, followed by students writing to teachers at
33.5. Experts writing to experts is the rarest, at 2.9 per day. Direction
matters: ties from experts to teachers average 14.3 per day, from teachers to
experts 5.8.

## Interpretation

The forum is sparse on any given day and dense in aggregate: one pair in three
hundred is connected on an average day, one in five over the course, and the
gap between the temporal density of 0.063 and the aggregate density of 0.216
says that most ties were live for only a fraction of the period. Activity is
concentrated on receivers rather than senders, and on teachers: the most
common exchange is between teachers, students address teachers more than
anyone else, and the most central participant over the whole period is a
teacher. Everything posted by participant 444 could in principle reach every
active participant, but the median took more than two weeks and the longest
route seven weeks, so the forum spreads information slowly even though it
connects everyone.

## Limitations

A reply that stays active until its thread closes is a modelling choice; a
shorter tie life would give sparser windows and longer latencies. The
analysis of structure and centrality is restricted to the 45 participants
with more than 20 ties, so it describes the active core, not the course.
Participants who wrote or received no surviving exchange are absent, and the
observation period is taken from the data rather than the course calendar.

## References

Bonacich, P. (1987). Power and centrality: A family of measures. *American
Journal of Sociology*, 92(5), 1170–1182.

Freeman, L. C. (1979). Centrality in social networks: Conceptual
clarification. *Social Networks*, 1(3), 215–239.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.

Newman, M. E. J. (2003). Mixing patterns in networks. *Physical Review E*,
67(2), 026126.

Saqr, M. (2024). Temporal network analysis: Introduction, methods and
analysis with R. In M. Saqr & S. López-Pernas (Eds.), *Learning analytics
methods and tutorials: A practical guide using R*. Springer.
https://doi.org/10.1007/978-3-031-54464-4_17

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319).
ACM.

Wasserman, S., & Faust, K. (1994). *Social network analysis: Methods and
applications*. Cambridge University Press.
