---
title: "Getting started with Dynet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with Dynet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5)
library(Dynet)
```

# Getting started with Dynet

Temporal networks map the dynamics of relationships as they occur over time, preserving when interactions occur, their duration, and their temporal order. Two common representations are **contact sequences**, in which interactions are treated as instantaneous events, and **interval networks**, in which relationships have onset and termination times. Dynet represents these relationships as relational spells. Each spell identifies two relational endpoints and their period of connection; instantaneous contacts have equal onset and termination times.

Temporal order determines which paths are available between vertices. For example, if A shares information with B on Monday and B interacts with C on Tuesday, the information can potentially reach C through B. An interaction between B and C on Sunday could not carry information received on Monday. A sequence that follows the order and availability of interactions is called a **time-respecting path**.

A static network combines relationships without preserving this order. Binary aggregation also discards their duration and multiplicity, although weighted aggregation can retain summaries of these quantities. Consequently, a path in the aggregate network may not be temporally traversable. Relationships also begin and end at different times, so a network that appears densely connected in aggregate may contain periods of limited connectivity.

This vignette demonstrates these distinctions using simulated classroom contacts. The analysis covers network construction, graph-level measures over time, participants’ centrality trajectories, time-respecting paths, and the timing and duration of relational spells.

## Data

`school_contacts` is a simulated dataset of directed face-to-face contacts among fourteen students, supplied in tidy format. The variables `from` and `to` identify the student initiating a contact and the student addressed. The variables `start` and `end` record contact onset and termination. In this example, numeric times are interpreted as days since the beginning of observation; decimal values allow contacts to begin and end within a day.

```{r data}
head(school_contacts)
```

## Building the network

`dynet()` constructs a temporal network from the supplied relational data. Because `school_contacts` contains recognised endpoint and interval-boundary columns, the constructor identifies it as interval data without requiring explicit column arguments.

Column recognition is case-insensitive. Endpoint names such as `from`/`to`, `sender`/`receiver`, and `source`/`target` are recognised, as are `start`/`end` and `onset`/`terminus` for interval boundaries. Explicit column specification is needed only when names do not match recognised aliases or their intended interpretation is ambiguous. A duration may also be supplied in place of `end`, in which case termination is calculated as `start + duration`.

```{r build}
dn <- dynet(school_contacts)
dn
```

The constructor records `duration = end - start` and assigns `weight = 1` when no multiplicity variable is supplied or recognised. Positive-duration spells are active on the half-open interval $[\text{start}, \text{end})$: onset is included, while termination is excluded. Two spells that meet at a boundary therefore do not overlap at that instant.

The resulting network contains 14 vertices, 240 relational spells, and 110 distinct ordered pairs, observed from time 0 to 21.52. Numeric input retains its supplied scale and is labelled `step`; here, one step represents one day. The default construction interval of 1 provides the one-day measurement intervals used below.

Because the network is directed, a relationship from Ana to Cara is distinct from a relationship from Cara to Ana. With 14 vertices and self-links excluded, there are $14 \times 13 = 182$ possible ordered pairs. Of these, 110—approximately 60%—are connected at least once during the observation period.

`summary()` reports network properties in tabular form, including the number of measurement bins and mean snapshot density.

```{r summary}
summary(dn)
```

The observation period covers 22 daily bins, with the final bin ending at time 21.52. Mean snapshot density is 0.0829: approximately 8.3% of possible directed connections are present in an average bin, compared with about 60% across the full period. This difference shows how the aggregate network combines relationships that occur at different times.

The activity plot displays spell onsets, terminations, and active-spell counts over time, providing an initial view of changes in relational activity.

```{r activity}
plot(dn, type = "activity")
```

## Measuring in windows

Graph-level measures describe the structure of the network as a whole. Calculating them within successive temporal windows shows how connectivity changes during the observation period. For each window, Dynet identifies the connections active during that interval and computes the requested measures on the resulting snapshot.

Four arguments control the timing of measurements. `start` and `end` specify the first and last measurement times. `step` determines the interval between measurements and defaults to the network’s construction interval. `window` specifies the duration covered by each measurement, beginning at its reported time. By default, `window` equals `step`, producing non-overlapping intervals. A larger `window` produces overlapping measurements, while `window = 0` evaluates connectivity at individual time points.

Window width determines the temporal detail of the analysis. Short windows distinguish changes over brief periods but may contain few connections. Longer windows combine more relationships and provide a broader summary, while obscuring changes within each interval. Connections included in the same window need not all be active simultaneously.

### Daily density

**Density** is the proportion of possible connections present in the network. For a directed network with $n$ vertices and no self-links, density within window $w$ is

$$D_w = \frac{m_w}{n(n-1)},$$

where $m_w$ is the number of ordered vertex pairs with at least one active spell during the window. Repeated or overlapping spells between the same endpoints contribute one connection, so density ranges from 0 to 1.

The following call calculates density using the default one-day intervals:

```{r density}
density <- metrics(dn, measure = "density")
density
```

The result is a tidy data frame containing `time`, the beginning of the measurement window; `measure`, the requested statistic; and `value`, its calculated value.

On day 0, density is 0.055, corresponding to 10 of the 182 possible ordered pairs. On day 6, it increases to 0.159, corresponding to 29 pairs. These counts include every pair connected at some point during the respective day.

`summary()` describes the resulting time series. It reports the number of measurements (`n`), mean, standard deviation (`sd`), minimum, maximum, and `peak_time`, which identifies the beginning of the window with the highest value.

```{r density-summary}
summary(density)
```

Across the 22 bins, mean density is 0.083 (SD 0.039). Density peaks at 0.165 on day 14, when 30 ordered pairs are connected. The minimum, 0.033, occurs in the final bin, which covers only the period from day 21 to day 21.52. Because this interval is shorter than a full day, it provides less time for contacts to occur. Among complete daily intervals, the minimum density is 0.038 on day 18.

### Rolling weekly density

A rolling window describes connectivity over a longer period while retaining frequent measurements. Setting `step = 1` and `window = 7` calculates density daily using the seven-day interval beginning at each measurement time. Successive windows overlap, so adjacent values share much of their underlying data.

```{r density-rolling}
density_weekly <- metrics(dn, measure = "density", step = 1, window = 7)
plot(density_weekly)
```

A pair contributes to weekly density if it is connected at any time during that seven-day interval. With the same vertex population, weekly density cannot be lower than daily density measured from the same starting time, because the weekly window includes the daily window.

Windows beginning after day 14.52 extend beyond the observation period and therefore contain fewer than seven observed days. The final values should be interpreted with this decreasing observation duration in mind.

## Vertex trajectories

Graph-level measures describe the network as a whole, whereas vertex-level centrality measures describe individual positions within it. `centrality_series()` computes the selected measures for each vertex within successive temporal windows. The resulting trajectories show when participants become more or less connected. The arguments `start`, `end`, `step`, and `window` specify the measurement intervals in the same way as for `metrics()`.

### Daily degree

**Degree** counts direct connections within a measurement window. In a directed network, indegree counts distinct vertices with incoming ties to the focal vertex, while outdegree counts distinct vertices reached by its outgoing ties. The default `mode = "all"` adds these two quantities. A reciprocated relationship therefore contributes twice to total degree, even though it involves a single partner. Repeated spells in the same direction between two vertices count once within a window.

```{r degree}
degree <- centrality_series(dn, measure = "degree")
degree
```

The result is a tidy data frame containing `time`, `node`, `measure`, and `value`. With fourteen students and 22 measurement intervals, the degree series contains 308 observations. Each value describes the connections active within the corresponding daily interval, rather than the number of individual contact events.

`summary()` summarises each participant’s trajectory, reporting the number of measurements, mean, standard deviation, minimum, maximum, and the time at which degree reaches its maximum.

```{r degree-summary}
summary(degree)
```

Mean daily total degree ranges from 1.64 to 2.86. The trajectories nevertheless differ in their variability and timing:

- **Relatively consistent connectivity.** Two students have at least one connection in every daily interval. The three smallest standard deviations range from 1.07 to 1.26, with maximum degrees between 4 and 6.
- **Episodic connectivity.** Other students alternate between intervals without connections and intervals with substantially higher degree. Five students reach maxima of 7 or 8, with standard deviations between 1.78 and 2.17. Their peaks occur on days 6, 13, and 14.

These descriptions represent variation along a continuum, rather than formally identified groups. Similar mean degrees can arise from different temporal patterns, so the mean alone does not describe when opportunities for interaction occur.

The heatmap displays all fourteen trajectories, with colour indicating degree for each participant and measurement interval.

```{r degree-heatmap}
plot(degree, type = "heatmap")
```

## Temporal centrality

Temporal centrality accounts for the timing and order of interactions when measuring a vertex’s position in the network. Temporal closeness measures how quickly a participant can reach others through time-respecting paths. Temporal betweenness measures the extent to which a participant acts as an intermediary on those paths. In Dynet, these measures are calculated with `path_centrality()`.

**Temporal closeness** measures how quickly a participant can reach others through time-respecting paths. Starting at the beginning of the observation period, Dynet determines the earliest time each reachable participant can be reached. The elapsed time includes waiting for subsequent interactions along the path. Temporal closeness is the inverse of the mean elapsed time across reachable participants; higher values indicate earlier reachability.

```{r closeness}
closeness <- path_centrality(dn, measure = "closeness")
closeness
```

In this example, temporal closeness ranges from 0.131 to 0.375, corresponding to mean elapsed times of approximately 7.6 and 2.7 days, respectively. Unreachable participants are excluded from the mean, so the measure describes the speed of reaching those who are reachable.

**Temporal betweenness** measures how often a participant acts as an intermediary on time-respecting paths between other participants. Dynet considers paths that arrive earliest and, among those arriving at the same time, use the fewest interactions. For each reachable ordered pair, a participant receives a contribution equal to the proportion of these paths that pass through them. Temporal betweenness sums these contributions across pairs.

```{r betweenness}
betweenness <- path_centrality(dn, measure = "betweenness")
betweenness
```

Here, temporal betweenness ranges from approximately 4.3 to 41.9. Higher values indicate a greater role in connecting others through these paths. The values are unnormalised sums, rather than percentages or counts of interactions. They describe potential routes through the observed network; they do not establish that information actually travelled along those routes.

## Time-respecting paths

`paths()` identifies which vertices can be reached from a specified source through time-respecting paths. The search begins at the start of the observation period unless `start` is specified. For each reachable destination, it identifies the earliest arrival time and then the fewest interactions needed to arrive at that time. These are called **shortest foremost paths**.

The following call searches for paths from Ana:

```{r paths}
from_ana <- paths(dn, from = "Ana")
from_ana
```

The result is a tidy data frame describing reachability from Ana to each vertex. `reachable` indicates whether a time-respecting path exists, and `arrival_time` records the earliest arrival. `latency` measures the elapsed time from the search start to arrival, including waiting between interactions. `n_hops` gives the number of interactions along a shortest foremost path, and `n_paths` counts the distinct paths satisfying these criteria.

Starting on day 0, Ana can reach all thirteen other students: three through one interaction, four through two, four through three, and two through four. Fewer interactions do not necessarily imply earlier arrival. For example, a two-hop path arrives on day 6.12, before two direct contacts become available on days 6.36 and 6.67. Arrival therefore depends on when the interactions occur as well as how they connect participants.

`summary()` reports the number and proportion of other vertices reached, together with the median and maximum latency and hop count.

```{r paths-summary}
summary(from_ana)
```

Ana can reach all thirteen other students, giving a reachable proportion of 1. The median arrival latency is 7.51 days and the maximum is 11.66 days. The median hop count is 2, and the maximum is 4. Although daily density never exceeds 0.165, interactions occurring in sequence across days allow Ana to reach every other student within twelve days.

Reachability also depends on when the search begins. Setting `start = 18` restricts the search to paths available from day 18 onward. Spells that ended before this time cannot contribute, while spells still active at the search start remain available.

```{r paths-late}
from_ana_late <- paths(dn, from = "Ana", start = 18)
summary(from_ana_late)
```

From day 18, Ana can reach five of the thirteen other students, giving a reachable proportion of 0.385. Among these students, the maximum arrival latency is 2.68 days and the maximum hop count is 3. Thus, the same participant can reach different sets of people depending on the search start and the interactions available during the remaining observation period.

A backward search identifies which participants could reach a specified vertex by a given deadline. With `direction = "backward"`, the default deadline is the end of the observation period. For each participant, `paths()` determines the latest departure time that would allow a time-respecting path to reach Ana by that deadline.

```{r paths-backward}
into_ana <- paths(dn, from = "Ana", direction = "backward")
summary(into_ana)
```

In backward searches, `arrival_time` records the latest departure boundary, and `latency` is the elapsed time from that boundary to the deadline. Because relational spells exclude their termination time, departure may be possible arbitrarily close to this boundary but not exactly at it. In such cases, `attained = FALSE`; the participant is nevertheless reachable in the backward search.

All thirteen other students could reach Ana by day 21.52, the end of observation. The median backward latency is 4.27 days and the maximum is 8.9 days. These values describe how far before the deadline participants would need to depart along the available paths.

## Routes

Several shortest foremost paths can follow the same sequence of vertices while using different relational spells. `pathways()` groups these paths by their vertex sequence, called a **route**, and counts the paths following each route.

```{r pathways}
pathways(dn, from = "Ana")
```

The result is a tidy data frame. `route` gives the sequence of vertices, `endpoint` identifies the destination, and `count` records the number of shortest foremost paths following that route. `share` gives the route’s proportion of all counted paths in the result. `n_hops` records the number of interactions along the route, and `arrival_time` gives the earliest arrival at its destination. Routes are ordered by decreasing count.

These counts describe the available paths through the observed interactions. They do not measure how often participants actually transmitted information along a route.

In this example, `pathways()` reports five routes containing seven shortest foremost paths. One route accounts for three paths (share 0.429) and ends at the last student reached, on day 11.66. These paths follow the same sequence of students through different relational spells. Each of the other four routes accounts for one path.

The reported routes end at leaves of the path tree. Paths that terminate at intermediate vertices are represented within longer routes rather than listed separately.

Calling `plot()` on the `paths()` result displays the paths as a tree:

```{r tree, fig.height = 5}
plot(from_ana)
```

Ana is the root, and each successive level adds one interaction. Node size and branch width indicate the number of shortest foremost paths using that part of the tree. Labels identify the participant and path count.

Each tree node represents a participant reached at a particular time. The same participant can therefore appear more than once when paths reach them at different times. These arrival times determine which subsequent interactions remain available to extend each path.

`path_trajectories()` returns the path tree in tidy format, allowing its branches and counts to be examined directly.

```{r trajectories}
tree <- path_trajectories(from_ana)
tree
```

`node` identifies a tree node by the sequence of `vertex@time` steps leading to it, and `parent` identifies the preceding node. `depth` gives the number of interactions from the source. `vertex` and `time` record the participant and arrival time separately.

`count` gives the number of shortest foremost paths passing through or ending at a tree node. `probability` divides this count by the parent’s count, describing the proportion of the parent’s paths that continue along that branch. It is a proportion of counted paths, rather than an estimated probability of information transmission.

The tree contains 22 nodes. Its root, `Ana@0`, has a count of 19, representing all shortest foremost paths reported by `paths()`, including the zero-hop path at Ana. One first-hop branch accounts for seven paths and subsequently reaches six additional students. Another student appears on three first-hop branches, with arrival times of 2.12, 3.43, and 6.68 days. These branches show how different contact times can support paths through the same participants.

A shortest foremost path need not reach every intermediate participant at their earliest possible time. In this example, one path reaches its second participant on day 6.68, although that participant is reachable by day 6.12 through another path. Both arrivals nevertheless allow the next participant to be reached at their earliest arrival time, day 9.59. Arriving earlier at an intermediate participant therefore does not necessarily produce an earlier arrival at the destination.

## Tie dynamics

Tie dynamics describe when relational spells begin and end, how long they last, and how they recur over time. `events()` counts spell onsets (**formation**) and terminations (**dissolution**) within each measurement window. With the default `window = step`, windows do not overlap, so each onset and termination within the measurement period is counted once.

```{r events}
turnover <- events(dn)
summary(turnover)
```

Both series average 10.91 events per measurement bin: all 240 spells begin and end within the observation period, which spans 22 bins. Formation peaks at 29 spells on day 13, while dissolution peaks at 27 on day 14, the day with the highest density. At least one bin contains no onsets, whereas every bin contains at least three terminations.

`durations()` summarises the duration of relational spells for each ordered pair of vertices, with spell duration calculated as `end - start`. Two pairs may have the same number of spells but differ substantially in how long their connections last.

Setting `measure = "mean"` returns the mean spell duration for each pair. Other options include `"events"` for the number of spells, `"total"` for their summed duration, and `"median"` for their median duration.

```{r durations}
lengths <- durations(dn, measure = "mean")
lengths
```

The result is a tidy data frame containing `from`, `to`, `measure`, and `value`. Among pairs with Ana as the initiating student, mean spell duration ranges from 0.10 to 1.19 days. Weighting an aggregate network by spell counts alone would omit this variation.

Ana has direct contacts with eight students during the observation period, but shortest foremost paths reach only three of them directly. The other five are reached earlier through paths involving two to four interactions, before their direct contacts with Ana begin. A direct connection in the aggregate network therefore need not provide the earliest route to its destination.

**Burstiness** describes variation in the intervals between successive events. Here, an event is the onset of a spell involving a participant as either `from` or `to`. Let $\mu$ be the mean interval between consecutive events and $\sigma$ its population standard deviation. Burstiness is calculated as

$$B = \frac{\sigma - \mu}{\sigma + \mu}.$$

For equal, positive intervals, $B = -1$. Positive values indicate greater variation in spacing, consistent with closely spaced events separated by longer gaps. A value of 0 indicates that the standard deviation equals the mean, as in the theoretical exponential waiting-time distribution of a Poisson process. A value near 0 alone does not establish that events follow a Poisson process.

The **memory coefficient**, $M$, is the Pearson correlation between consecutive intervals. Positive values indicate that short intervals tend to follow short intervals and long intervals tend to follow long intervals. Negative values indicate a tendency for short and long intervals to alternate.

```{r burstiness}
rhythm <- burstiness(dn)
summary(rhythm)
```

Burstiness ranges from −0.070 to 0.246, with eight students having positive values and six having negative values. Memory ranges from −0.227 to 0.480 and is positive for nine of the fourteen students. Participants have between 28 and 46 recorded spell onsets. These measures describe different aspects of interaction timing: burstiness captures variation in event spacing, while memory captures the association between successive intervals.

The results illustrate how interaction frequency, direct connectivity, and temporal position can differ. The student with the fewest events (28) and the lowest mean daily degree has the highest temporal closeness and the third-highest temporal betweenness. Conversely, the student with the most events (46) and the highest mean daily degree has the second-lowest temporal betweenness.

Frequent interaction therefore does not necessarily imply earlier reachability or a greater intermediary role. These temporal properties depend on how a participant’s interactions connect with those of others in time. Examining event counts, degree trajectories, and temporal centrality together provides a fuller description of participation than any single measure.

## Next steps

`vignette("building-networks")` explains how to construct temporal networks from contact, threaded, and co-presence data. It also covers direction, self-links, weights, vertex attributes, sessions, observation periods, and vertex activity spells. `vignette("ch17-temporal-networks")` demonstrates an analysis of a MOOC discussion forum using Dynet.

`?metrics` and `?centrality_series` document the available graph-level and vertex-level measures. `?paths` describes the traversal rules, including `traversal_time` for specifying a fixed duration per interaction and `as.data.frame(x, what = "steps")` for inspecting the individual steps of each path.
