---
title: "Getting started with Dynet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with Dynet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(Dynet)
```

## A network that knows what time it is

A static network says that two people are connected. A temporal network says
*when* they were connected, *for how long*, and — the part that matters most —
*in which order*.

That ordering is not decoration. If Ana talks to Jonas on Tuesday and Jonas
talks to Eve on Wednesday, something can travel Ana → Jonas → Eve. If the same
two conversations happen in the opposite order, nothing can. A flattened
network scores both situations identically: two edges, one path of length two.
Only one of them is real.

The usual workaround is to cut time into slices and analyse each slice as a
static network. That helps, but it throws away exactly the thing that makes the
data temporal: a snapshot cannot contain a chain that spans snapshots. In the
classroom data below, Ana and Eve never once make contact, so no slice will
ever link them — yet a time-respecting path carries Ana to Eve in four hops.
Aggregate instead and you make the opposite mistake: every ordering becomes
legal, and connectivity is inflated.

Dynet keeps time in the object and in every result. This vignette is the tour.

## Building a network in one line

The bundled `school_contacts` is an *interval log*: each row is one contact,
with its own start and end time.

```{r build}
dn <- dynet(school_contacts)
dn
```

`dynet()` inferred the format from the column names it found. `summary()`
describes what it built.

```{r summary}
summary(dn)
```

Fourteen pupils, 240 contact spells over 110 distinct pairs, observed across
about 21.5 time steps. Note the last two rows of that table: the *mean snapshot
density* and the *temporal density* are not the same number, and the gap
between them is the subject of this vignette.

One constructor covers the four shapes relational logs actually arrive in —
interval logs, contact logs of instantaneous events, threaded logs such as
forum data, and co-presence logs where sharing a group creates a tie. The
companion vignette on building networks works through all four; here we stay
with the one line above.

```{r activity, fig.width = 7, fig.height = 4.5}
plot(dn, type = "activity")
```

## Two commitments

Everything below rests on two promises the package makes.

**Vertices are named, never numbered.** You ask for `"Ana"`, not for vertex 1.
There is no index to look up first, and no result that hands you an integer
where you expected a person.

**Every verb returns a tidy data frame.** One row per observation, with the
columns you would expect. No matrices, no nested lists, nothing that asks you
to reach into an object with `$` or to subset a result before it is readable.

Both are visible in a single call:

```{r commitments}
head(dyn_centrality(dn, measure = "degree"))
```

One row per node per time point, vertices named, a `value` column you can plot
directly. The printed header is a courtesy; the object underneath is an
ordinary data frame.

```{r tidy}
str(as.data.frame(metrics(dn, measure = "density")))
```

Because results are already tidy, the tidying rituals you would normally write
by hand are arguments instead. `start`, `end`, `step`, `window`, `sessions`,
`measure`, `mode` — you narrow a result by asking the verb for it, not by
slicing what it returned.

## What a snapshot cannot see

Start with the graph-level view. `metrics()` computes a structural measure once
per time bin.

```{r density}
head(metrics(dn, measure = "density"))
```

```{r density-summary}
summary(metrics(dn, measure = "density"))
```

The busiest single instant links about one pair in six; on average it is closer
to one in twelve. Plotted, the classroom breathes.

```{r density-plot, fig.width = 7, fig.height = 4}
plot(metrics(dn, measure = "density"), palette = "okabe")
```

If those snapshots were all you had, you would conclude that this classroom is
sparse and fragmented. Now ask the temporal question instead.

## Time-respecting paths

A time-respecting path may only use edges whose timing runs forward. It can
wait at a vertex, but it can never travel back in time. `paths()` follows every
such path out of one named vertex.

```{r paths}
paths(dn, from = "Ana")
```

Four columns carry the temporal content:

- **`arrival_time`** — the earliest clock time at which the journey can get
  there. It is a point on the network's timeline, not a count of steps.
- **`latency`** — how long the journey took: `arrival_time` minus the time the
  source became available. Ana starts at 0 here, so the two coincide; start
  later and they part company.
- **`n_hops`** — how many edges the journey used. Dynet optimises *shortest
  foremost*: it minimises arrival time first, then hop count among the
  journeys that achieve it.
- **`n_paths`** — how many distinct optimal journeys there are. Two vertices
  can share an arrival time and hop count and still differ in how many ways
  the network gets you there.

`summary()` aggregates the whole reachable set.

```{r paths-summary}
summary(paths(dn, from = "Ana"))
```

So Ana reaches every one of the other thirteen pupils, in at most four hops —
in a network where no single snapshot ever connected more than a sixth of the
pairs. That is the information a snapshot-by-snapshot analysis destroys.

The result is not a property of the network alone; it is a property of the
network *and when you start*. Ask the same question from late in the window and
most of those journeys no longer exist, because the contacts that carried them
have already happened.

```{r paths-late}
paths(dn, from = "Ana", start = 18)
```

Reach also has a direction. `direction = "backward"` traces who could have
reached Ana rather than whom Ana can reach — the question you want when you are
tracing an exposure backwards rather than forwards.

```{r paths-backward}
summary(paths(dn, from = "Ana", direction = "backward"))
```

## Seeing the journeys

`plot()` on a paths result draws the whole family of optimal journeys as a
trajectory tree.

```{r trajectory, fig.width = 7, fig.height = 5}
journeys <- paths(dn, from = "Ana")
plot(journeys)
```

Read it outward from the root. The source sits at hop 0 on the left; each step
to the right is one more hop, so a vertex's horizontal position is the number
of edges its journey used. Node size and branch width both encode the number of
optimal routes running through that point, and each node is labelled with the
vertex name and the time the hop fires, so nothing is carried by size or colour
alone.

The feature that has no static counterpart is this: **a vertex appears more than
once when it was reached under a different temporal history.** `Jonas` shows up
three times in this tree. Those are not three different pupils and not three
different places in the graph — they are the same vertex entered at 2.12, at
3.43 and at 6.68. Each entry time leaves a different set of onward contacts
still in the future, so each one is genuinely a different starting point for
the rest of the journey. That is why `paths()` reported `n_paths = 3` for Ben
earlier: the three optimal routes to Ben share the vertex sequence
Ana → Jonas → Kira → Ben and differ only in when the hops fire. A flattened
network would collapse all three into one path and call the matter settled.

`measure` re-encodes the same tree. `"time"` weights it by arrival time rather
than route count, which turns the picture into a schedule of when the reachable
set fills up.

```{r trajectory-time, fig.width = 7, fig.height = 5}
plot(journeys, measure = "time")
```

`orientation = "vertical"` turns the tree on its side for a tall, narrow layout,
`min_count` prunes branches carrying fewer than that many routes, and
`measure = "predictability"` shades each branch by how strongly its parent
commits to it.

The tree is data as well as a picture. `path_trajectories()` returns it tidily,
one row per tree node, with the route as a readable string, its `parent`,
`depth`, the `count` of routes through it, the branching `probability` given its
parent, and the bare `vertex` and `time`.

```{r trajectories}
head(path_trajectories(journeys))
```

## A tour of the measurement verbs

Every verb takes the network first and a `measure` (or an attribute) by name.
Every one returns a tidy frame keyed by time, by vertex, or by both.

### `dyn_centrality()` — who matters, and when

At `scope = "snapshot"` a centrality is recomputed in every bin, so `summary()`
gives you each pupil's trajectory in one table, including when they peaked.

```{r centrality}
summary(dyn_centrality(dn, measure = "degree"))
```

```{r centrality-plot, fig.width = 7, fig.height = 4.5}
plot(dyn_centrality(dn, measure = "degree"), type = "heatmap", palette = "okabe")
```

At `scope = "temporal"` the measure is computed on time-respecting paths across
the whole window instead — one value per vertex, not one per bin. Temporal
closeness is the *inverse* of mean forward latency over the vertices a person
reaches, so a high score means "gets to the rest of the class quickly" rather
than "has many neighbours right now".

```{r centrality-temporal}
head(dyn_centrality(dn, measure = "closeness", scope = "temporal"))
```

### `metrics()` — graph-level structure over time

Density is one of forty available measures. `temporal_density` is a
useful contrast: it discounts pairs that are adjacent in a snapshot but never
in a time-respecting order.

```{r metrics-temporal}
head(metrics(dn, measure = "temporal_density"))
```

### `events()` — ties appearing and disappearing

Structure is a stock; formation and dissolution are the flows that produce it.

```{r events}
head(events(dn, measure = "formation"))
```

```{r events-dissolution}
head(events(dn, measure = "dissolution"))
```

### `burstiness()` — is contact regular or clumped?

Human contact is rarely Poisson. Burstiness compares the variability of the
gaps between a vertex's events against the exponential reference: `1` is the
bursty limit, `0` is Poisson, `-1` is perfectly regular.

```{r burstiness}
head(burstiness(dn))
```

### `durations()` — how long ties actually last

By default `durations()` counts events per pair; `measure = "mean"` gives the
average length of a spell, and `unit` moves the question from pairs to spells
or to vertex activity.

```{r durations}
head(durations(dn, measure = "mean"))
```

### `mixing()` — who connects across which groups

Mixing needs vertex attributes, so this one uses the forum data, where
`forum_people` supplies a `role` for each participant. `forum_posts` is a
*threaded* log: a post keeps its tie alive until the thread falls silent.

```{r forum}
fn <- dynet(forum_posts, thread = "thread", nodes = forum_people)
fn
```

```{r mixing}
mixing(fn, attribute = "role", step = 60)
```

A `step` wider than the observation window collapses the counts into a single
bin; leave it out and you get the same table once per bin, which is what you
want when you are asking whether cross-role contact changes over the term.

## Where to go next

- `vignette("building-networks", package = "Dynet")` — the four log formats in
  detail, node attributes, observation windows, sessions, and the editing verbs
  (`add_ties()`, `set_observations()`, `induce_subgraph()` and friends).
- **`?paths`** and **`?plot_path_trajectories`** — the full traversal
  semantics: `traversal_time` for a cost charged per hop, session-bounded
  search, `as.data.frame(x, what = "steps")` for the reconstructed routes
  themselves, and the remaining trajectory-tree controls.
- **`?dyn_centrality`** and **`?metrics`** — the complete measure catalogues.
  Passing a measure that does not exist prints the list of ones that do.
- `projection()`, `pshifts()`, `snapshots()` and `collapse_network()` cover
  two-mode projection, participation shifts, per-bin edge lists and static
  aggregation.

`browseVignettes("Dynet")` lists everything installed alongside this one.
