---
title: "Getting started with Dynet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with Dynet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5)
library(Dynet)
```

Temporal network analysis studies relations together with the times at which
they occur. A face-to-face conversation has a beginning and an end, a message
has a timestamp, and a reply in a discussion thread follows the post it
answers. Such data come in two types: *contact sequences*, in which each
relation is an instantaneous event, and *interval graphs*, in which each
relation is active over a period of time. Dynet stores both as **spells**. A
spell is an ordered pair of vertices together with the interval over which the
tie between them is active, and a temporal network is the set of all spells
over an observation period.

Aggregating the spells into one static graph discards three properties of the
data: how often a pair interacted, how long each interaction lasted, and the
order in which the interactions occurred. Order has the largest consequences,
because it constrains what can flow through the network. A sequence of ties
from A through B to C can carry information from A to C only if the tie
between B and C is active at or after the time the tie between A and B is
used. Such a sequence is a **time-respecting path**. Every time-respecting
path is a path of the aggregated graph, but the converse does not hold, and
reachability along time-respecting paths is not transitive: A may reach B and
B may reach C while A cannot reach C. The timing of relationships can
therefore leave far fewer vertices reachable than the aggregated graph
suggests, and uneven timing of contacts slows spreading compared with the same
contacts shuffled in time.

Collaborative learning is a natural case for these methods. The order in which
students contribute is part of what is being studied, so a model of who
interacted with whom that ignores when they interacted discards the object of
study.

This vignette applies these ideas to one log of classroom contacts. The
analysis moves from the network as a whole, to individual vertices, to paths
between vertices, and finally to the spells themselves:

1. Construct the temporal network and describe its size and time span.
2. Measure density in successive windows, which gives a graph-level time
   series.
3. Measure the degree of every vertex in every window, which gives one
   trajectory per vertex.
4. Compute temporal closeness and temporal betweenness on time-respecting
   paths.
5. Trace the time-respecting paths from one source, forward and backward in
   time, and represent them as routes and as a tree.
6. Describe tie dynamics: formation and dissolution, duration and burstiness.

## Data

`school_contacts` is a simulated interval log of face-to-face contacts among
fourteen students over about three weeks. Each row is one contact. `from`
names the student who initiated the contact and `to` the student who was
addressed, so every contact has a direction. `start` and `end` give the times
at which the contact began and ended, in days since the start of observation.
Times are decimal, so a contact can begin and end within one day.

To inspect the log, we call `head()`, which prints its first six rows.

```{r data}
head(school_contacts)
```

## Building the network

`dynet()` constructs a temporal network from a relational log. It recognises
four log formats (interval, contact, threaded and co-presence) from the
columns and arguments it receives. A log with both a `start` and an `end`
column is read as an interval log, which gives an interval graph. Each spell
is active on the half-open interval $[\text{start}, \text{end})$: it includes
its start time and excludes its end time, so a spell that ends at time $t$ and
a spell that begins at $t$ are never active together.

To construct the network, we call `dynet()` on the log. The print method
reports the format, the direction, the size and the observed time span,
followed by the first spells. `dynet()` adds a `duration` column, equal to
`end` minus `start`, and a `weight` column.

```{r build}
dn <- dynet(school_contacts)
dn
```

The network has 14 vertices, 240 spells and 110 distinct ordered pairs, and
is observed from time 0 to 21.52. Times are stored as plain numbers with the
unit `step`; in this log one step is one day. The bin width of 1 sets the
default resolution of every windowed measure below to one day. Every spell
has weight 1.

Because the network is directed, the pair (Ana, Cara) is distinct from
(Cara, Ana), and a time-respecting path traverses a spell only from its
`from` vertex to its `to` vertex. With $n = 14$ vertices there are
$n(n - 1) = 182$ possible ordered pairs, so 110 of 182 pairs, about 60%, were
in contact at least once. The aggregated graph is therefore dense. The rest of
the vignette asks how much of that connectivity is present at any one time,
and how much is available along time-respecting paths.

To obtain the same description as a table with one row per property, we call
`summary()`. It adds the number of time bins and the mean snapshot density,
the average of the daily densities computed in the next section.

```{r summary}
summary(dn)
```

The observation period spans 22 daily bins, and the mean snapshot density is
0.0829. On an average day about one ordered pair in twelve is in contact,
against 60% over the whole period. The gap between the two figures is a first
sign that the aggregated graph overstates the connectivity present at any
given time.

To see when spells form and dissolve over the observation period, we call
`plot()` with `type = "activity"`. The activity plot places spell formation
and dissolution on a common time axis, which shows periods of high and low
turnover before any measure is computed.

```{r activity}
plot(dn, type = "activity")
```

## Measuring in windows

A temporal network can be measured in two ways. The first divides the
observation period into windows, aggregates the spells in each window into a
static snapshot, and computes a static measure on every snapshot, which gives
a time series. The second computes measures directly on time-respecting paths
and is introduced in the section on temporal centrality. This section and the
next use the first.

Every Dynet function that returns a time series takes the same four timing
arguments. `start` and `end` set the first and last measurement times and
default to the observed range. `step` is the interval between successive
measurements and defaults to the bin width of the network. `window` is the
length of time each measurement covers, counted forward from the measurement
time, so the measurement at time $t$ uses the spells active in
$[t, t + \text{window})$. It defaults to `step`, which divides the
observation period into non-overlapping bins. The window length is an
analytic choice: a short window resolves change finely but gives sparse
snapshots, and a long window gives denser snapshots at the cost of temporal
resolution.

**Density** is the proportion of possible ties that are present. In a
directed network with $n$ vertices, the density of window $w$ is

$$D_w = \frac{m_w}{n(n-1)},$$

where $m_w$ is the number of ordered pairs with at least one spell active at
some time in the window. A pair counts once however many spells it has in the
window, so $D_w$ lies between 0 and 1.

To obtain daily density, we call `metrics()` with `measure = "density"`. The
result has one row per window, with the columns `time` (the start of the
window), `measure` and `value`.

```{r density}
density <- metrics(dn, measure = "density")
density
```

On day 0 the density is 0.055: 10 of the 182 ordered pairs were in contact at
some time that day. On day 6 it rises to 0.159, or 29 pairs. The final bin
begins at day 21 and ends with the observation at day 21.52, so it covers
about half a day, and its density is not comparable with that of a full day.

To reduce the series to one row, we call `summary()`. The columns are the
number of windows `n`, the `mean`, standard deviation `sd`, `min` and `max` of
the values, and `peak_time`, the start of the window with the largest value.

```{r density-summary}
summary(density)
```

Across the 22 bins the mean daily density is 0.083 (SD 0.039). The maximum,
0.165, falls on day 14, when 30 of the 182 ordered pairs were in contact. The
minimum, 0.033, belongs to the truncated final bin; among full days the lowest
value is 0.038, on day 18.

Daily snapshots vary from one day to the next, partly because a contact near a
bin boundary falls on one side or the other of it. A rolling window reduces
this variation. Setting `window` larger than `step` makes successive windows
overlap: with `step = 1` and `window = 7`, a measurement is taken every day
and each covers the seven days $[t, t + 7)$. A pair is present if it was in
contact at any time in those seven days, so the weekly density starting at
time $t$ is never lower than the daily density starting at $t$. Windows that
start after day 14.52 extend past the end of observation and cover fewer than
seven days, so the final values of the series rest on progressively shorter
windows.

To obtain the rolling weekly series and draw it, we call `metrics()` with
`step = 1` and `window = 7`, and then `plot()`.

```{r density-rolling}
density_weekly <- metrics(dn, measure = "density", step = 1, window = 7)
plot(density_weekly)
```

## Vertex trajectories

Density describes the network as a whole. A **centrality** index describes
the position of one vertex. `dyn_centrality()` computes a centrality index
for every vertex in every window and returns one row per vertex per window,
so each vertex has a trajectory over time. `measure` selects the index, and
`start`, `end`, `step` and `window` work as in `metrics()`.

**Degree** counts the distinct partners of a vertex within a window. In a
directed network the default, `mode = "all"`, adds in-degree, the number of
distinct vertices with a spell to the vertex, and out-degree, the number of
distinct vertices the vertex has a spell to. Several spells between the same
ordered pair in one window count once.

To obtain daily degree, we call `dyn_centrality()` with
`measure = "degree"`. The result has the columns `time`, `node`, `measure`
and `value`.

```{r degree}
degree <- dyn_centrality(dn, measure = "degree")
degree
```

Each row is one vertex in one window, so 14 students over 22 days give 308
rows. In the twelve rows shown for day 0, five students have two partners and
the other seven one. The first day's network is a scatter of isolated
contacts, not a connected structure.

To reduce each trajectory to one row per vertex, we call `summary()`. It
reports the number of windows `n`, the `mean`, `sd`, `min` and `max` of the
vertex's daily degree, and `peak_time`, the start of the window in which that
degree was highest.

```{r degree-summary}
summary(degree)
```

Mean daily degree varies little across the class, from 1.64 to 2.86 partners
per day. Students differ more in how their contacts are distributed over days
than in how many contacts they have. The summary shows two contrasting
patterns of contact.

- **Steady contact.** The student has a partner on most days and never many
  at once. Two students have at least one partner on every day (`min` of 1),
  and the three smallest standard deviations, 1.07 to 1.26, belong to
  students whose maximum is 4 to 6.
- **Episodic contact.** Days without any partner alternate with days of many
  partners. Five students reach a maximum of 7 or 8, and their standard
  deviations lie between 1.78 and 2.17. Their peaks fall on days 6, 13 and 14,
  the three densest days of the observation period.

Mean degree does not separate the two patterns, because the means of both
fall within the same narrow range. The patterns are the two ends of a
continuum, not discrete groups. They matter for anything that spreads through
contact: a steadily connected student can pass something on at almost any
time, while an episodically connected student can do so mainly when the whole
class is active.

To display all fourteen trajectories together, we call `plot()` with
`type = "heatmap"`. It draws one tile per vertex and window, shaded by degree.

```{r degree-heatmap}
plot(degree, type = "heatmap")
```

## Temporal centrality

Windowed degree treats each day as a separate static network. A contact on
one day and a contact on the next are never combined, so a windowed measure
cannot show whether something could pass from one student to another through
a third over several days. Temporal centrality indices address this by
computing distances along time-respecting paths over the whole observation
period. Setting `scope = "temporal"` in `dyn_centrality()` selects this
approach and returns one value per vertex.

A time-respecting path can be optimal in three different senses. A
**foremost** path reaches its endpoint at the earliest possible time, a
**shortest** path uses the fewest hops, and a **fastest** path has the
smallest difference between arrival and departure. Dynet uses **shortest
foremost** paths: among the paths that arrive earliest, it keeps those with
the fewest hops. A path may wait at a vertex for any length of time, and its
hop times must not decrease, so two spells active at the same instant can be
chained.

The **latency** from a source $s$ to a vertex $z$ is $a_z - o_s$, where $a_z$
is the earliest arrival time at $z$ and $o_s$ is the time at which the search
leaves $s$, here the start of observation. Temporal closeness is built from
these latencies. With $R_s$ the set of vertices other than $s$ that are
reachable from $s$, Dynet computes

$$C(s) = \frac{|R_s|}{\sum_{z \in R_s} (a_z - o_s)},$$

the reciprocal of the mean latency to the reachable vertices, in inverse days.
A student whose contacts connect early to the rest of the class has high
temporal closeness, whatever that student's degree on any single day.

To obtain temporal closeness, we call `dyn_centrality()` with
`measure = "closeness"` and `scope = "temporal"`.

```{r closeness}
closeness <- dyn_centrality(dn, measure = "closeness", scope = "temporal")
closeness
```

Temporal closeness ranges from 0.131 to 0.375, a mean latency of 7.6 days at
the low end and 2.7 days at the high end. The time a student needs to reach
the rest of the class therefore varies almost threefold, while mean daily
degree varies less than twofold. The ordering does not follow contact volume.
The two highest values, 0.375 and 0.367, belong to the students with the
lowest (1.64) and the highest (2.86) mean daily degree. Closeness is set by
timing. A student reaches the class quickly when that student's early
contacts are with partners who are about to meet others, however many
contacts the student has.

**Temporal betweenness** measures how often a vertex lies on the optimal
paths between other vertices. For each ordered pair of a source $s$ and a
target $t$ reachable from $s$, one unit of credit is divided equally among the
shortest foremost paths from $s$ to $t$. Each intermediate vertex receives the
share of those paths that pass through it, and the source and target receive
nothing. Summing over all ordered pairs gives

$$B(v) = \sum_{s \neq v \neq t} \frac{\sigma_{st}(v)}{\sigma_{st}},$$

where $\sigma_{st}$ is the number of shortest foremost paths from $s$ to $t$
and $\sigma_{st}(v)$ is the number of those paths that pass through $v$. Dynet
does not normalise the sum; with $n = 14$ vertices it lies between 0 and
$(n - 1)(n - 2) = 156$.

To obtain temporal betweenness, we call `dyn_centrality()` with
`measure = "betweenness"` and the same scope.

```{r betweenness}
betweenness <- dyn_centrality(dn, measure = "betweenness", scope = "temporal")
betweenness
```

Temporal betweenness ranges from 4.3 to 41.9, almost tenfold, so it
separates the students far more than degree does. It measures a different
role from closeness:

- **Reach** (closeness): how soon a student's own contacts connect that
  student to everyone else.
- **Relay** (betweenness): how often a student's contacts fall after a path
  from one student has arrived and before it continues to another.

In these data the two roles are held mostly by different students. The two
highest betweenness values, 41.9 and 41.8, belong to students ranked seventh
and eighth of fourteen in closeness. The student with the second-highest
closeness, who also has the highest mean daily degree, has the second-lowest
betweenness, 7.0. Only one student holds both roles, with the highest
closeness and the third-highest betweenness (32.6). Neither a large number of
contacts nor fast reach makes a student a relay. A student relays only when
the timing of that student's contacts bridges the timing of other students'
contacts.

## Time-respecting paths

Temporal centralities summarise every source at once. `paths()` examines one
source: it returns the shortest foremost paths from the vertex named in
`from` to every other vertex. The search begins at the start of observation
unless `start` is set. The result has one row per vertex. `reachable` records
whether any time-respecting path from the source reaches the vertex, and
`arrival_time` the earliest time a path reaches it. `attained` records
whether a path arrives at exactly that time, which matters for backward
searches. `latency` is `arrival_time` minus the search start, `n_hops` the
number of spells on a shortest foremost path, and `n_paths` the number of
distinct shortest foremost paths.

To obtain the paths leaving Ana, we call `paths()` with `from = "Ana"`.

```{r paths}
from_ana <- paths(dn, from = "Ana")
from_ana
```

The search starts at day 0. The source reaches three students in one hop,
four in two, four in three and two in four. Hop count and arrival time are
only loosely related. A two-hop path arrives on day 6.12, before two of the
one-hop paths (days 6.36 and 6.67), and a four-hop path arrives on day 7.98,
two days before a three-hop path (day 10.00). In a static graph the vertex
fewer hops away is the nearer one. Along time-respecting paths the nearer
vertex is the one whose connecting contacts happen sooner. Three destinations
are reached by more than one shortest foremost path, which the tree below
shows as parallel branches. The source's temporal closeness of 0.131 can be
recovered from this table: 13 students reached, divided by the sum of their
latencies, 98.96 days.

To summarise the reachable set, we call `summary()`. It reports the number
and share of vertices reached, and the median and maximum of latency and of
hops.

```{r paths-summary}
summary(from_ana)
```

Ana reaches all 13 other students (reachable share 1). The median latency is
7.51 days and the maximum 11.66 days; the median path uses 2 hops and the
longest 4. On no single day are more than 16.5% of the ordered pairs in
contact (maximum daily density 0.165), yet contacts chained across days
connect Ana to every other student within twelve days.

Reachability depends on when the search begins. A spell can extend a path
only if it is active at or after the time the path reaches its `from` vertex,
so spells that ended before the search start cannot be used. To begin the
search on day 18, we call `paths()` with `start = 18`; only spells active
during the last 3.52 days of observation are then available.

```{r paths-late}
from_ana_late <- paths(dn, from = "Ana", start = 18)
summary(from_ana_late)
```

From day 18, Ana reaches 5 of the 13 other students (reachable share 0.385),
with a maximum latency of 2.68 days and at most 3 hops. From day 0 the same
source reaches every student. In a temporal network reachability is a
property of a vertex at a given time, not of the vertex alone.

`paths()` can also search backward in time. With `direction = "backward"`,
the search starts at the end of observation, day 21.52, and finds for every
other vertex the latest time at which it could begin a time-respecting path
that still reaches Ana by the end. `arrival_time` then holds this latest
departure time, and `latency` is the end time minus it. The vertices reached
this way form the **source set** of Ana, the vertices from which information
could have reached Ana. A spell excludes its end time, so when the latest
departure time equals the end of a spell, no path can depart at exactly that
instant; the vertex is still reachable, and `attained` is `FALSE`.

To obtain the backward search into Ana, we call `paths()` with
`direction = "backward"`.

```{r paths-backward}
into_ana <- paths(dn, from = "Ana", direction = "backward")
summary(into_ana)
```

All 13 other students could have reached Ana by the end of observation. The
median backward latency is 4.27 days and the maximum 8.9 days.

## Routes

`paths()` returns one row per destination. `pathways()` returns one row per
distinct **route**, the sequence of vertices that a shortest foremost path
visits. Routes are identified by their vertex sequence alone, so paths that
visit the same students through different spells are combined into one row
and their counts are added. The columns are `route`, `endpoint`, `count` (the
number of shortest foremost paths that follow the route), `share` (the
route's fraction of all counted paths), `n_hops` and `arrival_time`, with the
most frequent route first.

To obtain the routes leaving Ana, we call `pathways()` with `from = "Ana"`.

```{r pathways}
pathways(dn, from = "Ana")
```

Route use is concentrated. One route carries 3 of the 7 counted paths
(share 0.429) and ends at the last student reached, on day 11.66. The other
four routes carry one path each. The dominant route is a single sequence of
students used at three different times. It is a channel the network
reproduces, not a single chance chain of contacts. The seven counted paths
are the ones that end at a leaf of the tree drawn below. A path that stops at
an intermediate vertex is the beginning of a longer route and is not listed
on its own.

To draw the shortest foremost paths as a tree, we call `plot()` on the
`paths()` result. The source is the root, and each level to the right adds one
hop. Node size and branch width show how many shortest foremost paths use the
branch, and each node is labelled with its vertex and path count. A tree node
is a vertex together with the time at which the path reached it, so a vertex
appears more than once when paths reach it at different times. The arrival
time matters because it determines which later spells the path can still use.

```{r tree, fig.height = 5}
plot(from_ana)
```

To obtain the tree as a table, we call `path_trajectories()` on the `paths()`
result. Each row is one node of the tree. `node` identifies the node by its
route, written as `vertex@time` steps joined by arrows. `parent` is the node
one hop shorter, `depth` the number of hops from the source, and `count` the
number of shortest foremost paths that pass through or end at the node.
`probability` is the node's `count` divided by its parent's `count`, the
proportion of the parent's paths that continue along this branch. `vertex`
and `time` give the vertex and its arrival time separately.

```{r trajectories}
tree <- path_trajectories(from_ana)
tree
```

The tree has 22 nodes. The root, `Ana@0`, has a count of 19, one for each
shortest foremost path in the `paths()` result, including the zero-hop path
at the source. The first-hop branches take two different forms.

- **Broad branch.** One contact of the source fans out. A single first-hop
  branch carries 7 of the 19 paths (probability 0.368) and spreads them over
  six further students along two sub-branches.
- **Repeated branch.** One chain is used again and again. The same first-hop
  student appears on three branches, reached on days 2.12, 3.43 and 6.68.
  Together these branches carry 9 paths, and every one continues through the
  same three students. This is why two destinations have three shortest
  foremost paths each.

A broad branch spreads the source's reach across the class. A repeated branch
gives several chances to reach the same few students, so reach along it does
not depend on any single contact.

The third repeated branch also shows a property with no static counterpart.
It reaches its second-hop vertex on day 6.68, later than that vertex's
earliest arrival on day 6.12, yet still reaches the next vertex on day 9.59,
that vertex's earliest arrival. In a static graph every part of a shortest
path is itself a shortest path. A shortest foremost path does not have this
property: it need not reach each intermediate vertex at the earliest possible
time.

## Tie dynamics

The analyses so far concern who can reach whom. A temporal network is also
described by the dynamics of its ties: when they begin and end, how long they
last, and how their occurrences are spaced in time.

`events()` counts, for each window, the spells that begin in it
(**formation**) and the spells that end in it (**dissolution**). With the
default `window` equal to `step`, the windows do not overlap, so each spell's
formation and dissolution are each counted once. To obtain both series and
reduce each to one row, we call `events()` and then `summary()`.

```{r events}
turnover <- events(dn)
summary(turnover)
```

Both series average 10.91 spells per day, which is 240 spells divided by 22
bins: every spell begins and ends inside the observation period. Formation
peaks at 29 spells on day 13 and dissolution at 27 on day 14, the day of
highest density. At least one day has no new spell (minimum formation 0),
while at least 3 spells end on every day (minimum dissolution 3).

Duration is what separates an interval graph from a contact sequence. Two
pairs with the same number of spells can differ in how long their contacts
last, and an aggregated graph weighted by spell counts gives them the same
weight. `durations()` summarises the spell durations of each ordered pair,
where the duration of a spell is `end` minus `start`. `measure = "mean"`
gives the mean duration of the pair's spells; `"events"` gives their number,
`"total"` their summed duration and `"median"` their median. The result has
one row per ordered pair that was ever in contact, with the columns `from`,
`to`, `measure` and `value`.

To obtain the mean spell duration of every pair, we call `durations()` with
`measure = "mean"`.

```{r durations}
lengths <- durations(dn, measure = "mean")
lengths
```

The first rows are the pairs with Ana as the `from` student. They show that
durations vary widely even within one student's ties. Mean spell length runs
from 0.10 to 1.19 days, a twelvefold difference that a graph weighted by
spell counts cannot show.

The same rows show a second distinction, between an **aggregated tie** and a
**usable tie**. The source has direct spells to eight students, but only
three of those ties are used by the shortest foremost paths in the `paths()`
table. The other five students are reached by relayed paths of two to four
hops, because every direct spell to them begins after the relayed path has
already arrived. A tie in the aggregated graph says nothing about when that
tie can be used.

**Burstiness** describes how the events of a vertex are distributed in time.
Here an event of a vertex is the start of a spell that involves it, as the
`from` or the `to` student. With $\tau_1, \ldots, \tau_k$ the gaps between
consecutive events, $\mu$ their mean and $\sigma$ their population standard
deviation,

$$B = \frac{\sigma - \mu}{\sigma + \mu}.$$

$B$ equals 0 when $\sigma = \mu$, which holds for the exponentially
distributed gaps of a Poisson process. It approaches 1 when a few long gaps
separate clusters of short gaps, and it equals −1 when all gaps are equal. The
**memory** coefficient $M$ is the Pearson correlation between consecutive
gaps, $\tau_i$ and $\tau_{i+1}$. A positive $M$ means that short gaps tend to
follow short gaps and long gaps tend to follow long gaps. Burstiness matters
for processes on the network: bursty contact timing slows spreading compared
with the same contacts shuffled in time.

To obtain $B$, $M$ and the event count of every vertex, we call
`burstiness()` and then `summary()`, which lists one row per vertex and
measure. Each row summarises a single value, so `mean`, `min` and `max`
coincide and `sd` is `NA`.

```{r burstiness}
rhythm <- burstiness(dn)
summary(rhythm)
```

Burstiness lies between −0.070 and 0.246, so every student is within 0.25 of
the Poisson reference of 0. Eight students have slightly positive values,
meaning mildly clustered contact, and six have slightly negative values,
meaning mildly regular contact. No student departs far from random timing, so
burstiness does not divide the class into bursty and regular students. Memory
varies more, from −0.227 to 0.480, and is positive for nine of the fourteen
students: for most students a short gap between contacts tends to be followed
by another short gap. Event counts range from 28 to 46.

Two kinds of measure have described the students in this vignette.
**Volume** measures, the event count and mean daily degree, record how much a
student interacts. **Timing** measures, temporal closeness and temporal
betweenness, record where those interactions fall relative to everyone
else's. In these data the two kinds disagree. The student with the fewest
events (28) and the lowest mean degree has the highest closeness and the
third-highest betweenness. The student with the most events (46) and the
highest mean degree has the second-lowest betweenness. Volume also varies
within a narrow band (mean degree 1.64 to 2.86), while betweenness varies
almost tenfold (4.3 to 41.9). A student's temporal position is set by when
that student's contacts occur, not by how many there are. This is the premise
of temporal network analysis.

## Next steps

`vignette("building-networks")` covers the other three log formats (contact,
threaded and co-presence), direction, loops, weights and vertex attributes,
sessions, observation windows and vertex activity spells.
`vignette("ch17-temporal-networks")` reproduces the analysis of a MOOC
discussion forum with Dynet. `?metrics` and `?dyn_centrality` list every
available measure. `?paths` documents the traversal rules, including
`traversal_time` for a fixed duration per hop and
`as.data.frame(x, what = "steps")` for the step-by-step reconstruction of
each path.
