---
title: "Trees of Thought: temporal-network reproduction with Dynet"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  data_dir: "/Users/mohammedsaqr/Library/CloudStorage/GoogleDrive-saqr@saqr.me/Other computers/My MacBook Pro (2)/My_Data/Zips/Trees_of_thought 2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE, echo = TRUE,
  fig.width = 10, fig.height = 6, dpi = 120
)
input_file <- knitr::current_input(dir = TRUE)
repo_dir <- normalizePath(file.path(dirname(input_file), "../../.."),
                          mustWork = TRUE)
devtools::load_all(repo_dir, quiet = TRUE)
data_dir <- normalizePath(params$data_dir, mustWork = TRUE)
```

# Rule 0

This is a Dynet reproduction, not a second implementation hidden in a
notebook. Every temporal-network result below comes directly from an exported
Dynet function. The notebook imports the saved network, calls those functions,
and plots their returned objects. It does not reconstruct statistics, reshape
them into substitute statistics, perform package-parity tests, or silently
alter the supplied network.

# Import the saved temporal network

```{r import}
dn <- as_dynet(readRDS(file.path(
  data_dir, "Out", "Dynamic_network_Weightedloops.RDS"
)))
dn
summary(dn)
```

The saved network contains ten codes, weighted directed edge spells, and nine
self-loop spells. They remain in the imported object. Each Dynet statistic
applies its documented loop convention.

# See the temporal network

## Edge-spell timeline

```{r edge-spell-timeline, fig.height=8}
plot(dn, type = "timeline")
```

## Formation, dissolution, and active ties

```{r edge-activity}
plot(dn, type = "activity")
```

## Whole-window network through cograph

```{r whole-window-network, fig.height=8}
cograph::splot(
  dn,
  theme = "dynet",
  tna_styling = FALSE,
  psych_styling = FALSE,
  layout = "oval",
  node_size = 4.5,
  label_size = 0.55,
  arrow_size = 0.35,
  edge_width_range = c(0.15, 2),
  edge_alpha = 0.30,
  edge_labels = FALSE,
  edge_label_style = "none"
)
```

## Duration-weighted collapsed network

```{r collapsed-network, fig.height=8}
collapsed <- collapse_network(dn, weight = "union_duration")
collapsed
cograph::splot(
  collapsed,
  theme = "dynet",
  tna_styling = FALSE,
  psych_styling = FALSE,
  layout = "oval",
  node_size = 4.5,
  label_size = 0.55,
  arrow_size = 0.35,
  edge_width_range = c(0.15, 2.5),
  edge_alpha = 0.40,
  edge_labels = FALSE,
  edge_label_style = "none"
)
```

## Temporal snapshots

```{r snapshots, fig.height=9, fig.width=10}
plot(dn, type = "snapshots", panels = 9)
```

## Proximity timeline with phase networks

```{r proximity-networks, fig.height=8, fig.width=12}
plot(dn, type = "proximity", measure = "degree",
     phases = 5, slices = 80)
```

## Proximity trajectories alone

```{r proximity-lines, fig.height=7, fig.width=12}
plot(dn, type = "proximity", measure = "betweenness",
     networks = FALSE, slices = 80)
```

# Graph-level trajectories

## Main structural descriptives

```{r graph-main}
graph_main <- dyn_metrics(
  dn,
  measure = c(
    "density", "efficiency", "connectedness", "reciprocity",
    "edges", "mean_degree"
  ),
  step = 0.1, window = 0.1
)
graph_main
plot(graph_main, type = "ridge")
```

## Dyad census

```{r dyad-census}
dyads <- dyn_metrics(
  dn, measure = c("mutual", "asymmetric", "null"),
  step = 0.1, window = 0.1
)
dyads
plot(dyads)
```

## Triad census

```{r triad-census, fig.height=8}
triads <- dyn_metrics(dn, measure = "triads", step = 0.1, window = 0.1)
triads
plot(triads, type = "heatmap")
```

## Lightweight ERGM-style descriptives

```{r ergm-descriptives}
ergm_descriptives <- dyn_metrics(
  dn,
  measure = c(
    "indegree_1_5", "outdegree_1_5", "triangles",
    "in_2stars", "out_2stars", "two_paths"
  ),
  step = 0.1, window = 0.1
)
ergm_descriptives
plot(ergm_descriptives, type = "ridge")
```

# Node-level trajectories

## Centrality

```{r centrality-heatmap, fig.height=11}
centrality <- dyn_centrality(
  dn,
  measure = c(
    "degree", "closeness", "betweenness", "eigenvector",
    "flow_betweenness", "diffusion"
  ),
  step = 0.1, window = 0.1
)
centrality
plot(centrality, type = "heatmap")
```

```{r centrality-trajectories, fig.height=10}
plot(centrality, type = "ridge", top = 10)
```

## In-degree and out-degree

```{r directed-degree, fig.height=7}
directed_degree <- dyn_centrality(
  dn, measure = "degree", mode = "in",
  step = 0.1, window = 0.1
)
out_degree <- dyn_centrality(
  dn, measure = "degree", mode = "out",
  step = 0.1, window = 0.1
)
plot(directed_degree, type = "heatmap")
plot(out_degree, type = "heatmap")
```

## Prestige

```{r prestige, fig.height=7}
prestige_indegree <- dyn_centrality(
  dn, measure = "prestige", prestige = "indegree",
  step = 0.1, window = 0.1
)
prestige_eigenvector <- dyn_centrality(
  dn, measure = "prestige", prestige = "eigenvector",
  step = 0.1, window = 0.1
)
prestige_domain <- dyn_centrality(
  dn, measure = "prestige", prestige = "domain",
  step = 0.1, window = 0.1
)
prestige_proximity <- dyn_centrality(
  dn, measure = "prestige", prestige = "domain.proximity",
  step = 0.1, window = 0.1
)
plot(prestige_indegree, type = "heatmap")
plot(prestige_eigenvector, type = "heatmap")
plot(prestige_domain, type = "heatmap")
plot(prestige_proximity, type = "heatmap")
```

# Code-to-code mixing

```{r mixing-heatmap, fig.height=14}
mixing <- dyn_mixing(dn, attribute = "name", step = 0.1, window = 0.1)
mixing
plot(mixing, type = "heatmap")
```

The original scripts singled out mixing involving group regulation. The same
selection is passed to Dynet's plot method as a highlight; no mixing statistic
is rebuilt in the notebook.

```{r mixing-group-regulation, fig.height=7}
group_regulation_flows <- grep(
  "Group_regulation", unique(mixing$measure), value = TRUE, fixed = TRUE
)
plot(mixing, highlight = group_regulation_flows)
```

# Tie and vertex duration

## Incident tie counts and durations by node

```{r node-tie-duration, fig.height=10}
node_ties_all <- dyn_durations(
  dn, unit = "node_ties", mode = "all",
  measure = c("events", "total", "union")
)
node_ties_in <- dyn_durations(
  dn, unit = "node_ties", mode = "in",
  measure = c("events", "total", "union")
)
node_ties_out <- dyn_durations(
  dn, unit = "node_ties", mode = "out",
  measure = c("events", "total", "union")
)
plot(node_ties_all)
plot(node_ties_in)
plot(node_ties_out)
```

## Edge-pair duration

```{r pair-duration, fig.height=10}
pair_duration <- dyn_durations(
  dn, unit = "pair",
  measure = c("events", "total", "union", "mean")
)
pair_duration
plot(pair_duration)
```

## Vertex activity duration

```{r vertex-duration, fig.height=8}
vertex_duration <- dyn_durations(dn, unit = "vertex_activity")
vertex_duration
plot(vertex_duration)
```

# Temporal reachability and centrality

```{r reachability, fig.height=8}
reachability <- dyn_reachability(
  dn, direction = "both", measure = c("reach", "reach_count")
)
reachability
plot(reachability)
```

```{r temporal-centrality, fig.height=9}
temporal_centrality <- dyn_centrality(
  dn,
  measure = c("closeness", "betweenness", "reach", "reach_count"),
  scope = "temporal"
)
temporal_centrality
plot(temporal_centrality)
```

# Forward temporal paths from every code

Each source gets a left-to-right trajectory tree. The tree repeats a node when it
occurs under a different temporal history; branch width and fill show the
number of optimal routes. The original scripts charged
`graph.step.time = 0.1`; Dynet receives the same value as
`traversal_time = 0.1`.

```{r forward-path-atlas, results='asis', fig.height=8, fig.width=12}
for (source_node in as.data.frame(dn, what = "nodes")$name) {
  cat("\n\n## ", source_node, "\n\n", sep = "")
  forward_paths <- dyn_paths(
    dn, from = source_node, direction = "forward", traversal_time = 0.1
  )
  print(plot_path_trajectories(
    forward_paths, measure = "frequency", orientation = "horizontal"
  ))
}
```

# Backward temporal paths to every code

The paper's latest-departure views used a deadline within the five-day display
window. Here the same question is asked with `end = 4`, where the supplied
interval network has attained backward routes.

```{r backward-path-atlas, results='asis', fig.height=8, fig.width=12}
for (target_node in as.data.frame(dn, what = "nodes")$name) {
  cat("\n\n## ", target_node, "\n\n", sep = "")
  backward_paths <- dyn_paths(
    dn, from = target_node, direction = "backward", start = 0, end = 4,
    traversal_time = 0.1
  )
  print(plot_path_trajectories(
    backward_paths, measure = "frequency", orientation = "horizontal"
  ))
}
```

# Course-group temporal subnetwork

```{r group-subnetwork-import}
full_dn <- as_dynet(readRDS(file.path(data_dir, "Out", "DyanmicFULL.RDS")))
full_ties <- as.data.frame(full_dn)
group_dn <- induce_subgraph(
  full_dn, ties = full_ties$course_group == "295_437"
)
group_dn
```

```{r group-subnetwork-splot, fig.height=8}
cograph::splot(
  group_dn,
  theme = "dynet",
  tna_styling = FALSE,
  psych_styling = FALSE,
  layout = "oval",
  node_size = 4.5,
  label_size = 0.55,
  arrow_size = 0.35,
  edge_width_range = c(0.15, 2.5),
  edge_alpha = 0.40,
  edge_labels = FALSE,
  edge_label_style = "none"
)
```

```{r group-subnetwork-timeline, fig.height=8}
plot(group_dn, type = "timeline")
```

```{r group-subnetwork-snapshots, fig.height=9}
plot(group_dn, type = "snapshots", panels = 9)
```

```{r group-subnetwork-metrics}
plot(dyn_metrics(
  group_dn,
  measure = c("density", "edges", "reciprocity", "connectedness"),
  step = 0.1, window = 0.1
), type = "ridge")
```

# Session information

```{r session-info}
sessionInfo()
```
