---
title: "The MOOC forum from the published files"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
```

In this article we build the MOOC forum network of
`vignette("ch17-temporal-networks")` from the two CSV files that chapter 17 of
*Learning Analytics Methods and Tutorials* (Saqr, 2024) publishes, rather than
from the bundled `mooc_posts` and `mooc_people`. The files are read and
prepared with `rio` and `dplyr`, as a reader following the chapter would, and
the network they give is identical to the vignette's, so the analysis is not
repeated here.

## Data

The chapter reads an edge list with one row per post and a node list with one
row per participant. To read them, we call `rio::import()` on each URL;
`janitor::clean_names()` puts the edge list's column names in snake case.

```{r import}
library(Dynet)
library(dplyr)
library(rio)
net_edges <- import("https://raw.githubusercontent.com/lamethods/data/main/6_snaMOOC/DLT1%20Edgelist.csv") |>
  janitor::clean_names()
net_nodes <- import("https://raw.githubusercontent.com/lamethods/data/main/6_snaMOOC/DLT1%20Nodes.csv")
head(net_edges)
```

`dynet()` matches the node table to the log by a column named `name`, so the
participant identifier is renamed. The experience code is recoded into the
three labels the chapter mixes on, and the columns the analysis does not read
are dropped.

```{r nodes}
net_nodes <- net_nodes |>
  rename(name = UID) |>
  mutate(expert_level = as.character(factor(experience, levels = 1:3,
                                            labels = c("Expert", "Student",
                                                       "Teacher")))) |>
  select(name, experience, expert_level)
head(net_nodes)
```

## The network

Following Saqr and Nouri (2020), a reply is active from the moment it is
posted until its thread falls silent. To build the network, we call `dynet()`
with `from` and `to` for the participant columns, `time` for the timestamp,
which is parsed from the chapter's month/day/year strings, `thread` for the
discussion title, `nodes` for the participant table, `time_unit = "days"`, and
`min_thread_posts = 2` to drop threads that never became an exchange.
Self-replies are dropped by default.

```{r build}
dn_full <- dynet(net_edges, from = "sender", to = "receiver",
                 time = "timestamp", thread = "discussion_title",
                 nodes = net_nodes, time_unit = "days",
                 min_thread_posts = 2)
summary(dn_full)
```

The network has 441 vertices and 2,406 spells on 1,907 pairs over 73 one-day
bins, the same as the vignette's. To restrict it to the participants with
more than 20 ties, we call `induce_subgraph()` with a condition on the vertex
table.

```{r active}
dn <- induce_subgraph(dn_full, degree > 20)
dn
```

The active subnetwork has 45 vertices, 686 spells and 428 pairs, observed
from day 0.11 to day 72.01. Every measure of `vignette("ch17-temporal-networks")`
applies to this object unchanged.

## References

Saqr, M. (2024). Temporal network analysis: Introduction, methods and
analysis with R. In M. Saqr & S. López-Pernas (Eds.), *Learning analytics
methods and tutorials: A practical guide using R*. Springer.
https://doi.org/10.1007/978-3-031-54464-4_17

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319).
ACM.
