---
title: "Function catalogue"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
library(Dynet)
```

Dynet exports 38 functions in six categories, and registers 50 S3 methods
across its result classes. Every function addresses vertices by name and
returns a data frame with one row per observation. This page lists each
function with what it takes and what it returns; the help page of each
function gives the full contract.

| Category | Functions | Purpose |
|---|---:|---|
| Construction | 4 | Build a network from a relational log and read it back per time bin |
| Editing | 17 | Rewrite spells and vertex activity without breaking time |
| Measurement | 8 | Centrality, reachability, structure, mixing, turn-taking, burstiness, duration, similarity |
| Paths | 5 | Time-respecting paths and four views of the result |
| Structure | 3 | Turn the network into another object |
| Animation | 1 | Render the measurement grid as an animation |

## Construction

`dynet()` builds a temporal network from a relational log. It takes a data
frame of ties and a set of column selectors, infers one of four log formats
from which selectors are named, and returns an object of class
`c("dynet", "netobject", "cograph_network")` carrying the spell table, the
vertex table and the construction metadata. The four formats are interval logs
with a start and an end per tie, contact logs of instantaneous events, threaded
logs in which a tie stays active until its thread falls silent, and co-presence
logs in which actors sharing a group become connected. To build a network from
an interval log, we call `dynet()` with the log; the columns are recognised by
name.

```{r construction}
dn <- dynet(school_contacts)
dn
```

The network has 14 vertices and 240 spells over 110 distinct pairs, observed
from 0 to 21.52 time units.

`as_dynet()` converts an object of another class to a temporal network. It
takes that object and returns a `dynet`. The method for `networkDynamic`
imports edge spells, vertex spells, the observation period and per-edge
attributes. The method for `dynet` returns its input unchanged, so the call is
safe on an object that is already a temporal network.

`events()` counts tie formation and dissolution over time. It takes a network,
one or more measure names in `measure`, and the four grid arguments, and
returns a `dynet_metric` at graph level with one row per time point and
measure. Eight measures are available: `"formation"`, `"dissolution"`,
`"active"`, `"new_pairs"`, `"formation_fraction"`, `"dissolution_fraction"`,
`"formation_rate"` and `"dissolution_rate"`. The two rates divide the
transitions by the exact eligible pair-time.

`snapshots()` lists the ties active in each time bin. It takes a network and
the grid arguments, or a single time in `at`, and returns a `dynet_snapshot`
data frame with one row per active tie per bin, holding `time`, `from`, `to`,
`weight` and `n_spells`, preceded by `session` when the network has sessions.
A pair joined by more than one spell inside one bin is one tie, and `n_spells`
records how many spells produced it. To list the ties in four-unit bins, we
call `snapshots()` with `step` set to 4.

```{r snapshots}
bins <- snapshots(dn, step = 4)
bins
```

Six bins hold 216 tie rows. In the first bin Ana and Jonas are joined by two
spells, which `n_spells` records, and `weight` sums them.

## Editing

Seventeen functions rewrite the spell table, the vertex table or the vertex
activity table. Each takes a network in `dn`, returns a new network of class
`c("dynet", "netobject", "cograph_network")`, and leaves its input unchanged.
Every edit goes through a rebuild, so the canonical spell identifiers can
renumber after any of them.

### Adding

`add_nodes()` adds vertices. It takes in `data` a character vector of names or
a data frame with a `name` column and static attributes, and returns a network
in which the new vertices are implicit always-active isolates until ties or
vertex activity are supplied.

`add_ties()` adds tie spells. It takes in `data` a data frame of endpoints and
times, and returns a network with the temporal ties and every flattened cograph
field rebuilt together. Columns beyond the spell fields are carried as tie
attributes. `loops` decides whether a self-tie is accepted.

`add_arcs()` is the directed counterpart of `add_ties()` and takes the same
arguments.

`add_vertex_spells()` declares periods in which a vertex is present. It takes
in `data` a data frame with `node`, `start` and `end`, and returns a network in
which the existing activity and the supplied spells are canonicalised together,
so a spell that overlaps or abuts an existing one for the same vertex is merged
into it.

### Removing

`remove_nodes()` removes vertices. It takes their names in `nodes` and a
`cascade` flag, and returns a network without them and, under
`cascade = TRUE`, without their ties and activity spells. A selection that
leaves no vertex or no tie raises `dynet_empty_network`.

`remove_ties()` removes tie spells. It takes row positions or a condition over
the spell table in `ties`, or the endpoint and time selectors `from`, `to`,
`start`, `end` and `session`, and returns a network without the matched spells.
A request that matches nothing raises `dynet_tie_not_found`.

`remove_arcs()` is the directed counterpart of `remove_ties()` and takes the
same arguments.

`remove_vertex_spells()` drops declared activity components. It takes integer
positions or a logical mask in `spells` and returns a network with the rest
canonicalised again, so the remaining spell identifiers renumber. A vertex left
with no declaration becomes implicitly always active.

### Updating and renaming

`update_nodes()` adds or replaces static vertex attributes. It takes in `data`
a data frame with a `name` key and one or more attribute columns, and returns a
network with the same spells, activity and metadata, and the attributes written
onto the named vertices. Vertices not named keep their values.

`update_ties()` edits selected spells. It takes a selection in `ties` and a data
frame of replacement values in `data`, and returns a network holding the
unselected spells unchanged and the selected ones with the supplied values
substituted.

`update_vertex_spells()` edits selected activity components. It takes positions
in `spells` and a data frame of replacement fields in `data`, and returns a
network in which the updated components are canonicalised with the retained
ones. A column outside the vertex-spell schema raises `dynet_unknown_column`.

`rename_nodes()` renames vertices. It takes in `mapping` a named character
vector, a two-column data frame named `old` and `new`, or the name of a vertex
attribute whose values become the names, and returns a network with tie
endpoints, vertex attributes, vertex activity, cograph labels and groups renamed
together.

`rename_sessions()` renames session labels. It takes the same mapping forms in
`mapping` and returns a network with tie and vertex session labels renamed
together and the session scheme in the metadata updated. Labels absent from the
mapping are left alone.

### Declaring support

`set_observations()` replaces the observation window. It takes either a
two-column data frame of components in `data` or a `start` and `end` pair, and
returns a network whose raw spells are unchanged and whose measurement view
alone is replaced, so `as.data.frame(x)` still returns the originals.

`clear_observations()` restores implicit support. It takes a network and
returns one observed continuously from its earliest raw start to its latest
raw end, with every explicit observation field dropped from the metadata.

`set_tie_sessions()` assigns or removes session walls. It takes in `session` a
character vector of length one or of the raw tie count, and returns a network
with a `session` column on the spell table and the scheme recorded in the
metadata, or with both removed under `session = NULL`. A full-length vector is
matched positionally against the sorted spell table, so labels should be
derived from `as.data.frame(dn)`.

`set_vertex_spells()` replaces declared vertex activity. It takes in `data` a
vertex-spell data frame, or the string `"ties"` to declare each vertex present
from its first tie to its last, and returns a network whose declared activity
is exactly that. To declare presence from the ties, we call it with `"ties"`,
and to read the declared activity we call `as.data.frame()` with `what` set to
`"vertex_spells"`.

```{r editing}
present <- set_vertex_spells(dn, "ties")
present
activity <- as.data.frame(present, what = "vertex_spells")
head(activity)
```

The tie spells are unchanged, so the network prints as before. The activity
table has one row per vertex, running from the vertex's first contact to its
last.

## Measurement

Eight functions measure a network over time. Each returns a `dynet_metric`: a
data frame whose columns are ordered `session`, `time`, the unit of
observation, `measure` and `value`. The `session` column appears only under
`sessions = "separate"`, the one mode that keeps session labels apart.

The grid arguments are not shared by all eight. `dyn_centrality()`,
`metrics()`, `mixing()` and `similarity()` take all four; `dyn_reachability()`
and `pshifts()` take `start` and `end` only; `burstiness()` and `durations()`
take none and summarise the whole observed period.

`dyn_centrality()` computes vertex centrality. It takes one or more measure
names in `measure` and a `scope`, and returns one row per vertex, time point
and measure. Nineteen measures are available at `scope = "snapshot"`, which
treats each time bin as a static graph, and four at `scope = "temporal"`,
`"closeness"`, `"betweenness"`, `"reach"` and `"reach_count"`, which are
measured along time-respecting paths and report no `time` column because the
whole window yields one value per vertex. On a directed network `mode` selects
incoming, outgoing or all ties. To measure betweenness in each unit bin, we
call it with `measure` set to `"betweenness"`.

```{r centrality}
between <- dyn_centrality(dn, measure = "betweenness")
between
```

The result has 14 vertices at 22 time points, 308 rows in all. Among the rows
printed for the first bin, Iris and Kira are the only vertices with a nonzero
value.

`dyn_reachability()` computes how much of the network each vertex can reach.
It takes a `direction` and a `measure`, `"reach"` for the proportion or
`"reach_count"` for the count, and returns one row per vertex per measure at
node level. The measures are labelled `forward_reach` and `backward_reach`, or
`forward_reach_count` and `backward_reach_count`. A vertex is excluded from its
own reachable set.

`metrics()` computes graph-level structure. It takes one or more of forty
measure names in `measure` and returns one row per time point and measure. The
measure `"triads"` contributes sixteen rows per time point, one per triad
class. To measure density in each unit bin, we call it with `measure` set to
`"density"`.

```{r metrics}
density <- metrics(dn, measure = "density")
density
summary(density)
```

Density is measured at 22 time points. `summary()` reduces the series to one
row with its mean, spread, range and the time of its peak: the mean density is
0.083, and the peak of 0.165 falls in the bin that begins at 14.

`mixing()` computes how much tie activity runs within and between vertex
groups. It takes the name of a vertex attribute in `attribute` and returns one
row per time point and ordered group pair, with `from_group` and `to_group`
naming the pair and `value` counting the pairs of vertices with an active tie
from the first group to the second.

`pshifts()` classifies consecutive turns into the thirteen participation-shift
types of Gibson (2003). It takes a directed network and an `output` shape, and
returns a `dynet_pshifts` data frame with columns `shift`, `family` and
`count`. Under `output = "final"` it has thirteen rows, one per shift type,
including the types that never occurred. Under `output = "cumulative"` it has
one block of thirteen rows per classified turn, carrying the running count.

`burstiness()` measures how unevenly each vertex's events are spaced. It takes
a measure name in `measure`, `"burstiness"`, `"memory"` or `"events"`, and
returns one row per vertex and measure, with no time column. The burstiness
coefficient is 1 in the bursty limit, 0 at the Poisson reference and -1 for
perfectly regular activity (Goh and Barabási, 2008); memory is the lag-one
correlation between consecutive inter-event gaps.

`durations()` measures how long relationships lasted. It takes a `unit` that
selects the row identity and a `measure`, `"events"`, `"total"` or `"mean"`,
and returns an edge-level table under `unit = "pair"` or `"spell"`, and a
node-level table under `"vertex_activity"`, `"vertex_spell"` or `"node_ties"`.
`censored` decides whether spells whose boundaries were never observed are
retained.

`similarity()` compares the network at each pair of time points. It takes a
coefficient in `method`, `"jaccard"`, `"overlap"`, `"hamming"`, `"cosine"` or
`"pearson"`, and returns a `dynet_similarity` data frame with one row per
ordered pair of time bins and columns `time`, `other`, `measure` and `value`.
The diagonal is included and is one for every coefficient except `"hamming"`,
where identical layers differ in nothing and score zero.

### The shared grid

Four arguments decide when the network is measured, following
`tsna::tSnaStats()`:

| Argument | Meaning | tsna equivalent |
|---|---|---|
| `start`, `end` | First and last measurement | `start`, `end` |
| `step` | Interval between measurements | `time.interval` |
| `window` | Length of time each measurement covers | `aggregate.dur` |

`window` equal to `step` tiles the period, `window` larger than `step` slides,
`window = 0` samples an instant, and `window = "all"` treats the whole observed
period as one bin.

## Paths

`paths()` finds time-respecting journeys from one vertex. It takes a network,
a source vertex name in `from` and a `direction`, and returns a `dynet_paths`
data frame with one row per vertex: `node`, `reachable`, `arrival_time`,
`attained`, `latency`, `n_hops` and `n_paths`. A journey is time-respecting
when each hop leaves no earlier than the previous hop arrived (Kempe, Kleinberg
and Kumar, 2002). The criterion is earliest arrival first and fewest contacts
second. To find the journeys from one student, we call it with `from`.

```{r paths}
routes <- paths(dn, from = "Ana")
routes
```

Ana reaches all 13 other students. Jonas is reached at time 2.12 in one hop;
Eve is reached at 11.66 in four hops, by three equally early routes.

`pathways()` counts the routes those journeys take. It takes a network and,
optionally, one source in `from`, and returns a `dynet_pathways` data frame
with one row per distinct route, most frequent first: `route` as the vertex
sequence joined by arrows, `endpoint`, `count`, `share` as the route's fraction
of all counted routes, `n_hops` and `arrival_time`.

`path_network()` builds the union of the optimal routes. It takes a `paths()`
result and returns a static `dynet_path_network`, whose tie table has one row
per hop used by at least one optimal route and whose vertex table has one row
per reached vertex. Unreachable vertices are absent rather than present with
missing values.

`path_trajectories()` arranges the routes as a counted prefix tree. It takes a
`paths()` result and a `min_count` threshold, and returns a
`dynet_path_trajectories` data frame with one row per tree node: `node` as the
route prefix, `parent`, `depth`, `count`, `probability`, `vertex`, `time`,
`session` and `branch`.

`plot_path_trajectories()` draws that tree. It takes a `paths()` or
`path_trajectories()` result and a `measure` that decides what the fill
encodes, `"frequency"`, `"time"` or `"predictability"`, and returns a `ggplot`
object. Each vertex name is printed beside its value, so no distinction rests
on colour alone.

A vertex with declared spells begins its search at its own first presence
inside the window rather than at the window edge. A vertex never present in the
window reaches nothing.

## Structure and animation

`projection()` expands the network into vertex-time states. It takes the grid
arguments and an interlayer coupling weight `omega`, and returns a
`dynet_projection` holding two tables, obtained with
`as.data.frame(x, what = "vertices")` for the states and
`as.data.frame(x, what = "edges")` for the directed arcs between them.
Coupling is ordinal: each slice is joined to the next one only.

`collapse_network()` reduces temporal activity to one static weighted network.
It takes a `weight` rule that decides what a tie weight means, `"binary"`,
`"union_duration"`, `"total_duration"`, `"duration_fraction"`,
`"spell_count"`, `"weight_sum"`, `"weighted_duration"` or `"latest_weight"`,
and returns a `dynet_collapsed` cograph network with a tie table and a vertex
table. Under `sessions = "separate"` it returns a `dynet_collapsed_list` with
one network per session. To collapse the classroom network, we call it with the
network; the default rule is binary.

```{r collapse}
static <- collapse_network(dn)
static
```

The collapsed network has 14 vertices and 110 ties, one per distinct pair, and
its tie table carries every weight rule as a column alongside the first and
last contact of the pair.

`induce_subgraph()` extracts a subgraph. It takes a condition over the vertex
table in `nodes`, a condition over the spell table in `ties`, or both, and
returns a network carrying only the selected spells, the vertices they touch,
those vertices' activity spells and all static attributes. Centralities are
available as columns inside the `nodes` condition, computed over the whole
observed period.

`animate()` renders the measurement grid as an animation. It takes the same
four grid arguments as every measuring function, a `file` whose extension
selects the encoder, `layout`, `measure` for vertex size, `tween` and `fps`
for the motion, and `absent`, `isolates` and `ease` for how presence and
transitions are drawn. It returns a `dynet_animation` data frame with one row
per bin holding `bin`, `frame`, `time`, `window_start`, `window_end`, `nodes`,
`idle`, `ties`, `forming`, `dissolving` and `file`. The table is returned
invisibly, since writing the file is the purpose of the call. GIF output needs
`gifski`; mp4 and webm output need `av`.

## Result classes

Fifty S3 methods are registered across the result classes: `print`, `summary`,
`plot`, `as.data.frame`, `head` and `tail`, and the two `as_dynet` methods. A
secondary table always comes out through an argument of `as.data.frame()`,
never by reaching into the object with `$`.

| Class | Returned by |
|---|---|
| `dynet` | `dynet()`, `as_dynet()`, and every editing function |
| `dynet_metric` | `dyn_centrality()`, `dyn_reachability()`, `metrics()`, `mixing()`, `burstiness()`, `durations()`, `events()` |
| `dynet_paths` | `paths()` |
| `dynet_pathways` | `pathways()` |
| `dynet_path_network` | `path_network()` |
| `dynet_path_trajectories` | `path_trajectories()` |
| `dynet_snapshot` | `snapshots()` |
| `dynet_similarity` | `similarity()` |
| `dynet_pshifts` | `pshifts()` |
| `dynet_projection` | `projection()` |
| `dynet_collapsed` | `collapse_network()` |
| `dynet_collapsed_list` | `collapse_network()` under `sessions = "separate"` |
| `dynet_animation` | `animate()` |

Secondary tables are obtained as `as.data.frame(x, what = "steps")` for the
hops of a `paths()` result, `as.data.frame(x, what = "vertex_spells")` for the
declared activity of a network, `as.data.frame(x, session = "s1")` for one
session, and the like.

On a measuring function, `plot = TRUE` follows `hist()`: drawing is a side
effect, and the tidy result is still returned, invisibly when it has been
drawn.

## References

Gibson, D. R. (2003). Participation shifts and institutional change in
relational systems. *Social Forces*, 81(4), 1335–1380.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.
