---
title: "Function catalogue"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
library(Dynet)
```

Dynet provides functions for temporal network construction, editing, measurement, path analysis, aggregation, and animation. Vertices are identified by name, and analytical results are returned in tidy formats with methods for printing, summarising, and plotting. This catalogue describes the principal inputs and outputs of the 39 exported functions. Individual help pages document their arguments and conditions in detail.

| Category | Functions | Purpose |
|---|---:|---|
| Construction | 4 | Construct or convert networks and inspect spells and snapshots |
| Editing | 17 | Modify vertices, relational spells, sessions, and observation periods |
| Measurement | 9 | Measure centrality, reachability, structure, mixing, turn-taking, timing, duration, and similarity |
| Paths | 5 | Find time-respecting paths and summarise or visualise their routes |
| Structure | 3 | Project, aggregate, or subset a network |
| Animation | 1 | Animate successive temporal snapshots |

## Construction

`dynet()` constructs a temporal network from relational data. It returns an object of class `c("dynet", "netobject", "cograph_network")`, containing relational spells, vertices, and construction metadata.

Four input formats are supported. Interval data supply onset and termination, or onset and duration. Contact data supply instantaneous interaction timestamps. Threaded data derive termination from the last retained interaction in each thread. Co-presence data connect actors attending the same occasion.

With `format = "auto"`, specifying `actor` and `group` selects co-presence; otherwise, specifying `thread` selects threaded construction. A specified or recognised termination or duration column selects interval data when neither preceding condition applies. Otherwise, contact data are selected. `format` can also be specified explicitly.

Column matching is case-insensitive. Recognised endpoint aliases include `from`/`to`, `sender`/`receiver`, and `source`/`target`; interval boundaries include `start`/`end` and `onset`/`terminus`. Explicit column specification is needed only for unrecognised or ambiguous names. The constructor calculates `duration = end - start` and assigns `weight = 1` when no multiplicity variable is supplied or recognised.

```{r construction}
dn <- dynet(school_contacts)
dn
```

The example contains fourteen vertices, 240 relational spells, and 110 distinct ordered pairs, observed from time 0 to 21.52.

`as_dynet()` converts supported network objects to `dynet`. Its `networkDynamic` method imports edge spells, vertex activity spells, observation periods, and edge attributes. Applied to a `dynet` object, it returns the input unchanged.

`events()` measures spell formation, dissolution, and activity over time. It accepts `measure` and the measurement-grid arguments and returns a graph-level `dynet_metric`. Available measures are `"formation"`, `"dissolution"`, `"active"`, `"new_pairs"`, `"formation_fraction"`, `"dissolution_fraction"`, `"formation_rate"`, and `"dissolution_rate"`. Formation and dissolution rates divide transition counts by eligible pair-time.

`snapshots()` returns the connections active within each measurement bin as a `dynet_snapshot`. Its tidy columns include `time`, `from`, `to`, `weight`, and `n_spells`, with session identifiers when applicable. Multiple spells on the same pair are combined within a bin: `weight` sums their weights and `n_spells` counts them. `at` selects a single measurement time.

```{r snapshots}
bins <- snapshots(dn, step = 4)
bins
```

The six four-unit bins contain 216 pair-by-bin observations. In the first bin, two spells connect Ana and Jonas; `n_spells` records both and `weight` sums their contributions.

## Editing

Editing functions return a new temporal network and leave their input unchanged. They update relational spells, vertex attributes, and associated network representations together. Rebuilding may change spell identifiers, so subsequent selections should refer to the updated object.

### Adding

`add_nodes()` adds vertices from a character vector or a data frame containing `name` and optional attributes. New vertices are treated as eligible throughout observation unless activity spells are declared.

`add_ties()` adds relational spells from a data frame of endpoints and times. Additional columns are retained as spell attributes, and `loops` controls whether self-links are accepted. The endpoints must already exist in the network. `add_arcs()` is the directed counterpart with the same arguments.

`add_vertex_spells()` adds activity periods from a table containing `node`, `start`, and `end`. Overlapping or adjacent periods for the same vertex are merged with existing declarations.

### Removing

`remove_nodes()` removes vertices selected by name. With `cascade = TRUE`, it also removes their incident spells and activity declarations. A selection leaving no vertices or no ties raises `dynet_empty_network`.

`remove_ties()` selects spells through positions or a condition in `ties`, or through `from`, `to`, `start`, `end`, and `session`. A selection matching no spells raises `dynet_tie_not_found`. `remove_arcs()` is the directed counterpart.

`remove_vertex_spells()` removes declared activity components selected by integer positions or a logical mask in `spells`. Remaining components are merged where needed and assigned updated identifiers. A vertex with no remaining declaration becomes implicitly eligible throughout observation.

### Updating and renaming

`update_nodes()` adds or replaces vertex attributes from a table keyed by `name`. Vertices absent from the supplied table retain their existing attributes.

`update_ties()` replaces fields of spells selected through `ties`, using replacement values supplied in `data`. Unselected spells are retained.

`update_vertex_spells()` replaces fields of activity components selected through `spells`. The updated and retained periods are combined, merging overlaps and adjacent periods. Unrecognised fields raise `dynet_unknown_column`.

`rename_nodes()` updates vertex names throughout the network. `mapping` accepts a named character vector, a data frame with `old` and `new` columns, or a vertex attribute whose values provide the new names. Relational endpoints, vertex attributes, activity declarations, and associated labels are updated together.

`rename_sessions()` updates session identifiers using the supported mapping forms and updates the session metadata. Unmapped session labels are retained.

### Observation periods and activity

`set_observations()` replaces the observation calendar using a table of intervals in `data` or a `start` and `end` pair. The original relational spells remain available through `as.data.frame()`; measurements use their intersections with the declared periods.

`clear_observations()` removes the explicit calendar and restores continuous observation from the earliest raw onset to the latest raw termination.

`set_tie_sessions()` assigns session labels through `session`, using either a single label or a vector matching the number of raw spells. A full-length vector follows the sorted spell-table order and should therefore be derived from `as.data.frame(dn)`. Setting `session = NULL` removes the assignments. The function also supports assigning sessions from onset-time boundaries through `breaks` and `labels`.

`set_vertex_spells()` replaces declared vertex activity using a table in `data`. The special value `"ties"` derives activity from each vertex’s first spell onset to its last termination.

```{r editing}
present <- set_vertex_spells(dn, "ties")
present
activity <- as.data.frame(present, what = "vertex_spells")
head(activity)
```

The relational spells are unchanged. The extracted activity table records a derived participation period for each vertex; these boundaries need not represent independently observed arrival or departure times.

## Measurement

Graph-level measures describe network structure as a whole. Vertex-level measures describe individual positions, while pair-level measures describe relationships between endpoints. Dynet returns these quantities in tidy results identified by the relevant vertices or pairs, `measure`, and `value`, with `time` and `session` where applicable.

The functions differ in their temporal arguments. `centrality_series()`, `metrics()`, `mixing()`, and `similarity()` accept `start`, `end`, `step`, and `window`. `reachability()`, `path_centrality()`, and `pshifts()` accept `start` and `end`. `burstiness()` and `durations()` summarise the observation period without a measurement grid.

`centrality_series()` applies the selected centrality measures to successive snapshots. For measures supporting direction selection, `mode` selects incoming, outgoing, or all connections. `path_centrality()` calculates `"closeness"` and `"betweenness"` using time-respecting paths; its results describe the selected search period and have no `time` column.

```{r centrality}
between <- centrality_series(dn, measure = "betweenness")
between
```

The example calculates snapshot betweenness for fourteen vertices across 22 bins, giving 308 observations. Among the values printed for the first bin, only Iris and Kira have nonzero betweenness.

`reachability()` calculates the number or proportion of other vertices connected to each vertex through time-respecting paths. `direction` selects forward or backward search, and `measure` selects `"reach"` or `"reach_count"`. Results are labelled `forward_reach`, `backward_reach`, `forward_reach_count`, or `backward_reach_count`. A vertex is excluded from its own reachable set.

`metrics()` calculates graph-level measures selected through `measure`. Results are indexed by measurement time and measure. Requesting `"triads"` returns counts for the sixteen directed triad classes at each time.

```{r metrics}
density <- metrics(dn, measure = "density")
density
summary(density)
```

The example measures density in 22 bins. `summary()` reports its mean, standard deviation, range, and peak time. Mean density is 0.083, and the maximum of 0.165 occurs in the bin beginning at time 14.

`mixing()` counts connections within and between groups defined by a vertex attribute. `attribute` selects the grouping variable. The result identifies ordered group pairs through `from_group` and `to_group`, with `value` counting distinct connected vertex pairs in each window. Connections in the same window need not be simultaneous, and these counts are not probabilities.

`pshifts()` classifies consecutive directed interactions into the thirteen participation-shift types of Gibson (2003). The returned `dynet_pshifts` contains `shift`, `family`, and `count`. `output = "final"` reports all thirteen types, including zero counts. `output = "cumulative"` reports running counts after each classified turn.

`burstiness()` describes the spacing of spell onsets involving each vertex. `measure` selects `"burstiness"`, `"memory"`, or `"events"`. Burstiness is $(\sigma-\mu)/(\sigma+\mu)$, where $\mu$ and $\sigma$ are the mean and population standard deviation of inter-event intervals. It equals −1 for equal positive intervals and approaches 1 with increasing relative variability. A value of 0 matches the theoretical exponential waiting-time reference but does not establish Poisson timing. Memory is the correlation between successive intervals.

`durations()` summarises observed spell durations. `unit` selects relational pairs (`"pair"`), individual spells (`"spell"`), vertex activity (`"vertex_activity"` or `"vertex_spell"`), or incident relational spells (`"node_ties"`). Available measures depend on this unit. Pair-level defaults are `"events"`, `"total"`, and `"mean"`; `"union"` and `"median"` are also available. Individual-spell output defaults to `"duration"`. `censored` controls inclusion of spells with unobserved boundaries.

`similarity()` compares the connection sets of temporal snapshots using `"jaccard"`, `"overlap"`, `"hamming"`, `"cosine"`, or `"pearson"`. It returns a `dynet_similarity` indexed by `time` and `other`, with the selected coefficient in `measure` and its value in `value`. Self-comparisons are included. Hamming distance is zero for identical snapshots; the other coefficients express similarity. At least two measurement bins are required.

### The shared grid

For functions supporting window-based measurement, four arguments define the grid. Their relationship to `tsna::tSnaStats()` is shown below.

| Argument | Meaning | tsna equivalent |
|---|---|---|
| `start`, `end` | First and last measurement times | `start`, `end` |
| `step` | Interval between measurements | `time.interval` |
| `window` | Duration covered from each measurement time | `aggregate.dur` |

By default, `step` uses the network’s construction interval and `window` equals `step`, producing non-overlapping windows. A larger `window` produces overlapping windows. `window = 0` evaluates individual time points, while `window = "all"` aggregates the observation period into a single window. The last option is unsuitable for `similarity()`, which requires multiple snapshots.

## Paths

`paths()` finds time-respecting paths from the vertex named in `from`. Each interaction must be available at or after arrival at its starting vertex. The default search selects earliest arrival first and then the fewest interactions among paths arriving at that time: the shortest foremost criterion.

The returned `dynet_paths` describes each destination through `node`, `reachable`, `arrival_time`, `attained`, `latency`, `n_hops`, and `n_paths`. In a forward search, latency is elapsed time from the search start to arrival, including waiting. A backward search instead identifies the latest departure boundary from which the specified vertex can be reached by the deadline. `attained` distinguishes an achievable boundary time from a limiting time excluded by a spell’s termination.

```{r paths}
routes <- paths(dn, from = "Ana")
routes
```

Ana reaches all thirteen other students. Jonas is reached at time 2.12 in one hop; Eve is reached at time 11.66 in four hops through three shortest foremost paths.

`pathways()` groups paths by their sequence of vertices. It accepts a network and an optional source in `from`, and returns a `dynet_pathways` containing `route`, `endpoint`, `count`, `share`, `n_hops`, and `arrival_time`. Routes correspond to leaves of the path tree, so an intermediate prefix is not listed separately. Paths following the same vertex sequence through different spells contribute to the same route count. `share` is the route’s proportion of counted paths, and routes are ordered by decreasing count.

`path_network()` converts a `paths()` result to a static `dynet_path_network` containing the connections used by optimal paths and the vertices they reach. Unreachable vertices are omitted.

`path_trajectories()` represents the paths as a tree whose branches share common initial steps. It accepts a `paths()` result and an optional `min_count` threshold. The returned `dynet_path_trajectories` contains `node`, `parent`, `depth`, `count`, `probability`, `vertex`, `time`, `session`, and `branch`. `probability` expresses a tree node’s count relative to its parent’s count, rather than an empirical probability of transmission.

`plot_path_trajectories()` draws a `paths()` or `path_trajectories()` result. `measure` selects `"frequency"`, `"time"`, or `"predictability"` for fill, and the function returns a `ggplot` object. Labels identify vertices and values.

A vertex with declared activity begins its search at its first eligible time within the search period. A vertex absent throughout that period reaches no other vertex.

## Structure and animation

`projection()` constructs a network of vertex-time states. It accepts the grid arguments and interlayer coupling weight `omega`. The returned `dynet_projection` provides state and edge tables through `as.data.frame(x, what = "vertices")` and `as.data.frame(x, what = "edges")`. Interlayer connections join consecutive slices.

`collapse_network()` aggregates temporal activity into a static weighted network. `weight` selects `"binary"`, `"union_duration"`, `"total_duration"`, `"duration_fraction"`, `"spell_count"`, `"weight_sum"`, `"weighted_duration"`, or `"latest_weight"`. The default is binary presence. The result is a `dynet_collapsed`; with `sessions = "separate"`, a `dynet_collapsed_list` contains a network for each session.

```{r collapse}
static <- collapse_network(dn)
static
```

The collapsed classroom network contains fourteen vertices and 110 connected pairs. Its edge table includes the available weight summaries and the first and last observed contact times for each pair.

`induce_subgraph()` selects vertices through `nodes`, spells through `ties`, or both. It retains the selected network’s attributes and vertex activity. Centralities computed over the observation period are available within vertex-selection conditions.

`animate()` displays successive snapshots using `start`, `end`, `step`, and `window`. `file` selects the output path and encoder; `layout` controls positions, `measure` controls vertex size, and `tween` and `fps` control frame generation. `absent`, `isolates`, and `ease` govern presence and transitions.

The function writes the animation and invisibly returns a `dynet_animation` containing `bin`, `frame`, `time`, `window_start`, `window_end`, `nodes`, `idle`, `ties`, `forming`, `dissolving`, and `file`. GIF output requires `gifski`; MP4 and WebM output require `av`.

## Result classes

Result classes provide methods appropriate to their contents, including printing, summarising, plotting, and conversion to data frames. Use `as.data.frame()` with the documented `what` argument to extract secondary tables.

| Class | Returned by |
|---|---|
| `dynet` | `dynet()`, `as_dynet()`, and editing functions |
| `dynet_metric` | `centrality_series()`, `path_centrality()`, `reachability()`, `metrics()`, `mixing()`, `burstiness()`, `durations()`, `events()` |
| `dynet_paths` | `paths()` |
| `dynet_pathways` | `pathways()` |
| `dynet_path_network` | `path_network()` |
| `dynet_path_trajectories` | `path_trajectories()` |
| `dynet_snapshot` | `snapshots()` |
| `dynet_similarity` | `similarity()` |
| `dynet_pshifts` | `pshifts()` |
| `dynet_projection` | `projection()` |
| `dynet_collapsed` | `collapse_network()` |
| `dynet_collapsed_list` | `collapse_network()` with `sessions = "separate"` |
| `dynet_animation` | `animate()` |

For example, `as.data.frame(x, what = "steps")` extracts individual path steps, `as.data.frame(x, what = "vertex_spells")` extracts declared vertex activity, and `as.data.frame(x, session = "s1")` selects a session where supported.

For measurement functions supporting `plot = TRUE`, plotting is a side effect: the analytical result is still returned, invisibly after the plot is drawn.

## References

Gibson, D. R. (2003). Participation shifts: Order and differentiation in
group conversation. *Social Forces*, 81(4), 1335–1380.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.
