---
title: "Function catalogue"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
# Prefer the working tree, fall back to the installed package, so the article
# builds whether or not it is rendered from inside the repository.
package_root <- normalizePath(file.path(dirname(knitr::current_input(dir = TRUE)),
                                        "../.."), mustWork = FALSE)
if (dir.exists(package_root) && requireNamespace("devtools", quietly = TRUE)) {
  devtools::load_all(package_root, quiet = TRUE)
} else {
  library(Dynet)
}
```

Dynet exports 38 functions in six categories, plus 50 S3 methods across the
result classes. Every verb addresses vertices by name and returns a tidy
one-row-per-observation data frame.

| Category | Verbs | What it does |
|---|---:|---|
| Construction | 4 | Build a network from a relational log; read it back as a time series |
| Editing | 17 | Rewrite spells and vertex activity without breaking time |
| Measurement | 8 | Centrality, reachability, structure, mixing, turn-taking, burstiness, duration, similarity |
| Paths | 5 | Time-respecting paths and four views of the result |
| Structure | 3 | Turn the network into another object |
| Animation | 1 | Render the measurement grid as a film |

## Construction

`dynet()` builds a temporal network from a relational log. It takes a data frame
of ties and a set of column selectors, infers one of four log formats from which
selectors are named, and returns an object of class
`c("dynet", "netobject", "cograph_network")` carrying the spell table, the node
table and the construction metadata. The four formats are interval logs with
explicit start and end times, contact logs of instantaneous events, threaded logs
where an edge stays active until its thread falls silent, and co-presence logs
where actors sharing a group become connected.

```{r construction}
dn <- dynet(school_contacts)
dn
```

`as_dynet()` converts an object of another class to a temporal network. It takes
that object and returns a `dynet`. The method for `networkDynamic` imports edge
spells, vertex spells, the observation period and per-edge attributes. The method
for `dynet` is the identity and returns its input unchanged, so the call is safe
on an object that is already a temporal network.

`events()` counts edge formation and dissolution over time. It takes a network, a
measure name, and the four grid arguments, and returns a `dynet_metric` at graph
level with one row per time point and measure. Eight measures are available, from
raw `"formation"` and `"dissolution"` counts through to the formation and
dissolution rates, which divide confirmed transitions by exact eligible
pair-time.

`snapshots()` slices the network into one edge list per time bin. It takes a
network and the grid arguments and returns a `dynet_snapshot` data frame with one
row per active edge per bin, holding `time`, `from`, `to`, `weight` and
`n_spells`, preceded by `session` when the network has sessions. A pair joined by
more than one spell inside one bin is a single edge, and `n_spells` records how
many spells produced it.

```{r snapshots}
bins <- snapshots(dn, step = 4)
bins
```

## Editing

Seventeen verbs rewrite the spell table, the node table or the vertex-activity
table. Each takes a network, returns a new one of class
`c("dynet", "netobject", "cograph_network")`, and leaves its input unchanged.
Edits go through a rebuild, so canonical spell identifiers can renumber after any
of them.

### Adding

`add_nodes()` adds vertices. It takes a character vector of names or a data frame
with a `name` column and static attributes, and returns a network in which the
new vertices are implicit always-active isolates until ties or vertex activity
are supplied.

`add_ties()` adds tie spells. It takes a data frame of endpoints and times, and
returns a network with the canonical temporal ties and every flattened cograph
field rebuilt together. Columns beyond the canonical spell fields are carried as
tie attributes.

`add_arcs()` is the directed counterpart of `add_ties()` and delegates to it.

`add_vertex_spells()` declares periods in which a vertex is present. It takes a
data frame with `node`, `start` and `end`, and returns a network whose existing
activity and the supplied spells are canonicalised together, so an added spell
that overlaps or abuts an existing one for the same vertex is merged into it.

### Removing

`remove_nodes()` removes vertices. It takes their names and a `cascade` flag, and
returns a network without them and, under `cascade = TRUE`, without their ties
and activity spells. At least one temporal tie must remain.

`remove_ties()` removes tie spells. It takes row positions, a condition over the
spell table, or endpoint and time selectors, and returns a network without the
matched spells. A request that matches nothing raises `dynet_tie_not_found`.

`remove_arcs()` is the directed counterpart of `remove_ties()`.

`remove_vertex_spells()` drops declared activity components. It takes integer
positions or a logical mask and returns a network with the rest canonicalised
again, so the remaining spell identifiers renumber. A vertex left with no
declaration becomes implicitly always active.

### Updating and renaming

`update_nodes()` adds or replaces static vertex attributes. It takes a data frame
with a `name` key and one or more attribute columns, and returns a network with
the same spells, activity and metadata, and the attributes written onto the named
vertices. Unnamed vertices keep their existing values.

`update_ties()` edits selected spells. It takes a selection and a data frame of
replacement values, and returns a network holding the unselected spells unchanged
and the selected ones with the supplied values substituted.

`update_vertex_spells()` edits selected activity components. It takes positions
and a data frame of replacement fields, and returns a network in which the
updated components are canonicalised with the retained ones. A column outside the
vertex-spell schema raises `dynet_unknown_column`.

`rename_nodes()` renames vertices. It takes a named character vector, or an
`old`/`new` data frame, and returns a network with edge endpoints, node
attributes, vertex activity, cograph labels and groups renamed together.

`rename_sessions()` renames session labels. It takes the same mapping forms and
returns a network with edge and vertex session labels renamed together and the
session scheme in the metadata updated. Labels absent from the mapping are left
alone.

### Declaring support

`set_observations()` replaces the observation window. It takes either a
two-column data frame of components or a `start` and `end` pair, and returns a
network whose raw spells are unchanged and whose measurement view alone is
replaced, so `as.data.frame(x)` still returns the originals.

`clear_observations()` restores implicit support. It takes a network and returns
one observed continuously from its earliest raw start through its latest raw end,
with every explicit observation field dropped from the metadata.

`set_tie_sessions()` assigns or removes session walls. It takes a character
vector of length one or of the raw tie count, and returns a network with a
`session` column on the spell table and the scheme recorded in metadata, or with
both removed under `session = NULL`. A full-length vector is matched positionally
against the sorted spell table, so labels should be derived from
`as.data.frame(dn)`.

`set_vertex_spells()` replaces declared vertex activity outright. It takes a
vertex-spell data frame, or the string `"ties"` to declare each vertex present
from its first tie to its last, and returns a network whose declared activity is
exactly that.

```{r editing}
present <- set_vertex_spells(dn, "ties")
present
```

## Measurement

Eight verbs measure a network over time. Each returns a `dynet_metric`: a tidy
data frame whose columns are ordered `session`, `time`, the unit of observation,
`measure` and `value`. The `session` column appears only under
`sessions = "separate"`, which is the one mode that keeps session labels apart.

The grid arguments are not shared by all eight. `dyn_centrality()`, `metrics()`,
`mixing()` and `similarity()` take all four; `dyn_reachability()` and `pshifts()`
take `start` and `end` only; and `burstiness()` and `durations()` summarise the
whole observed period.

`dyn_centrality()` computes vertex centrality. It takes one or more measure names
and a `scope`, and returns one row per vertex, time point and measure. Nineteen
measures are available at `scope = "snapshot"`, which treats each time bin as a
static graph, and four at `scope = "temporal"`, which measures along
time-respecting paths and reports no `time` column because the whole window
yields one value per vertex.

```{r centrality}
between <- dyn_centrality(dn, measure = "betweenness")
between
```

`dyn_reachability()` computes how much of the network each vertex can reach. It
takes a direction and a measure name and returns one row per vertex per measure
at node level. Proportions are labelled `forward_reach` and `backward_reach`,
counts `forward_reach_count` and `backward_reach_count`. A vertex is excluded
from its own reachable set.

`metrics()` computes graph-level structure. It takes one or more of forty measure
names and returns one row per time point and measure. The measure `"triads"`
contributes sixteen rows per time point, one per triad class.

```{r metrics}
density <- metrics(dn, measure = "density")
density
```

`mixing()` computes how much tie activity runs within and between vertex groups.
It takes the name of a node attribute and returns one row per time point and
ordered group pair, with `from_group` and `to_group` naming the pair.

`pshifts()` classifies consecutive turns into Gibson's thirteen
participation-shift types. It takes a directed network and an `output` shape, and
returns a `dynet_pshifts` data frame. Under `output = "final"` that is thirteen
rows, one per shift class, always all thirteen even when a class never occurred.
Under `"cumulative"` it is thirteen rows per classified turn, carrying the
running count.

`burstiness()` measures how unevenly each vertex's events are spaced. It takes a
measure name and returns one row per vertex and measure, with no time column. The
burstiness coefficient is 1 in the bursty limit, 0 at the Poisson reference and
-1 for perfectly regular activity; memory is the lag-1 correlation between
consecutive inter-event gaps.

`durations()` measures how long relationships lasted. It takes a `unit` that
selects the row identity and a measure name, and returns an edge-level table
under `unit = "pair"` or `"spell"`, and a node-level table under the
vertex-activity units. The `censored` argument decides whether spells whose
boundaries were never observed are retained.

`similarity()` compares the network at each pair of time points. It takes a
coefficient name and returns a `dynet_similarity` data frame with one row per
ordered pair of time bins and columns `time`, `other`, `measure` and `value`. The
diagonal is included and is one for every coefficient except `"hamming"`, where
identical layers differ in nothing and score zero.

### The shared grid

Four arguments decide when the network is measured, following
`tsna::tSnaStats()`:

| Argument | Meaning | tsna equivalent |
|---|---|---|
| `start`, `end` | First and last measurement | `start`, `end` |
| `step` | How often to measure | `time.interval` |
| `window` | How much time each measurement covers | `aggregate.dur` |

`window == step` tiles, `window > step` slides, `window == 0` samples an instant,
and `window = "all"` treats the whole observed period as one bin.

## Paths

`paths()` finds time-respecting journeys from one vertex. It takes a network, a
source vertex name, and a direction, and returns a `dynet_paths` data frame with
one row per vertex: `node`, `reachable`, `arrival_time`, `attained`, `latency`,
`n_hops` and `n_paths`. A journey is time-respecting when each hop leaves no
earlier than the previous hop arrived. The criterion is foremost-then-shortest:
earliest arrival first, fewest contacts second.

```{r paths}
routes <- paths(dn, from = "Ana")
routes
```

`pathways()` counts the routes those journeys take. It takes a network,
optionally one source, and returns a `dynet_pathways` data frame with one row per
distinct route, most frequent first: `route` as the vertex sequence joined by
arrows, `endpoint`, `count`, and `share` as the route's fraction of all counted
routes.

`path_network()` builds the union of the optimal routes. It takes a `paths()`
result and returns a static `dynet_path_network`, whose edge table has one row
per hop used by at least one optimal route and whose node table has one row per
reached vertex. Unreachable vertices are absent rather than present with missing
values.

`path_trajectories()` arranges the routes as a counted prefix tree. It takes a
`paths()` result and a `min_count` threshold and returns a
`dynet_path_trajectories` data frame with one row per tree node: `node` as the
route prefix, `parent`, `depth`, `count`, `probability`, `vertex`, `time`,
`session` and `branch`.

`plot_path_trajectories()` draws that tree. It takes a `paths()` or
`path_trajectories()` result and a `measure` that decides what the fill encodes,
and returns a `ggplot` object. Each vertex name is printed beside its value, so
no distinction rests on colour alone.

A vertex with declared spells begins its search at its own first presence inside
the window rather than at the window edge. A vertex never present in the window
reaches nothing.

## Structure and animation

`projection()` expands the network into vertex-time states. It takes the grid
arguments and an interlayer coupling weight `omega`, and returns a
`dynet_projection` holding two tables, reached with
`as.data.frame(x, what = "vertices")` for the states and
`as.data.frame(x, what = "edges")` for the directed arcs between them. Coupling
is ordinal: each slice is joined to the next one only.

`collapse_network()` reduces temporal activity to one static weighted network. It
takes a `weight` rule that decides what an edge weight means, from `"binary"`
through to summed durations and spell counts, and returns a `dynet_collapsed`
cograph netobject with an edge table and a node table. Under
`sessions = "separate"` it returns one such object per session.

```{r collapse}
static <- collapse_network(dn)
static
```

`induce_subgraph()` extracts a subgraph. It takes a condition over the node
table, a condition over the spell table, or both, and returns a network carrying
only the selected spells, the vertices they touch, those vertices' activity
spells, and all static attributes. Centralities are available as columns inside
the `nodes` condition.

`animate()` renders the measurement grid as a film. It takes the same four grid
arguments as every measuring verb, a `file` whose extension selects the encoder,
and returns a `dynet_animation` data frame with one row per bin holding `bin`,
`frame`, `time`, `window_start`, `window_end`, `nodes`, `idle`, `ties`, `forming`
and `dissolving`. The table is returned invisibly, since writing the file is the
purpose of the call. GIF output requires gifski; mp4 and webm require av.

## Result classes

Every result class carries `print`, `summary`, `plot` and `as.data.frame` -- 50
S3 methods in total. A secondary table always comes out through an argument,
never by reaching in with `$`.

| Class | Returned by |
|---|---|
| `dynet` | `dynet()`, `as_dynet()`, and every editing verb |
| `dynet_metric` | `dyn_centrality()`, `dyn_reachability()`, `metrics()`, `mixing()`, `burstiness()`, `durations()`, `events()` |
| `dynet_paths` | `paths()` |
| `dynet_pathways` | `pathways()` |
| `dynet_path_network` | `path_network()` |
| `dynet_path_trajectories` | `path_trajectories()` |
| `dynet_snapshot` | `snapshots()` |
| `dynet_similarity` | `similarity()` |
| `dynet_pshifts` | `pshifts()` |
| `dynet_projection` | `projection()` |
| `dynet_collapsed` | `collapse_network()` |
| `dynet_animation` | `animate()` |

Secondary tables arrive as `as.data.frame(x, what = "steps")`,
`as.data.frame(x, what = "vertex_spells")`,
`as.data.frame(x, session = "s1")` and the like.

On a measuring verb, `plot = TRUE` follows `hist()`: drawing is a side effect and
the tidy result still comes back, invisibly when it has drawn.
