---
title: "The MOOC forum data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
package_root <- normalizePath(file.path(dirname(knitr::current_input(dir = TRUE)),
                                        "../.."), mustWork = FALSE)
if (dir.exists(package_root) && requireNamespace("devtools", quietly = TRUE)) {
  devtools::load_all(package_root, quiet = TRUE)
} else {
  library(Dynet)
}
```

`mooc_posts` records 2,529 posts to the discussion forum of a massive open
online course, and `mooc_people` records the 445 participants who wrote or
received them. The two tables are the data behind chapter 17 of *Learning
Analytics Methods and Tutorials*, and they are the package's worked example of a
**threaded** log: a forum post is not an interval with a stated end, and it is
not an instantaneous contact either. A post stays relevant until its thread
falls silent.

This article builds the network from those posts and reports what it holds.

## The log

```{r log}
head(mooc_posts)
```

Each row is one post: who wrote it, who it replied to, when, and which
discussion it belongs to. There are 338 distinct discussions.

```{r people}
head(mooc_people)
```

`experience` is the self-reported level the chapter recodes into expert,
student and teacher.

## Building the network

`dynet()` reads the log as threaded because `thread =` is named. An edge from a
post stays active until the last post in the same discussion, so a reply late in
a thread extends every earlier tie in it.

```{r build}
dn <- dynet(mooc_posts, from = "sender", to = "receiver",
            time = "timestamp", thread = "discussion")
dn
```

The network has 442 vertices and 2,443 spells. Three of the 445 participants in
`mooc_people` never appear as a sender or a receiver, so they are absent from
the network the log implies.

Times are read as dates and converted to an offset in days from the first post,
which is what `time_unit` reports.

## What the forum looked like over time

`metrics()` measures graph-level structure on the grid. Over the whole observed
period the network's density is:

```{r density}
density <- metrics(dn, measure = "density", window = "all")
density
```

Measured day by day instead, activity is concentrated rather than even:

```{r edges-per-day}
per_day <- metrics(dn, measure = "edges", step = 7, window = 7)
per_day
```

Each row is one week of the course. The counts show where the forum was busy
and where it went quiet.

## Who was central

`dyn_centrality()` measures each vertex on the same grid. Over the whole period,
the ten participants with the highest degree are:

```{r degree}
degree <- dyn_centrality(dn, measure = "degree", window = "all")
head(degree, 10)
```

## Mixing between experience levels

`mooc_people` supplies the attribute the chapter mixes on. The network is built
with that table attached, and `mixing()` then reports how much tie activity ran
within and between levels.

```{r mixing}
labelled <- dynet(mooc_posts, from = "sender", to = "receiver",
                  time = "timestamp", thread = "discussion",
                  nodes = mooc_people)
role_mixing <- mixing(labelled, attribute = "experience", window = "all")
role_mixing
```

Each row is an ordered pair of levels. `value` is the share of tie activity
running from the first group to the second, so the rows where `from_group`
equals `to_group` measure how much the forum kept conversation inside an
experience level.

## Cost

`summary()` on this network takes about 28 seconds, because it measures every
graph-level statistic on all 74 daily bins. `print()` is immediate, and a single
measure on a stated grid — as used throughout this article — costs a few
hundredths of a second. Reach for `metrics()` with the measure you want rather
than `summary()` when the network is this size.

## Source

Kaliisa, R., Gudmundsdottir, G. B., & Jahn, T. (2022). The MOOC forum data used
in chapter 17 of *Learning Analytics Methods and Tutorials*. The bundled tables
are the anonymised sender, receiver, timestamp and discussion columns of that
log, with participant identifiers replaced by integers.
