---
title: "The MOOC forum data"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE,
                      message = FALSE)
```

In this article we describe the bundled MOOC forum data, `mooc_posts` and
`mooc_people`, build the threaded network from them, and read its size, its
activity over time, the degree of its participants and the mixing between
experience levels. The same data carry the case study in
`vignette("ch17-temporal-networks")`.

## The log

Each row of `mooc_posts` is one post: the participant who wrote it, the
participant it answers, the time it was written and the discussion it belongs
to.

```{r log}
library(Dynet)
head(mooc_posts)
nrow(mooc_posts)
```

The log holds 2,529 posts. The timestamps are date-times and the discussion
titles are the thread identifiers.

`mooc_people` has one row per participant with the self-reported experience
level, coded 1 for expert, 2 for student and 3 for teacher, and its label in
`expert_level`, which serves as the mixing attribute.

```{r people}
head(mooc_people)
nrow(mooc_people)
```

There are 445 participants.

## Building the network

To build the network, we call `dynet()` with the log, `from` and `to` for the
endpoint columns, `time` for the timestamp and `thread` for the discussion.
Naming `thread` selects the threaded format: each tie starts at its post and
ends at the last post of the same discussion, so a late reply extends every
earlier tie in its thread. Printing the network gives its format, its size and
the first spells.

```{r build}
dn <- dynet(mooc_posts, from = "sender", to = "receiver",
            time = "timestamp", thread = "discussion")
dn
```

The network has 442 vertices, 2,443 spells and 1,936 distinct ordered pairs,
observed over 73.03 days. Two differences from the log are worth noting. The
spell count is lower than the post count because a post that answers its own
author is a self-loop, and `dynet()` drops self-loops unless `loops = TRUE` is
set. The vertex count is lower than the participant count because three of the
445 participants in `mooc_people` are never a sender or a receiver of a kept
post, so they are not part of the network the log implies. Times are read as
date-times and converted to days since the first post; the unit is stated in
the header of every result.

## Activity over time

Graph-level measures are computed on the grid shared by every measuring
function: `start` and `end` bound the period, `step` is the interval between
measurements and `window` the length of time each one covers. To measure the
network over the whole observed period as a single window, we call `metrics()`
with `measure` set to `"density"` and `window` set to `"all"`.

```{r density}
density <- metrics(dn, measure = "density", window = "all")
density
```

The density over the whole period is 0.0099: about one ordered pair in a
hundred was connected at some point during the course.

To count the ties active in each week, we call `metrics()` with `measure` set
to `"edges"`, and `step` and `window` both set to seven days, so that the bins
tile the period.

```{r edges-per-day}
per_day <- metrics(dn, measure = "edges", step = 7, window = 7)
per_day
```

Each row is one week of the course. The first week holds 451 active ties. The
count rises to 1,013 in the week that begins on day 28, stays above 900 until
the week that begins on day 49, and falls to 231 in the last bin, which begins
on day 70 and covers only the final three days of observation.

## Degree

To obtain the number of distinct contacts of every participant over the whole
period, we call `dyn_centrality()` with `measure` set to `"degree"` and
`window` set to `"all"`. The result has one row per vertex. `head()` prints the
first ten rows, in the order of the vertex table.

```{r degree}
degree <- dyn_centrality(dn, measure = "degree", window = "all")
head(degree, 10)
```

Among the first ten participants, participant 7 has the highest degree with 48
distinct contacts and participant 1 the next with 40; participant 3 has 6.

## Mixing between experience levels

To count the ties within and between experience levels, the network must carry
the level as a vertex attribute. We build it again with `nodes` set to
`mooc_people`, so that the attribute travels with the network, and call
`mixing()` with `attribute` set to `"experience"` and `window` set to `"all"`.

```{r mixing}
labelled <- dynet(mooc_posts, from = "sender", to = "receiver",
                  time = "timestamp", thread = "discussion",
                  nodes = mooc_people)
role_mixing <- mixing(labelled, attribute = "experience", window = "all")
role_mixing
```

Each row is one ordered pair of levels, named by `from_group` and `to_group`,
and `value` is the number of ordered pairs of participants with at least one
active tie from the first level to the second. Within-level counts are 73 for
level 1, 184 for level 2 and 425 for level 3. The three largest counts are the
ties directed into level 3, the teachers: 257 from level 1, 383 from level 2
and 425 from level 3 itself.

## Cost

`summary()` on a network describes it in one table and computes every
graph-level statistic in every daily bin of the observed period. On a network
of this size that is the one slow call; the single-measure calls used in this
article, each on a stated grid, are not. When only one quantity is wanted,
`metrics()` with that measure is the call to make.

## Source

The log is the chapter 17 data of *Learning Analytics Methods and Tutorials*
(Saqr, 2024), taken from the `6_snaMOOC` directory of the book's data
repository at <https://github.com/lamethods/data>. The bundled table keeps the
four columns the chapter reads, `sender`, `receiver`, `timestamp` and
`discussion`, and drops the category hierarchy and comment identifiers of the
published file. `?mooc_posts` documents the preparation.

## References

Saqr, M. (2024). Temporal network analysis: Introduction, methods and analysis
with R. In M. Saqr & S. López-Pernas (Eds.), *Learning analytics methods and
tutorials: A practical guide using R*. Springer.

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319). ACM.
