---
title: "Temporal networks of interactions in course discussions"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6, dpi = 120)
```

This vignette demonstrates temporal network analysis of coded contributions to asynchronous course discussions. The `thought_chains` dataset provides reply relationships in tidy format, with contributions classified into categories such as inquiring, arguing, and approving. Contribution categories constitute the vertices, and directed ties connect the category of a reply to the category of the message it addresses.

The analysis examines how relationships among contribution categories evolve over time, which categories receive replies, and how categories are connected through time-respecting paths. Network measures, centrality indices, relational durations, and temporal paths describe the structure, timing, and reachability of these relationships.

## Data

The dataset is synthetic and based on the original data analysed by [Saqr, López-Pernas, and Törmänen (2026)](https://doi.org/10.1007/s11412-025-09464-5). It is used to demonstrate temporal network construction, measurement, and visualisation. Results reported in this vignette describe the supplied dataset and should not be interpreted as empirical findings from the original study.

The synthetic dataset contains 23,017 reply links among nine contribution categories, spanning 1,169 discussions in 29 groups across five courses. The relational endpoints are `from`, the category assigned to the reply, and `to`, the category assigned to the message it addresses. The variable `time` records when the reply was posted; `discussion`, `course`, and `group` identify the discussion thread, course, and course group, respectively. The dataset includes 9,452 self-links, where the reply and the message it addresses share the same category.

```{r data}
library(Dynet)
head(thought_chains)
```

## The network

Each reply is represented as a relational spell with onset at the time of posting and termination at the end of its discussion. This specification treats the relationship as active for the remaining duration of the discussion.

The call to `dynet()` explicitly identifies `discussion` as the thread variable through `thread = "discussion"`. Setting `thread_clock = "relative"` expresses time in days since the first recorded reply in each discussion, while `loops = TRUE` retains self-links and `observation_end = 4` specifies a four-day observation period. The constructor automatically recognises `course` as the session variable.

```{r import}
dn <- dynet(thought_chains, thread = "discussion", thread_clock = "relative",
            loops = TRUE, observation_end = 4)
dn
summary(dn)
```

Ties are directed from the category of a reply to the category of the message it addresses. Paths following these ties therefore represent directed reply relationships between contribution categories. They do not, by themselves, establish the forward dissemination of ideas or the transmission of particular content.

Setting `thread_clock = "relative"` aligns each discussion to its first recorded reply. Temporal overlap consequently represents comparable elapsed times within discussions, rather than simultaneous activity on a shared calendar. The constructor recognises `course` as the session variable. With the default `sessions = "bounded"`, each temporal path remains within a single course, although it may combine ties from different discussions or groups within that course. These paths describe category-level reachability under the specified alignment and session boundaries.

The resulting temporal network contains nine vertices and 23,017 relational spells distributed across 80 of the 81 possible ordered pairs, including self-pairs. The mean snapshot density is 0.8958. This high density reflects the specification that relational spells remain active until their discussions terminate.

## Ties over time

`events()` computes relational spell onset and termination counts over successive half-day intervals. Setting both `step` and `window` to 0.5 produces non-overlapping intervals.

```{r events}
tie_events <- events(dn, measure = c("formation", "dissolution"),
                     step = 0.5, window = 0.5)
tie_events
```

The first three intervals contain 13,094, 4,435, and 2,442 onsets, respectively, compared with 10 in the final interval. Terminations peak in the third interval, with 5,348 spells ending. Relational spell onsets are therefore concentrated early in the observation period.

The timeline displays relational activity for the 40 ordered pairs with the most spells, using an hourly grid. Colour represents the proportion of each interval during which the pair was active, counting overlapping spells once.

```{r timeline, fig.height=8}
plot(dn, type = "timeline", step = 1 / 24)
```

The activity plot displays onset and termination counts alongside the number of active relational spells over time.

```{r activity}
plot(dn, type = "activity")
```

## Drawing the network

Setting `type = "network"` displays the union of ties active during the observation period. The network is drawn using cograph, to which the `layout` argument is passed.

```{r network, fig.height=8}
plot(dn, type = "network", layout = "oval")
```

The binary union records whether each ordered pair was connected at any time during the observation period. To account for duration, `collapse_network()` with `weight = "union_duration"` weights each pair by the total duration for which at least one relational spell was active, counting overlapping spells once.

```{r collapsed, fig.height=8}
collapsed <- collapse_network(dn, weight = "union_duration")
plot(collapsed, layout = "oval")
```

Setting `type = "snapshots"` displays the network at nine equally spaced time points using a common layout.

```{r snapshots, fig.height=9, fig.width=10}
plot(dn, type = "snapshots", panels = 9)
```

The proximity timeline represents changes in the relationships among contribution categories. Network distances are computed within overlapping temporal slices and reduced to a single coordinate for each category. Lines connect these coordinates across slices, with proximity indicating shorter network distances within a slice. Line thickness represents degree by default. The following call specifies 80 slices and five accompanying phase networks.

```{r proximity, fig.height=8, fig.width=12}
plot(dn, type = "proximity", phases = 5, slices = 80)
```

Setting `measure = "betweenness"` maps line thickness to betweenness centrality, while `networks = FALSE` omits the phase networks.

```{r proximity-lines, fig.height=7, fig.width=12}
plot(dn, type = "proximity", measure = "betweenness",
     networks = FALSE, slices = 80)
```

Layered visualisations represent successive temporal slices as separate network layers. Setting `type = "layers"` and `step = 4 / 3` displays three slices as planes.

```{r layers, fig.height=9, fig.width=13}
plot(dn, type = "layers", step = 4 / 3, layout = "circle")
```

Setting `type = "stack"` and `cumulative = TRUE` produces a cumulative projection in which each plane includes all ties active up to that slice.

```{r stack, fig.height=10, fig.width=13}
plot(dn, type = "stack", step = 4 / 3, cumulative = TRUE)
```

## Structure over time

Dynet computes both graph-level and vertex-level metrics for temporal networks. Graph-level metrics characterise the network as a whole, describing properties such as density, reciprocity, and connectivity. Vertex-level metrics characterise the structural position of individual vertices through measures such as degree, closeness, and betweenness. In this vignette, vertices represent contribution categories, so vertex-level metrics describe the positions of these categories within the discussion network. Both levels can be examined over successive temporal intervals to quantify changes in network structure and vertex position.

`metrics()` computes graph-level measures over specified temporal intervals. Setting both `step` and `window` to `1 / 24` produces successive, non-overlapping hourly intervals. Density measures the proportion of possible directed ties present, while reciprocity measures the proportion of directed ties whose reverse is also present. Efficiency and connectedness are two of Krackhardt’s (1994) indices for characterising departures from an out-tree structure. The number of edges and mean degree provide additional descriptions of network connectivity.

```{r structure}
structure <- metrics(dn,
                     measure = c("density", "reciprocity", "efficiency",
                                 "connectedness", "edges", "mean_degree"),
                     step = 1 / 24, window = 1 / 24)
structure
plot(structure, type = "ridge")
```

Setting `type = "ridge"` displays a separate panel for each measure. In the first hour, the network contains 57 directed ties and has a density of 0.79. Connectedness is 1, indicating that all contribution categories belong to a single weakly connected component; this does not imply that every category is directly connected to every other category. Reciprocity is 0.91, indicating that most directed ties have a corresponding reverse tie within the interval. Density increases to 0.82 in the second hour.

The dyad census (Holland and Leinhardt, 1976) classifies unordered pairs of distinct vertices as mutual, asymmetric, or null, depending on whether both, one, or neither of the possible directed ties are present. The corresponding counts are requested through `measure`.

```{r dyads}
dyads <- metrics(dn, measure = c("mutual", "asymmetric", "null"),
                 step = 1 / 24, window = 1 / 24)
dyads
plot(dyads)
```

In the first hour, the 36 unordered pairs comprise 26 mutual, five asymmetric, and five null dyads. These counts describe connectivity between distinct contribution categories and exclude self-links.

Setting `measure = "triads"` computes the sixteen-class directed triad census, which classifies sets of three distinct vertices according to their directed ties. The heatmap displays changes in the frequency of each triad class across hourly intervals.

```{r triads, fig.height=8}
triads <- metrics(dn, measure = "triads", step = 1 / 24, window = 1 / 24)
triads
plot(triads, type = "heatmap")
```

Additional graph-level statistics describe local configurations of directed ties. These include two-stars, two-paths, triangles, and sums of indegrees and outdegrees raised to the power 1.5. Such statistics are also used as terms in exponential-family random graph models (Morris, Handcock and Hunter, 2008).

```{r ergm}
local <- metrics(dn,
                 measure = c("indegree_1_5", "outdegree_1_5", "triangles",
                             "in_2stars", "out_2stars", "two_paths"),
                 step = 1 / 24, window = 1 / 24)
local
plot(local, type = "ridge")
```

The first hourly interval contains 376 directed triangles and 325 two-paths; the corresponding counts in the second interval are 436 and 356.

## Vertex-level metrics

`centrality_series()` computes centrality measures for each contribution category over specified temporal intervals. The following call requests degree, closeness, and betweenness (Freeman, 1979), eigenvector centrality, flow betweenness (Freeman, Borgatti and White, 1991), and diffusion degree (Kundu, Murthy and Pal, 2011). These measures characterise the positions of categories within each hourly network. Paths used in these interval-specific calculations are evaluated within the corresponding network; temporal centrality based on time-respecting paths is considered separately below.

```{r centrality, fig.height=11}
centrality <- centrality_series(dn,
                                measure = c("degree", "closeness", "betweenness",
                                            "eigenvector", "flow_betweenness",
                                            "diffusion"),
                                step = 1 / 24, window = 1 / 24)
centrality
plot(centrality, type = "heatmap")
```

Setting `type = "heatmap"` displays centrality values by category and hourly interval. Setting `type = "ridge"` displays their trajectories over the observation period.

```{r centrality-ridge, fig.height=10}
plot(centrality, type = "ridge")
```

In the first hour, `Socialising` has the highest degree (18), and `Objecting` has the lowest (7). `Arguing` has a closeness centrality of 1, indicating that it reaches every other category in one step under the specified calculation. Across the 96 hourly intervals, `Objecting` has the lowest mean degree, closeness, and betweenness.

For directed networks, `mode` specifies the direction used to calculate degree. Because ties are directed from the reply category to the category being addressed, `mode = "in"` measures incoming connections from replying categories, whereas `mode = "out"` measures outgoing connections to addressed categories.

```{r directed-degree, fig.height=7}
in_degree <- centrality_series(dn, measure = "degree", mode = "in",
                               step = 1 / 24, window = 1 / 24)
out_degree <- centrality_series(dn, measure = "degree", mode = "out",
                                step = 1 / 24, window = 1 / 24)
plot(in_degree, type = "heatmap")
plot(out_degree, type = "heatmap")
```

Prestige indices characterise a vertex through its incoming relationships (Wasserman and Faust, 1994). Indegree prestige measures direct incoming connections. Domain proximity prestige incorporates indirect connections by dividing the proportion of other vertices that can reach the focal vertex by their mean directed distance to it. Setting `measure = "prestige"` requests prestige, with the variant specified through `prestige`.

```{r prestige, fig.height=7}
prestige_indegree <- centrality_series(dn, measure = "prestige",
                                       prestige = "indegree",
                                       step = 1 / 24, window = 1 / 24)
prestige_proximity <- centrality_series(dn, measure = "prestige",
                                        prestige = "domain.proximity",
                                        step = 1 / 24, window = 1 / 24)
plot(prestige_indegree, type = "heatmap")
plot(prestige_proximity, type = "heatmap")
```

## Flows between codes

`mixing()` summarises directed relationships between categories of a vertex attribute within each temporal interval. Setting `attribute = "name"` uses the vertex names, which represent contribution categories in this network. The result contains counts of distinct active ordered vertex pairs for each hourly interval and pair of categories. Repeated spells and their weights do not multiply these counts. Setting `type = "heatmap"` displays all 81 ordered pairs, including self-pairs.

```{r mixing, fig.height=14}
flows <- mixing(dn, attribute = "name", step = 1 / 24, window = 1 / 24)
flows
plot(flows, type = "heatmap")
```

The `highlight` argument selects a category for visual emphasis. The following plot displays incoming and outgoing relationships involving `Coordinating` in colour and the remaining relationships in grey.

```{r mixing-highlight, fig.height=7}
plot(flows, highlight = "Coordinating")
```

## Duration

`durations()` summarises relational spells and their durations over the observation period. Setting `unit = "node_ties"` groups spells by their incident vertex, while `mode = "all"` includes both incoming and outgoing incidences. The requested measures are the incident spell count (`"events"`), summed incident duration (`"total"`), and union duration (`"union"`). A self-link contributes twice to the additive counts and durations, once for each endpoint. Summed duration includes overlapping spells separately, whereas union duration counts each period of activity once.

```{r node-ties, fig.height=10}
node_ties <- durations(dn, unit = "node_ties", mode = "all",
                       measure = c("events", "total", "union"))
node_ties
plot(node_ties)
```

`Resourcing` has 20,999 incident spell counts with a summed duration of 22,781 days, compared with 109 and 71 days for `Objecting`. These summed durations accumulate time across spell incidences and can therefore exceed the observation period. Union duration is bounded by the four-day observation period; six of the nine categories have at least one incident spell active throughout this period.

Setting `unit = "pair"` computes the same summaries for each ordered pair of contribution categories. Including `"mean"` additionally returns the mean spell duration.

```{r pairs, fig.height=10}
pair_duration <- durations(dn, unit = "pair",
                           measure = c("events", "total", "union", "mean"))
pair_duration
plot(pair_duration)
```

Spell counts vary substantially across ordered pairs. For example, 1,353 spells connect `Approving` to `Resourcing`, compared with four connecting `Approving` to `Objecting`.

## Reachability and temporal centrality

Temporal reachability describes whether vertices can be connected through time-respecting paths, whose successive traversals follow chronological order and respect tie availability (Kempe, Kleinberg and Kumar, 2002). `reachability()` computes the proportion (`"reach"`) and number (`"reach_count"`) of other vertices reachable along these paths. Setting `direction = "both"` requests both forward reachability and backward reachability: which categories a focal category can reach and which can reach it.

```{r reachability, fig.height=8}
reach <- reachability(dn, direction = "both",
                      measure = c("reach", "reach_count"))
reach
plot(reach)
```

Every category can reach all eight other categories and can be reached from all eight. Under the specified spell durations and observation period, reachability therefore does not differentiate the categories. Temporal centrality and path analysis provide further information about arrival times and the routes connecting them.

Temporal closeness measures how quickly a vertex can reach other vertices, whereas temporal betweenness measures its contribution to earliest-arrival paths between other vertices (Pan and Saramäki, 2011). These measures are requested through `path_centrality()`.

When traversal has zero duration, paths composed of ties active at time zero can arrive immediately, yielding zero latency and potentially unbounded temporal closeness. Setting `traversal_time = 0.1` imposes a traversal duration of one tenth of a day per tie. This is an analytical assumption rather than an observed response time.

```{r temporal, fig.height=9}
temporal <- path_centrality(dn, measure = c("closeness", "betweenness"),
                            traversal_time = 0.1)
temporal
plot(temporal)
```

`Tutoring` has the highest temporal closeness (10), corresponding to the reciprocal of the specified traversal duration: it reaches every other category directly without waiting. `Objecting` has the lowest temporal closeness (5.5). Temporal betweenness is concentrated in `Socialising`, `Arguing`, and `Resourcing`, whereas `Approving`, `Drafting`, and `Objecting` are not intermediate vertices on any earliest-arrival path between other categories.

## Paths

`paths()` identifies earliest-arrival paths from the category specified by `from`, minimising hop count among paths with the same earliest arrival. The following call uses the same traversal duration as the temporal centrality analysis. In the result, `arrival_time` records the earliest arrival at each destination, `n_hops` records the number of traversed ties, and `n_paths` records the number of optimal paths. When the optimal session is unique, `path_session` identifies the course containing the ties that achieve that arrival time. The plot displays the resulting paths as a tree.

```{r paths, fig.height=8, fig.width=12}
from_inquiring <- paths(dn, from = "Inquiring", traversal_time = 0.1)
from_inquiring
plot(from_inquiring)
```

From `Inquiring`, every category is reachable in one hop. Arrival at `Arguing`, `Drafting`, `Resourcing`, and `Socialising` occurs at day 0.1, requiring only the specified traversal duration. `Objecting` is reached last, at day 0.189. Nine equally early paths reach `Resourcing`.

Backward path analysis identifies the latest departure times that permit arrival at a focal category by a specified deadline. Setting `direction = "backward"` makes `from` the focal destination, while `start` and `end` delimit the observation period. In this result, `arrival_time` contains the supremum of feasible departure times, and `latency` is the difference between the deadline and that value. An excluded interval endpoint may yield a supremum that cannot itself be attained; this is indicated by `attained = FALSE`.

```{r backward, fig.height=8, fig.width=12}
into_approving <- paths(dn, from = "Approving", direction = "backward",
                        start = 0, end = 4, traversal_time = 0.1)
into_approving
plot(into_approving)
```

`Arguing` and `Resourcing` can depart at day 3.9 and reach `Approving` by day 4. In contrast, the latest feasible departure from `Objecting` is day 2.83, corresponding to a latency of 1.17 days. This difference reflects the availability of ties towards the end of the observation period.

`pathways()` ranks routes across source vertices, with `top` specifying the number returned. Counts represent optimal paths grouped by vertex sequence, rather than observed frequencies of message sequences. The plot represents each category along a route at its arrival time. This call uses the default zero traversal duration.

```{r pathways, fig.height=8, fig.width=12}
top_routes <- pathways(dn, top = 12)
top_routes
plot(top_routes)
```

## Timing

`similarity()` compares network structure across temporal intervals. Setting both `step` and `window` to 0.5 produces non-overlapping half-day intervals. The default Jaccard coefficient measures similarity between their tie sets, and the plot displays the pairwise comparisons as a heatmap.

```{r similarity, fig.height=7, fig.width=8}
bin_similarity <- similarity(dn, step = 0.5, window = 0.5)
plot(bin_similarity)
```

`pshifts()` classifies consecutive directed interactions using Gibson’s (2003) thirteen participation shifts. These distinguish whether an interaction from A to B is followed by an interaction initiated by B, by A again, or by another vertex X. Here, vertices represent contribution categories, so the classifications describe changes in category-level relational endpoints.

```{r pshifts, fig.height=6, fig.width=10}
shifts <- pshifts(dn)
shifts
plot(shifts)
```

Of the 3,423 classified transitions, the largest family is turn usurping, in which a third category becomes the source of the next interaction. The most frequent individual shift is AB-XB, with 642 occurrences: a different source category addresses the same target category.

`burstiness()` characterises the temporal distribution of activity using inter-event intervals. The burstiness coefficient approaches 1 for highly heterogeneous intervals, is 0 for an exponential inter-event distribution, and is −1 for equal intervals (Goh and Barabási, 2008).

```{r burstiness, fig.height=6, fig.width=10}
bursts <- burstiness(dn)
summary(bursts)
plot(bursts)
```

All categories have positive burstiness coefficients, ranging from 0.43 for `Objecting` to 0.74 for `Resourcing`. The memory coefficients are close to zero, indicating little correlation between successive inter-event intervals.

## One course group

The `group` variable is retained as a spell attribute during network construction. A course group can therefore be selected by filtering the existing temporal network. `induce_subgraph()` with `ties = group == "A_01"` retains spells associated with group `A_01`.

```{r group}
one_group <- induce_subgraph(dn, ties = group == "A_01")
summary(one_group)
```

The resulting subnetwork contains 2,425 spells across 53 of the 81 possible ordered pairs, with a mean snapshot density of 0.559 and one session.

Setting `type = "events"` displays reply events with contribution categories on the vertical axis. Specifying `time = "clock"` positions events at their recorded times, while `blend = TRUE` interpolates link colours between their source and target categories.

```{r group-events, fig.height=8, fig.width=13}
plot(one_group, type = "events")
```

```{r group-events-clock, fig.height=8, fig.width=13}
plot(one_group, type = "events", time = "clock", blend = TRUE)
```

Graph-level metrics can be computed for the subnetwork using the same hourly intervals as the full network.

```{r group-metrics}
group_structure <- metrics(one_group,
                           measure = c("density", "edges", "reciprocity",
                                       "connectedness"),
                           step = 1 / 24, window = 1 / 24)
plot(group_structure, type = "ridge")
```

## Interpretation

In this synthetic example, `Resourcing` is incident to the largest number of reply links and is frequently reachable through multiple equally early paths. Together with `Socialising` and `Arguing`, it accounts for much of the temporal betweenness. `Objecting` has comparatively few incident replies, the lowest temporal closeness, and no temporal betweenness.

Under the specified construction, relational spells remain active until discussion termination, and every category is temporally reachable from every other category. Differences in arrival times and intermediate categories provide additional distinctions that reachability alone does not capture. These results illustrate the analytical methods applied to the synthetic dataset and do not constitute empirical findings about the original study.

## Limitations

Maintaining each relational spell until discussion termination is a modelling assumption that increases tie availability and contributes to high network density and complete reachability. Shorter spell durations could produce sparser networks, longer waiting times, or unreachable destinations. The four-day observation period also truncates longer discussions for measurement, without modifying their raw spell endpoints.

Vertices represent contribution categories rather than individual participants; consequently, the analysis characterises relationships among categories and does not identify interpersonal interaction patterns. Replies assigned multiple categories may contribute multiple ties, affecting spell counts and accumulated durations. Temporal path results additionally depend on the specified traversal duration.

The dataset is synthetic and based on the original study data. Its results serve to demonstrate package functionality; correspondence with empirical patterns in the original data requires separate assessment.

## References


Freeman, L. C. (1979). Centrality in social networks: conceptual
clarification. *Social Networks*, 1(3), 215–239.

Freeman, L. C., Borgatti, S. P., & White, D. R. (1991). Centrality in valued
graphs: a measure of betweenness based on network flow. *Social Networks*,
13(2), 141–154.

Gibson, D. R. (2003). Participation shifts: order and differentiation in group
conversation. *Social Forces*, 81(4), 1335–1380.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Holland, P. W., & Leinhardt, S. (1976). Local structure in social networks.
*Sociological Methodology*, 7, 1–45.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.

Krackhardt, D. (1994). Graph theoretical dimensions of informal organizations.
In K. M. Carley & M. J. Prietula (Eds.), *Computational organization theory*
(pp. 89–111). Lawrence Erlbaum.

Kundu, S., Murthy, C. A., & Pal, S. K. (2011). A new centrality measure for
influence maximization in social networks. In *Pattern Recognition and Machine
Intelligence* (Lecture Notes in Computer Science 6744, pp. 242–247). Springer.

Morris, M., Handcock, M. S., & Hunter, D. R. (2008). Specification of
exponential-family random graph models: terms and computational aspects.
*Journal of Statistical Software*, 24(4).

Pan, R. K., & Saramäki, J. (2011). Path lengths, correlations, and centrality
in temporal networks. *Physical Review E*, 84(1), 016105.

Saqr, M., López-Pernas, S., & Törmänen, T. (2026). A temporal network approach to reveal the longitudinal dynamics of CSCL group regulation and productive collaboration. *International Journal of Computer-Supported Collaborative Learning*, 21, 237–270. https://doi.org/10.1007/s11412-025-09464-5

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319). ACM.

Wasserman, S., & Faust, K. (1994). *Social network analysis: methods and
applications*. Cambridge University Press.
