---
title: "Case study: How kinds of contribution follow one another in course discussions"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6, dpi = 120)
```

In this article we analyse the reply table of the *Trees of Thought* study,
which coded the messages of asynchronous course discussions into kinds of
contribution such as inquiring, arguing or approving. The unit of analysis is
the code rather than the student: a tie runs from the code of a reply to the
code of the message it answers. We ask which kinds of contribution the
discussions return to, how quickly a contribution of one kind can lead to any
other, and which kinds sit on the routes between the rest.

## Data

`thought_chains` is the study's reply table, anonymised: 23,017 reply links
among nine codes, from 1,169 discussions in 29 groups across five courses.
Each row is one reply. `from` is the code of the reply, `to` the code of the
message it answers, `time` the moment of the reply, `discussion` the thread it
belongs to, and `course` and `group` the course and course group. A code that
answers itself is a self-link; there are 9,452 of them.

```{r data}
library(Dynet)
head(thought_chains)
```

## The network

A reply is a contribution to a conversation that is still open, so, following
Saqr and Nouri (2020), it becomes a tie that opens when the reply is posted
and closes when its discussion ends. To build the network, we call `dynet()`
with `thread` for the discussion column, `thread_clock = "relative"` so that
time is counted in days from the first post of each discussion, `loops = TRUE`
to keep self-links, and `observation_end = 4` for a four-day observation
window. The `course` column is recognised as the session column.

```{r import}
dn <- dynet(thought_chains, thread = "discussion", thread_clock = "relative",
            loops = TRUE, observation_end = 4)
dn
summary(dn)
```

The network has nine vertices and 23,017 spells on 80 of the 81 possible
ordered pairs of codes. Time is in days, and on an average day nine of every
ten ordered pairs are connected (mean snapshot density 0.8958). The network is
dense because a tie stays open for the life of its discussion.

## Ties over time

To count the ties that open and close in each half day, we call `events()`
with `step` and `window` of 0.5.

```{r events}
tie_events <- events(dn, measure = c("formation", "dissolution"),
                     step = 0.5, window = 0.5)
tie_events
```

13,094 ties open in the first half day, 4,435 in the second and 2,442 in the
third; only 10 open in the last. Closures peak in the third half day, at
5,348, when the shorter discussions end. Most of the network is therefore in
place from the start, and the analysis below asks what happens within it.

To draw the spells over time, we call `plot()` with `type = "timeline"`; each
bar is one spell, on an hourly grid, and the forty busiest pairs are drawn.

```{r timeline, fig.height=8}
plot(dn, type = "timeline", step = 1 / 24)
```

To draw the counts of opening and closing ties with the stock of active ties
beneath them, we set `type` to `"activity"`.

```{r activity}
plot(dn, type = "activity")
```

## Drawing the network

To draw the union of every tie active in the window, we set `type` to
`"network"`; the drawing is cograph's and `layout` is passed to it.

```{r network, fig.height=8}
plot(dn, type = "network", layout = "oval")
```

The binary union treats a tie active for an hour like one active for four
days. To weight each pair by the time it was active, counting overlapping
spells once, we call `collapse_network()` with `weight = "union_duration"` and
draw the static network it returns.

```{r collapsed, fig.height=8}
collapsed <- collapse_network(dn, weight = "union_duration")
plot(collapsed, layout = "oval")
```

To draw the network at nine equally spaced cross-sections with a shared
layout, we set `type` to `"snapshots"`.

```{r snapshots, fig.height=9, fig.width=10}
plot(dn, type = "snapshots", panels = 9)
```

The proximity timeline follows each code through time: the window is measured
in overlapping slices, the distances in each slice are reduced to one
dimension, and each code is a line whose height follows that coordinate, so
codes that interact sit close together. Line thickness follows degree. To draw
it with 80 slices and five phase networks, we set `type` to `"proximity"`.

```{r proximity, fig.height=8, fig.width=12}
plot(dn, type = "proximity", phases = 5, slices = 80)
```

To let thickness follow betweenness and omit the phase networks, we set
`measure` and `networks`.

```{r proximity-lines, fig.height=7, fig.width=12}
plot(dn, type = "proximity", measure = "betweenness",
     networks = FALSE, slices = 80)
```

Cutting the window into slices and treating each slice as one layer of a
multilayer network shows structure and time together. To draw three slices as
planes, we set `type` to `"layers"` and `step` to `4 / 3`; to draw them as
matrices keeping only cells with at least 50 active dyads, we set `type` to
`"heatmap"` and `threshold` to 50; to draw them as a projection in which each
plane carries every tie active up to it, we set `type` to `"stack"` and
`cumulative` to `TRUE`.

```{r layers, fig.height=9, fig.width=13}
plot(dn, type = "layers", step = 4 / 3, layout = "circle")
```

```{r heatmap-threshold, fig.height=9, fig.width=13}
plot(dn, type = "heatmap", step = 4 / 3, threshold = 50)
```

```{r stack, fig.height=10, fig.width=13}
plot(dn, type = "stack", step = 4 / 3, cumulative = TRUE)
```

## Structure over time

To measure the network in each hour, we call `metrics()` with `step` and
`window` of `1 / 24`. Density is the share of ordered pairs connected in the
hour, reciprocity the share of arcs whose reverse is also present, and
efficiency and connectedness two of Krackhardt's (1994) indices of how far a
directed network departs from an out-tree. To draw one panel per measure, we
call `plot()` with `type = "ridge"`.

```{r structure}
structure <- metrics(dn,
                     measure = c("density", "reciprocity", "efficiency",
                                 "connectedness", "edges", "mean_degree"),
                     step = 1 / 24, window = 1 / 24)
structure
plot(structure, type = "ridge")
```

In the first hour, 57 arcs give a density of 0.79, connectedness is 1, so every
code is joined to every other, and reciprocity is 0.91: almost every transition
between two codes occurs in both directions within the same hour. An hour
later the density is 0.82.

To obtain the dyad census (Holland and Leinhardt, 1976), we name the mutual,
asymmetric and null counts in `measure`; to obtain the sixteen-class triad
census, we set `measure` to `"triads"`.

```{r dyads}
dyads <- metrics(dn, measure = c("mutual", "asymmetric", "null"),
                 step = 1 / 24, window = 1 / 24)
dyads
plot(dyads)
```

```{r triads, fig.height=8}
triads <- metrics(dn, measure = "triads", step = 1 / 24, window = 1 / 24)
triads
plot(triads, type = "heatmap")
```

Of the 36 unordered pairs of codes, 26 are mutual in the first hour, 5
asymmetric and 5 null. No triad is empty; nine are of class 102, one mutual
dyad with two null dyads, eight of class 111U and seven of class 201.

To obtain the two-star, two-path and triangle counts and the degree sums raised
to 1.5 used as terms in exponential-family random graph models (Morris,
Handcock and Hunter, 2008), we name them in `measure`.

```{r ergm}
local <- metrics(dn,
                 measure = c("indegree_1_5", "outdegree_1_5", "triangles",
                             "in_2stars", "out_2stars", "two_paths"),
                 step = 1 / 24, window = 1 / 24)
local
plot(local, type = "ridge")
```

The first hour holds 376 directed triangles and 325 two-paths; the second, 436
and 356.

## Codes

To measure each code in each hour, we call `dyn_centrality()` with the indices
in `measure`: degree, closeness and betweenness (Freeman, 1979), eigenvector
centrality, flow betweenness (Freeman, Borgatti and White, 1991) and diffusion
degree (Kundu, Murthy and Pal, 2011). `plot()` with `type = "heatmap"` draws
one tile per code and hour, and `type = "ridge"` draws the trajectories as
lines.

```{r centrality, fig.height=11}
centrality <- dyn_centrality(dn,
                             measure = c("degree", "closeness", "betweenness",
                                         "eigenvector", "flow_betweenness",
                                         "diffusion"),
                             step = 1 / 24, window = 1 / 24)
centrality
plot(centrality, type = "heatmap")
```

```{r centrality-ridge, fig.height=10}
plot(centrality, type = "ridge")
```

In the first hour `Socialising` has the highest degree, 18, and `Objecting`
the lowest, 7; `Arguing` has a closeness of 1, so it reaches every other code
in one step. `Objecting` stays at the margin throughout: it has the lowest
mean degree, closeness and betweenness over the 96 hours.

On a directed network the direction of degree is the `mode` argument.

```{r directed-degree, fig.height=7}
in_degree <- dyn_centrality(dn, measure = "degree", mode = "in",
                            step = 1 / 24, window = 1 / 24)
out_degree <- dyn_centrality(dn, measure = "degree", mode = "out",
                             step = 1 / 24, window = 1 / 24)
plot(in_degree, type = "heatmap")
plot(out_degree, type = "heatmap")
```

Prestige indices describe a code by the ties it receives (Wasserman and Faust,
1994). Indegree prestige counts the distinct codes with a tie into a code;
domain proximity prestige discounts the share of codes with a directed path
into it by their mean distance. To obtain them, we set `measure` to
`"prestige"` and name the variant in `prestige`.

```{r prestige, fig.height=7}
prestige_indegree <- dyn_centrality(dn, measure = "prestige",
                                    prestige = "indegree",
                                    step = 1 / 24, window = 1 / 24)
prestige_proximity <- dyn_centrality(dn, measure = "prestige",
                                     prestige = "domain.proximity",
                                     step = 1 / 24, window = 1 / 24)
plot(prestige_indegree, type = "heatmap")
plot(prestige_proximity, type = "heatmap")
```

## Flows between codes

To count the active dyads from each code to each other code in each hour, we
call `mixing()` with `attribute = "name"`. Each row is one hour and one ordered
pair of codes, and `plot()` with `type = "heatmap"` draws all 81 pairs.

```{r mixing, fig.height=14}
flows <- mixing(dn, attribute = "name", step = 1 / 24, window = 1 / 24)
flows
plot(flows, type = "heatmap")
```

To follow one code's flows against the rest, we call `plot()` with
`highlight` naming the code. The flows into and out of `Coordinating` are
drawn in colour and the others in grey.

```{r mixing-highlight, fig.height=7}
plot(flows, highlight = "Coordinating")
```

## Duration

To measure how much tie time each code carries, we call `durations()` with
`unit = "node_ties"`, `mode = "all"` for ties in either direction, and
`measure` for the number of spells, their summed duration and the union of
their active time, which counts overlapping spells once.

```{r node-ties, fig.height=10}
node_ties <- durations(dn, unit = "node_ties", mode = "all",
                       measure = c("events", "total", "union"))
node_ties
plot(node_ties)
```

`Resourcing` is incident to 20,999 spells with a summed duration of 22,781
days, against 109 spells and 71 days for `Objecting`. The union column shows
the ceiling: no code is in contact for more than the four days of the window,
and six of the nine are in contact for all of it.

To obtain the same accounting per ordered pair of codes, with the mean spell
length added, we set `unit` to `"pair"`.

```{r pairs, fig.height=10}
pair_duration <- durations(dn, unit = "pair",
                           measure = c("events", "total", "union", "mean"))
pair_duration
plot(pair_duration)
```

The pairs differ by three orders of magnitude: 1,353 spells run from
`Approving` to `Resourcing` and 4 from `Approving` to `Objecting`.

## Reachability and temporal centrality

A time-respecting path may only use ties in chronological order (Kempe,
Kleinberg and Kumar, 2002). To obtain the share and the number of other codes
each code can reach, and be reached from, along such paths, we call
`dyn_reachability()` with `direction = "both"`.

```{r reachability, fig.height=8}
reachability <- dyn_reachability(dn, direction = "both",
                                 measure = c("reach", "reach_count"))
reachability
plot(reachability)
```

Every code reaches all eight others and is reached by all eight. With so many
ties open from the start this is expected, and the informative quantities are
how quickly and by how many routes.

Temporal closeness is the inverse of the mean time a code needs to reach the
others, and temporal betweenness the share of earliest routes between other
codes that pass through a code (Pan and Saramäki, 2011). Because the ties that
carry the routes are open at time 0, a hop that costs nothing arrives at once
and closeness is unbounded; `traversal_time = 0.1` charges a tenth of a day
per hop so that the indices separate. To obtain them, we call
`dyn_centrality()` with `scope = "temporal"`.

```{r temporal, fig.height=9}
temporal <- dyn_centrality(dn, measure = c("closeness", "betweenness"),
                           scope = "temporal", traversal_time = 0.1)
temporal
plot(temporal)
```

`Tutoring` has the highest temporal closeness, 10, the reciprocal of the
traversal cost: it reaches every other code in one hop without waiting.
`Objecting` has the lowest, 5.5. Betweenness is concentrated in `Socialising`,
`Arguing` and `Resourcing`, while `Approving`, `Drafting` and `Objecting`
relay no earliest route at all.

## Paths

To obtain the earliest route from one code to every other, we call `paths()`
with `from` for the source and the same traversal cost. `arrival_time` is the
earliest moment a code can be reached, `n_hops` the length of the earliest
route and `n_paths` the number of routes that arrive equally early;
`path_session` names the course whose ties achieve the optimum when it is
unique. To draw the routes as a tree, we call `plot()` on the result.

```{r paths, fig.height=8, fig.width=12}
from_inquiring <- paths(dn, from = "Inquiring", traversal_time = 0.1)
from_inquiring
plot(from_inquiring)
```

From `Inquiring` every code is reached in one hop. `Arguing`, `Drafting`,
`Resourcing` and `Socialising` are reached at exactly 0.1 days, the traversal
cost with no waiting, and `Objecting` last, at 0.189 days. `Resourcing` is
reached by nine equally early routes.

To ask which codes could have fed into a code, leaving as late as possible and
still arriving by the end of the window, we set `direction` to `"backward"`
with `start` and `end` as the window. `arrival_time` then holds the latest
moment a sender could depart, and `latency` the deadline minus that moment.

```{r backward, fig.height=8, fig.width=12}
into_approving <- paths(dn, from = "Approving", direction = "backward",
                        start = 0, end = 4, traversal_time = 0.1)
into_approving
plot(into_approving)
```

`Arguing` and `Resourcing` can depart as late as day 3.9 and still reach
`Approving` by the deadline, while `Objecting` must depart by day 2.83, a
latency of 1.17 days, because the network thins towards the end of the window.

To rank whole routes by how often they are used across every source, we call
`pathways()` with `top` for the number of routes. The plot places a point per
code at the moment the route reaches it.

```{r pathways, fig.height=8, fig.width=12}
top_routes <- pathways(dn, top = 12)
top_routes
plot(top_routes)
```

## Timing

To compare the network across half-day bins, we call `similarity()` with
`step` and `window` of 0.5. The default coefficient is Jaccard, and the plot
is a heatmap of bins against bins.

```{r similarity, fig.height=7, fig.width=8}
bin_similarity <- similarity(dn, step = 0.5, window = 0.5)
plot(bin_similarity)
```

To classify how one turn follows another, we call `pshifts()`. It counts
Gibson's (2003) thirteen participation shifts, in which A addresses B and the
next turn is taken by B, by A again, or by a third party X.

```{r pshifts, fig.height=6, fig.width=10}
shifts <- pshifts(dn)
shifts
plot(shifts)
```

Of the 3,423 classified transitions, the largest family is turn usurping, in
which a third code takes the turn. The single most frequent shift is AB-XB,
642 cases: a new code addresses the code that was just addressed.

To measure whether each code's activity is clustered in time, we call
`burstiness()`. Burstiness compares the gaps between a code's replies with the
exponential reference: 1 is the bursty limit, 0 the Poisson reference and -1
regular activity (Goh and Barabási, 2008).

```{r burstiness, fig.height=6, fig.width=10}
bursts <- burstiness(dn)
summary(bursts)
plot(bursts)
```

Every code is bursty, from 0.43 for `Objecting` to 0.74 for `Resourcing`, and
memory is close to zero throughout: replies arrive in clusters whose lengths
do not predict one another.

## One course group

The `group` column travels with every spell, so a course group is a selection
over the network rather than a second construction. To select group `A_01`,
we call `induce_subgraph()` with `ties` set to a condition on the spell table.

```{r group}
one_group <- induce_subgraph(dn, ties = group == "A_01")
summary(one_group)
```

The group has 2,425 spells on 53 of the 81 ordered pairs, a mean snapshot
density of 0.559 and one session, and it is small enough for every reply to be
read. To draw each reply at the moment it fires, with codes on the vertical
axis and each link leaving in its source's colour and arriving in its
target's, we set `type` to `"events"`; `time = "clock"` places each reply at
its exact time and `blend = TRUE` blends the two colours along the link.

```{r group-events, fig.height=8, fig.width=13}
plot(one_group, type = "events")
```

```{r group-events-clock, fig.height=8, fig.width=13}
plot(one_group, type = "events", time = "clock", blend = TRUE)
```

To measure the group on the same hourly grid, we call `metrics()` on the
subnetwork and draw the result.

```{r group-metrics}
group_structure <- metrics(one_group,
                           measure = c("density", "edges", "reciprocity",
                                       "connectedness"),
                           step = 1 / 24, window = 1 / 24)
plot(group_structure, type = "ridge")
```

## Interpretation

The discussions return to `Resourcing` more than to any other kind of
contribution: it is incident to the most replies, it is the code most often
reached by several equally early routes, and it is one of the three codes that
relay most of the earliest routes between the others, with `Socialising` and
`Arguing`. `Objecting` is the opposite case, rarely addressed, slowest to
reach and never a relay. Because ties stay open for the life of a discussion,
every code can reach every other; what separates the codes is how quickly and
through whom, which is what the temporal measures report.

## Limitations

A tie that stays open until its discussion ends is a modelling choice that
makes the network dense and reachability near-total; a shorter tie life would
give sparser snapshots and longer latencies. The vertices are codes, not
students, so the network says nothing about who talked to whom. The four-day
window truncates the longest discussions. The table is the study's, trimmed
and anonymised as described in `?thought_chains`, and repeated rows for a
reply carrying two codes are counted as separate ties.

## References

Freeman, L. C. (1979). Centrality in social networks: conceptual
clarification. *Social Networks*, 1(3), 215–239.

Freeman, L. C., Borgatti, S. P., & White, D. R. (1991). Centrality in valued
graphs: a measure of betweenness based on network flow. *Social Networks*,
13(2), 141–154.

Gibson, D. R. (2003). Participation shifts: order and differentiation in group
conversation. *Social Forces*, 81(4), 1335–1380.

Goh, K.-I., & Barabási, A.-L. (2008). Burstiness and memory in complex systems.
*EPL (Europhysics Letters)*, 81(4), 48002.

Holland, P. W., & Leinhardt, S. (1976). Local structure in social networks.
*Sociological Methodology*, 7, 1–45.

Kempe, D., Kleinberg, J., & Kumar, A. (2002). Connectivity and inference
problems for temporal networks. *Journal of Computer and System Sciences*,
64(4), 820–842.

Krackhardt, D. (1994). Graph theoretical dimensions of informal organizations.
In K. M. Carley & M. J. Prietula (Eds.), *Computational organization theory*
(pp. 89–111). Lawrence Erlbaum.

Kundu, S., Murthy, C. A., & Pal, S. K. (2011). A new centrality measure for
influence maximization in social networks. In *Pattern Recognition and Machine
Intelligence* (Lecture Notes in Computer Science 6744, pp. 242–247). Springer.

Morris, M., Handcock, M. S., & Hunter, D. R. (2008). Specification of
exponential-family random graph models: terms and computational aspects.
*Journal of Statistical Software*, 24(4).

Pan, R. K., & Saramäki, J. (2011). Path lengths, correlations, and centrality
in temporal networks. *Physical Review E*, 84(1), 016105.

Saqr, M., & Nouri, J. (2020). High resolution temporal network analysis to
understand and improve collaborative learning. In *Proceedings of the Tenth
International Conference on Learning Analytics & Knowledge* (pp. 314–319). ACM.

Wasserman, S., & Faust, K. (1994). *Social network analysis: methods and
applications*. Cambridge University Press.
