---
title: "Trees of Thought: temporal-network reproduction with Dynet"
subtitle: "The study's networkDynamic / tsna analysis, re-run on the bundled `thought_chains` reply table, with a gallery of every temporal view"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup}
# Messages and warnings are left on: anything the package says while building
# or drawing belongs in the record, not suppressed out of it.
knitr::opts_chunk$set(
  message = TRUE, warning = TRUE, echo = TRUE,
  fig.width = 10, fig.height = 6, dpi = 120
)
input_file <- knitr::current_input(dir = TRUE)
repo_dir <- normalizePath(file.path(dirname(input_file), "../../.."),
                          mustWork = TRUE)
devtools::load_all(repo_dir)
# The multilayer views below need cograph's repaired plane spacing, layer and
# node labelling, and slice handling, which live in its working tree.
devtools::load_all(normalizePath(file.path(repo_dir, "../cograph"),
                                 mustWork = TRUE))
```

# What this document reproduces

The *Trees of Thought* study coded asynchronous discussion messages into ten
interaction codes and studied how one code follows another **over time**. Its
unit of analysis is not the student but the code: a tie runs from the code of a
message to the code of the message it replies to, so the network is a
code-to-code process network whose vertices are `Inquiring`, `Arguing`,
`Drafting`, `Coordinating` and the other five categories of the shipped table
(the study's ten codes, with *Evaluation* and *Acceptance* merged into
*Approving*).

The original analysis pipeline was a chain of packages. `Craete.Rmd` assembled
the spell table and called `networkDynamic()` to build the dynamic object;
`Analyse_trees.Rmd` and `Centralities_trees.Rmd` swept it with `tsna`
(`tSnaStats()`, `tErgmStats()`, `tiedDuration()`, `tReach()`, `tPath()`) and
`ndtv` (`proximity.timeline()`, `transmissionTimeline()`, `plotPaths()`);
`CraeteGROUP.Rmd` split it by course discussion group.

This document asks whether **one package** can carry that whole analysis. It
imports the saved network the study itself produced and re-runs each step with
an exported Dynet verb. The mapping is:

| Original call (script) | Dynet verb here |
|---|---|
| `networkDynamic()` built from the reply table with time rescaled to days since the discussion start (`Craete.Rmd`) | `dynet(thread = "discussion", thread_clock = "relative")` |
| `tSnaStats(gden / efficiency / connectedness)` | `metrics(measure = c("density", "efficiency", "connectedness"))` |
| `tSnaStats("mutuality")` and `tSnaStats("grecip", measure = "edgewise")` | the dyad census `mutual` count and `metrics(measure = "reciprocity")` |
| `tErgmStats("edges" / "meandeg" / "triangle" / "idegree1.5" / "odegree1.5")` | `metrics(measure = ...)` |
| `tSnaStats("dyad.census")`, `tSnaStats("triad.census")` | `metrics(measure = c("mutual", "asymmetric", "null"))`, `metrics(measure = "triads")` |
| `tErgmStats('nodemix("Name")')` | `mixing(attribute = "name")` |
| four `while` loops over `tSnaStats(snafun = degree / evcent / flowbet / betweenness / closeness / prestige)` | one `dyn_centrality(measure = ...)` call |
| `centiserve::diffusion.degree(lambda = 1)` on the aggregate | `dyn_centrality(measure = "diffusion")` |
| `tiedDuration(mode = , neighborhood = )` | `durations(unit = , mode = )` |
| `tReach(direction = "fwd", start = , end = )` | `dyn_reachability()`, `dyn_centrality(scope = "temporal")` |
| `proximity.timeline(time.increment = 0.1)` (`Visualize.Rmd`) | `plot(dn, type = "proximity")` |
| `tPath(type = "earliest.arrive", graph.step.time = 0.1)` + `transmissionTimeline()` / `plotPaths()` inside a `while` loop over the codes (`Visualize.Rmd`) | `paths(direction = "forward", traversal_time = 0.1)` + `plot_path_trajectories()` |
| `tPath(direction = "bkwd", type = "latest.depart", start = 0)` | `paths(direction = "backward", ...)` |
| `split()` on `course_group` and a rebuild per group (`CraeteGROUP.Rmd`) | `induce_subgraph(ties = ...)` |

Two differences from the original settings are deliberate and are flagged
where they occur. First, the study swept its statistics day by day
(`time.interval = 1`, `aggregate.dur = 1`); every sweep below uses
`step = 1 / 24, window = 1 / 24`, a ten-times finer grid over the same window, which
resolves the first day rather than collapsing it into one point. Second, the
trajectory trees in the path sections are a Dynet visual with no counterpart
in the original scripts, which drew `tPath()` results as hierarchical network
plots; that section says so where it happens.

## How to read this document

This is a reproduction, not a second implementation hidden in a notebook.
Every temporal-network number below is returned by an exported Dynet function.
The document imports the saved network, calls those functions, and prints or
plots the objects they return. It does not recompute a statistic by hand,
reshape one into a substitute for another, run package-parity tests, or alter
the supplied network. The final section audits those claims with code rather
than asserting them.

# The saved temporal network

## Import

The study's own saved network is not redistributable. This document runs on
`thought_chains`, the study's reply table shipped with the package with every
identity removed, the bottom fifth of authors trimmed, the two sparse weekdays
dropped, the calendar shifted by whole weeks, and *Evaluation* merged with
*Acceptance* into *Approving*. Every row is a real reply link with its real
weekday and time of day; nothing is resampled. See `?thought_chains`.

The study built its network from exactly this kind of table (`Craete.Rmd`):
a tie opens at the message's offset from the start of its discussion and stays
active until the discussion's last post. That is Dynet's threaded format with
each thread on its own clock, so the construction is one call. Self-links (a
code answering itself) are kept, as the study kept them, and the five courses
are picked up as sessions. Discussions run for four days: 99% of them are over
within 3.3 days and only three of 1,169 (40 links) drag on past the fourth,
so the study window is declared as four days with `observation_end = 4`,
which clips every measurement to it without touching the spells.

```{r import}
dn <- dynet(thought_chains, thread = "discussion", thread_clock = "relative",
            loops = TRUE, observation_end = 4)
dn
summary(dn)
# One course group, used below wherever every single reply has to be legible.
one_group <- induce_subgraph(dn, ties = group == "A_01")
```

The imported network has `r nrow(as.data.frame(dn, what = "nodes"))` codes and
`r nrow(as.data.frame(dn))` directed edge spells over
`r nrow(as.data.frame(dn, what = "network"))` distinct ordered pairs, of which
`r with(as.data.frame(dn), sum(from == to))` are self-loops.

## What a spell means here

This matters for everything that follows. In `Craete.Rmd` time is rescaled by
`3600 * 24`, so the clock is **days since the start of the discussion the
message belongs to**; `thread_clock = "relative"` is that clock. A tie opens
at the day offset of the message that created it and stays active until its
discussion ends. Ties therefore accumulate rather than flicker, most of them
open early, and the window is the declared four days.

```{r formation-dissolution-table}
tie_events <- events(dn, measure = c("formation", "dissolution"),
                     step = 0.5, window = 0.5)
head(tie_events, 8)
```

The first row of that table is the whole story of this network's shape:
the large majority of ties form at time 0. Every reachability and path result
later in the document follows from it.

# Seeing the temporal network

## Edge-spell timeline

One horizontal bar per spell, ordered by onset: the study's raw material
before any statistic is computed.

```{r edge-spell-timeline, fig.height=8}
plot(dn, type = "timeline", step = 1 / 24)
```

The bin is one hour (`step = 1 / 24` on a network measured in days).

## Formation, dissolution, and active ties

The same tie events as counts over time, with the active-tie stock beneath
them. This is the plotted form of the `events()` table above.

```{r edge-activity}
plot(dn, type = "activity")
```

## Whole-window network through cograph

The union of everything that was ever active, drawn by cograph with Dynet's
rendering defaults; `layout` is the only choice made here.

```{r whole-window-network, fig.height=8}
plot(dn, type = "network", layout = "oval")
```

## Duration-weighted collapsed network

The binary union above treats a tie that lived one hour like a tie that lived
five days. `collapse_network(weight = "union_duration")` weights each pair by
the total time it was active instead, which is the aggregate the original
analysis approximated by setting `edge.lwd` from `tiedDuration()`.

```{r collapsed-network, fig.height=8}
collapsed <- collapse_network(dn, weight = "union_duration")
plot(collapsed, layout = "oval")
```

## Temporal snapshots

Nine equally spaced cross-sections of the same network, the static-panel view
of the process.

```{r snapshots, fig.height=9, fig.width=10}
plot(dn, type = "snapshots", panels = 9)
```

## Proximity timeline with phase networks

`Visualize.Rmd` drew `proximity.timeline()` at `time.increment = 0.1` to show
codes drifting together and apart. Dynet's proximity view is the same idea:
each code is a line whose vertical position tracks its distance from the
others, sliced finely and annotated here with five phase networks.

```{r proximity-networks, fig.height=8, fig.width=12}
plot(dn, type = "proximity", phases = 5, slices = 80)
```

## Proximity trajectories alone

The same construction driven by betweenness instead of degree, without the
phase networks.

```{r proximity-lines, fig.height=7, fig.width=12}
plot(dn, type = "proximity", measure = "betweenness",
     networks = FALSE, slices = 80)
```

# Multilayer views of the sliced network

The original pipeline had no multilayer view: `ndtv` drew the process as a
proximity timeline and as animated snapshots, and the code-by-code structure
was read off aggregate matrices. Cutting the window into slices and treating
each slice as a layer of one multilayer network shows both at once — who was
talking to whom, and when.

Each slice carries the full code set, so a code keeps its identity across the
stack. The slicing, the supra-adjacency, the layer membership and the display
labels are assembled by the package.

## Node-link layers

Three slices of the four-day window, each a plane of the same nine codes.

```{r ml-layers, fig.height=9, fig.width=13}
plot(dn, type = "layers", step = 4 / 3, layout = "circle")
```

`omega` weights the identity arcs carrying a code from one slice to the next —
the interlayer coupling. Zero leaves the slices independent.

```{r ml-layers-uncoupled, fig.height=9, fig.width=13}
plot(dn, type = "layers", step = 4 / 3, omega = 0)
```

## Heatmap planes

The same slices as matrices rather than diagrams. Row and column names are
drawn once against the front plane, so a cell reads as a code pair rather than
an anonymous square.

```{r ml-heatmap, fig.height=9, fig.width=13}
plot(dn, type = "heatmap", step = 4 / 3)
```

Thresholded, so only substantial transitions carry ink:

```{r ml-heatmap-threshold, fig.height=9, fig.width=13}
plot(dn, type = "heatmap", step = 4 / 3, threshold = 50)
```

At the hourly step used for the heatmaps below, the four days split into 96
slices. That is the right resolution for a grid or a line, and too many planes
to read as a stack — shown here so the limit is visible rather than asserted.

```{r ml-heatmap-fine, fig.height=12, fig.width=13}
plot(dn, type = "heatmap", step = 1 / 24, show_node_labels = FALSE)
```

## Projected stack

`cograph::plot_temporal()`'s oblique projection, with one colour per code held
across every plane. Empty windows are drawn rather than skipped, so a gap in
the process is visible as a gap.

```{r ml-stack, fig.height=10, fig.width=13}
plot(dn, type = "stack", step = 4 / 3)
```

Cumulative, so each plane carries everything that has happened up to it:

```{r ml-stack-cumulative, fig.height=10, fig.width=13}
plot(dn, type = "stack", step = 4 / 3, cumulative = TRUE)
```

# Graph-level trajectories

## Main structural descriptives

The original scripts collected these one call at a time and `cbind()`-ed the
results into a wide matrix. `metrics()` takes the whole set in one call and
returns one tidy row per time point and measure.

```{r graph-main}
graph_main <- metrics(
  dn,
  measure = c(
    "density", "efficiency", "connectedness", "reciprocity",
    "edges", "mean_degree"
  ),
  step = 1 / 24, window = 1 / 24
)
graph_main
plot(graph_main, type = "ridge")
```

`reciprocity` here is the edgewise reciprocity — the share of arcs that are
reciprocated — which is what `tSnaStats("grecip", measure = "edgewise")`
returned in `Centralities_trees.Rmd`. The study also reported
`tSnaStats("mutuality")`, the raw count of mutual dyads; that count is the
`mutual` column of the dyad census below.

## Dyad census

```{r dyad-census}
dyads <- metrics(
  dn, measure = c("mutual", "asymmetric", "null"),
  step = 1 / 24, window = 1 / 24
)
dyads
plot(dyads)
```

## Triad census

`Centralities_trees.Rmd` swept a triad census with
`tSnaStats("triad.census")`. Dynet returns the same sixteen MAN triad types,
one row per type per time point, drawn here as a heatmap.

```{r triad-census, fig.height=8}
triads <- metrics(dn, measure = "triads", step = 1 / 24, window = 1 / 24)
triads
plot(triads, type = "heatmap")
```

## Lightweight ERGM-style descriptives

The counterparts of the study's `tErgmStats()` terms — `idegree1.5`,
`odegree1.5`, `triangle` — plus the two-star and two-path counts that describe
how the local structure builds up.

```{r ergm-descriptives}
ergm_descriptives <- metrics(
  dn,
  measure = c(
    "indegree_1_5", "outdegree_1_5", "triangles",
    "in_2stars", "out_2stars", "two_paths"
  ),
  step = 1 / 24, window = 1 / 24
)
ergm_descriptives
plot(ergm_descriptives, type = "ridge")
```

# Node-level trajectories

## Centrality

The original centrality section was four `while` loops that called
`tSnaStats()` once per measure, transposed each result, and stitched the
pieces back together with `cbind()` and `rbind()` into
`Centralities_Combined_rounded.xlsx`. Here the same five measures — degree,
closeness, betweenness, eigenvector, flow betweenness — plus diffusion degree
(the study computed that one separately, with `centiserve`, and only on the
aggregate) come from a single call that already returns them tidily.

```{r centrality-heatmap, fig.height=11}
centrality <- dyn_centrality(
  dn,
  measure = c(
    "degree", "closeness", "betweenness", "eigenvector",
    "flow_betweenness", "diffusion"
  ),
  step = 1 / 24, window = 1 / 24
)
centrality
plot(centrality, type = "heatmap")
```

```{r centrality-trajectories, fig.height=10}
plot(centrality, type = "ridge", top = 10)
```

## In-degree and out-degree

The study ran separate loops with `cmode = "indegree"` and
`cmode = "outdegree"`; in Dynet the direction is the `mode` argument.

```{r directed-degree, fig.height=7}
directed_degree <- dyn_centrality(
  dn, measure = "degree", mode = "in",
  step = 1 / 24, window = 1 / 24
)
out_degree <- dyn_centrality(
  dn, measure = "degree", mode = "out",
  step = 1 / 24, window = 1 / 24
)
plot(directed_degree, type = "heatmap")
plot(out_degree, type = "heatmap")
```

## Prestige

The four prestige variants the study looped over — `indegree`, `eigenvector`,
`domain` and `domain.proximity` — are the four values of the `prestige`
argument.

```{r prestige, fig.height=7}
prestige_indegree <- dyn_centrality(
  dn, measure = "prestige", prestige = "indegree",
  step = 1 / 24, window = 1 / 24
)
prestige_eigenvector <- dyn_centrality(
  dn, measure = "prestige", prestige = "eigenvector",
  step = 1 / 24, window = 1 / 24
)
prestige_domain <- dyn_centrality(
  dn, measure = "prestige", prestige = "domain",
  step = 1 / 24, window = 1 / 24
)
prestige_proximity <- dyn_centrality(
  dn, measure = "prestige", prestige = "domain.proximity",
  step = 1 / 24, window = 1 / 24
)
plot(prestige_indegree, type = "heatmap")
plot(prestige_eigenvector, type = "heatmap")
plot(prestige_domain, type = "heatmap")
plot(prestige_proximity, type = "heatmap")
```

# Code-to-code mixing

`tErgmStats('nodemix("Name")')` produced one column per ordered pair of codes,
which the original scripts then had to un-widen and re-aggregate by hand.
`mixing()` returns the same 9 x 9 flows already long, one row per time point
and per ordered pair.

```{r mixing-heatmap, fig.height=14}
mixing_flows <- mixing(dn, attribute = "name", step = 1 / 24, window = 1 / 24)
mixing_flows
plot(mixing_flows, type = "heatmap")
```

The original scripts then pulled out the mixing columns that involved
regulation (`select(contains("reg"))`, which caught both `Group_regulation`
and `T.Regulation`, here `Coordinating` and `Tutoring`). The highlight below
narrows that to the coordinating flows, which keeps the lines distinguishable. The result carries `from_group`
and `to_group` columns, so the selection is a `subset()` of the returned table
handed straight back to the plot method; no mixing statistic is rebuilt
here.

```{r mixing-group-regulation, fig.height=7}
group_regulation_flows <- with(
  subset(as.data.frame(mixing_flows),
         from_group == "Coordinating" | to_group == "Coordinating"),
  unique(measure)
)
plot(mixing_flows, highlight = group_regulation_flows)
```

# Tie and vertex duration

## Incident tie counts and durations by node

`tiedDuration()` was called six times in `Visualize.Rmd` — count and duration,
crossed with the `in`, `out` and `combined` neighbourhoods — and the six
vectors were `cbind()`-ed into a table. `durations(unit = "node_ties")` gives
the same quantities per node, with `mode` selecting the neighbourhood and
`measure` selecting event count, total duration, or the union of active time
(which, unlike the total, does not double-count overlapping spells).

```{r node-tie-duration, fig.height=10}
node_ties_all <- durations(
  dn, unit = "node_ties", mode = "all",
  measure = c("events", "total", "union")
)
node_ties_in <- durations(
  dn, unit = "node_ties", mode = "in",
  measure = c("events", "total", "union")
)
node_ties_out <- durations(
  dn, unit = "node_ties", mode = "out",
  measure = c("events", "total", "union")
)
plot(node_ties_all)
plot(node_ties_in)
plot(node_ties_out)
```

## Edge-pair duration

The same accounting one level down, per ordered pair of codes, with the mean
spell length added.

```{r pair-duration, fig.height=10}
pair_duration <- durations(
  dn, unit = "pair",
  measure = c("events", "total", "union", "mean")
)
pair_duration
plot(pair_duration)
```

## Vertex activity duration

How long each code was itself active, as opposed to how long its ties were.

```{r vertex-duration, fig.height=8}
vertex_duration <- durations(dn, unit = "vertex_activity")
vertex_duration
plot(vertex_duration)
```

# Temporal reachability and centrality

`tReach()` was used to count, for each code, how many other codes it could
reach along time-respecting paths within a window. `dyn_reachability()`
answers the same question in both directions at once, and
`dyn_centrality(scope = "temporal")` returns the path-based measures computed
on the time-respecting paths themselves rather than on a snapshot.

```{r reachability, fig.height=8}
reachability <- dyn_reachability(
  dn, direction = "both", measure = c("reach", "reach_count")
)
reachability
plot(reachability)
```

```{r temporal-centrality, fig.height=9}
temporal_centrality <- dyn_centrality(
  dn,
  measure = c("closeness", "betweenness", "reach", "reach_count"),
  scope = "temporal"
)
temporal_centrality
plot(temporal_centrality)
```

# Forward temporal paths from every code

`Visualize.Rmd` looped over the codes with
`tPath(direction = "fwd", type = "earliest.arrive", graph.step.time = 0.1)`
and drew three pictures per code. `paths()` answers the same query — Dynet
receives that traversal cost as `traversal_time = 0.1`, one tenth of a day per
hop — and returns one tidy object per source.

Before the atlas, one source in full, to show what the objects contain.

```{r forward-path-example}
inquiring_paths <- paths(
  dn, from = "Inquiring", direction = "forward", traversal_time = 0.1
)
inquiring_paths
path_trajectories(inquiring_paths)
```

The printed table gives, per destination, the earliest attainable arrival
time, the latency from the source, the number of hops on the optimal route,
and how many distinct optimal routes achieve it. `path_trajectories()` is the
tidy tree behind the picture: one row per route prefix, so a code reached
under two different temporal histories is two rows and the branches never
cross misleadingly.

The atlas below repeats that for every code. Each panel is a left-to-right
trajectory tree: branch width and node size show how many optimal routes use a
branch, node fill shows the same count, and every node prints its vertex name
and value beneath its circle, so nothing is carried by colour alone. The table
above each tree is the path result the tree is drawn from.

```{r forward-path-atlas, results='asis', fig.height=8, fig.width=12}
# A loop, not an apply: each iteration emits a section heading, a table and a
# plot into the asis stream in order, which is a sequence of side effects
# rather than a value to collect.
code_names <- with(as.data.frame(dn, what = "nodes"), name)
for (source_node in code_names) {
  cat("\n\n## ", source_node, "\n\n", sep = "")
  forward_paths <- paths(
    dn, from = source_node, direction = "forward", traversal_time = 0.1
  )
  print(knitr::kable(
    as.data.frame(forward_paths), digits = 3,
    caption = paste("Earliest-arrival paths from", source_node)
  ))
  cat("\n\n")
  print(plot_path_trajectories(
    forward_paths, measure = "frequency", orientation = "horizontal"
  ))
  cat("\n\n")
}
```

Read the tables together and the picture is consistent: because almost every
tie opens at time 0, every code reaches every other code, and the `n_hops`
and `latency` columns show it doing so in one or two hops within a fraction
of a day. The trees are wide and shallow. That is a property of how the study built its spells —
a tie stays open until its discussion ends — not an artefact of the traversal
setting.

# Backward temporal paths to every code

The complementary query in `Visualize.Rmd` was
`tPath(direction = "bkwd", type = "latest.depart", start = 0)`, again at
`graph.step.time = 0.1`: which codes could have fed into this one, leaving as
late as possible. Dynet expresses it as `direction = "backward"` with an
explicit window. The deadline here is `end = 4`, four days into the roughly
eight-day span; a later deadline admits the same senders but with longer
latencies, so the earlier deadline keeps the trees interpretable.

```{r backward-path-atlas, results='asis', fig.height=8, fig.width=12}
# Same reason for the loop as above: ordered side effects per code.
for (target_node in code_names) {
  cat("\n\n## ", target_node, "\n\n", sep = "")
  backward_paths <- paths(
    dn, from = target_node, direction = "backward", start = 0, end = 4,
    traversal_time = 0.1
  )
  print(knitr::kable(
    as.data.frame(backward_paths), digits = 3,
    caption = paste("Latest-departure paths into", target_node)
  ))
  cat("\n\n")
  print(plot_path_trajectories(
    backward_paths, measure = "frequency", orientation = "horizontal"
  ))
  cat("\n\n")
}
```

In these trees the queried code is the root and the possible senders branch
away from it, so a branch is read right to left in time: `arrival_time` is the
latest moment a sender could have departed and still reach the root by the
deadline.

# Every temporal view, on this network

The study drew its process with `ndtv`: a proximity timeline and animated
snapshots. Dynet has nine views of a temporal network and a plot for every
result class. They are all shown here on the same object, so the reader can
pick the one that answers the question at hand. Node-link rendering is
cograph's throughout.

## Contacts as curved links on the time axis

`type = "events"` draws every contact at the moment it fires, actors on the
vertical axis, each link leaving its source in the source's colour and arriving
in the target's. On the whole network only the busiest pairs are legible, so
the first picture keeps the thirty busiest and the rest use one course group,
`A_01`, where every reply can be seen.

```{r events-top, fig.height=8, fig.width=13}
plot(dn, type = "events", top = 30)
```

```{r events-group, fig.height=8, fig.width=13}
plot(one_group, type = "events")
```

The link glyph is a choice. `"hook"` above is the default; the other four are
arcs, chevrons, waves and brackets, each with the same colour run.

```{r events-arc, fig.height=8, fig.width=13}
plot(one_group, type = "events", link = "arc")
```

```{r events-chevron, fig.height=8, fig.width=13}
plot(one_group, type = "events", link = "chevron")
```

```{r events-wave, fig.height=8, fig.width=13}
plot(one_group, type = "events", link = "wave")
```

```{r events-bracket, fig.height=8, fig.width=13}
plot(one_group, type = "events", link = "bracket")
```

`curvature` sets how far a link bows (`0` is straight), and `time = "clock"`
places each contact at its exact time rather than in its bin, with `blend`
mixing the two endpoint colours along the link (both with the default hook).

```{r events-straight, fig.height=8, fig.width=13}
plot(one_group, type = "events", curvature = 0)
```

```{r events-clock, fig.height=8, fig.width=13}
plot(one_group, type = "events", time = "clock", blend = TRUE)
```

## The other eight views

`"activity"`: ties forming and dissolving over time.

```{r view-activity, fig.height=6, fig.width=12}
plot(dn, type = "activity")
```

`"network"` at one bin, against the whole-window picture drawn earlier:

```{r view-network-at, fig.height=8, fig.width=9}
plot(dn, type = "network", at = 1)
```

`"snapshots"`, `"layers"`, `"heatmap"`, `"stack"` and `"proximity"` are drawn
in the sections above with the study's settings; `"timeline"` opens the
document. Two more proximity readings close the set: the lines alone, driven
by betweenness, and two codes highlighted against the rest.

```{r view-proximity-highlight, fig.height=7, fig.width=12}
plot(dn, type = "proximity", networks = FALSE,
     highlight = c("Arguing", "Approving"))
```

## Plots of every result class

Routes on a real time axis: `pathways()` ranks whole time-respecting routes by
how often they are used; the plot places a point per vertex at the moment the
route reaches it, so a horizontal gap is waiting time.

```{r result-pathways, fig.height=8, fig.width=12}
top_routes <- pathways(dn, top = 12)
plot(top_routes)
```

Similarity between every pair of half-day bins, as a heatmap:

```{r result-similarity, fig.height=7, fig.width=8}
bin_similarity <- similarity(dn, step = 0.5, window = 0.5)
plot(bin_similarity)
```

Gibson's participation shifts, the thirteen ways one turn follows another:

```{r result-pshifts, fig.height=6, fig.width=10}
shifts <- pshifts(dn)
plot(shifts)
```

Burstiness and memory of each code's activity:

```{r result-burstiness, fig.height=6, fig.width=10}
bursts <- burstiness(dn)
plot(bursts)
```

Tie durations by pair, and reachability, both already tabulated above:

```{r result-durations, fig.height=8, fig.width=10}
pair_totals <- durations(dn, unit = "pair", measure = "total")
plot(pair_totals)
```

```{r result-reachability, fig.height=6, fig.width=10}
reach <- dyn_reachability(dn)
plot(reach)
```

Snapshots as a result object rather than a view:

```{r result-snapshots, fig.height=8, fig.width=12}
daily <- snapshots(dn, step = 1, window = 1)
plot(daily)
```

The three trajectory-tree colourings, horizontal and vertical, for the
forward paths from *Inquiring*:

```{r result-trajectories, fig.height=8, fig.width=12}
inquiry <- path_trajectories(paths(dn, from = "Inquiring"))
plot_path_trajectories(inquiry, measure = "frequency")
plot_path_trajectories(inquiry, measure = "time", orientation = "vertical")
plot_path_trajectories(inquiry, measure = "predictability")
```

The path network of those same paths, as a cograph node-link drawing:

```{r result-path-network, fig.height=8}
from_inquiring <- paths(dn, from = "Inquiring")
inquiring_network <- path_network(from_inquiring)
plot(inquiring_network, layout = "oval")
```

# Course-group temporal subnetwork

`CraeteGROUP.Rmd` split the interaction table by `course_group` and rebuilt a
separate `networkDynamic` object for each group. Dynet keeps the tie
attributes from the import, so a group is a selection over the existing
network rather than a second construction: `induce_subgraph()` accepts a mask
over the spell table, which is exactly the legacy filter expressed as an
argument.

`thought_chains` keeps the course group as `group`, so the study's split is
one condition on the spell table: no mask has to be built first.

```{r group-subnetwork-import}
summary(one_group)
```

The subnetwork is a selection over the existing object, not a second
construction, and it is small enough to read code by code.

```{r group-subnetwork-splot, fig.height=8}
plot(one_group, type = "network", layout = "oval")
```

```{r group-subnetwork-timeline, fig.height=8}
plot(one_group, type = "timeline")
```

```{r group-subnetwork-snapshots, fig.height=9}
plot(one_group, type = "snapshots", panels = 9)
```

```{r group-subnetwork-metrics}
group_structure <- metrics(one_group,
                           measure = c("density", "edges", "reciprocity", "connectedness"),
                           step = 1 / 24, window = 1 / 24)
plot(group_structure, type = "ridge")
```

# Reproduction audit

The claims made at the top are checked here rather than asserted.

**Every verb used came from Dynet's public interface.** If any of these were
internal, or had been renamed, the check below would say so.

```{r audit-exports}
verbs_used <- c(
  "as_dynet", "collapse_network", "events", "metrics", "mixing",
  "durations", "dyn_centrality", "dyn_reachability", "paths",
  "path_trajectories", "plot_path_trajectories", "induce_subgraph",
  "pathways", "similarity", "pshifts", "burstiness", "snapshots",
  "projection", "collapse_network", "path_network"
)
data.frame(
  verb = verbs_used,
  exported = verbs_used %in% getNamespaceExports("Dynet")
)
```

**The analysed network is the saved network, unaltered.** Re-importing the
file after every statistic above has been computed must give an object
identical to the one they were computed on.

```{r audit-unchanged}
identical(dn, dynet(thought_chains, thread = "discussion",
                    thread_clock = "relative", loops = TRUE))
```

**The self-loops were kept, not quietly dropped.** The study built the network
with `loops = TRUE`. The count of loop spells in the imported network and the
self-pairs carried through to the pair-duration table must agree.

```{r audit-loops}
with(as.data.frame(dn), sum(from == to))
subset(as.data.frame(pair_duration), from == to & measure == "events")
```

What this document does **not** establish: it is not a numerical parity test
against `tsna`. The sweep grid differs from the original by design
(hourly against the study's daily interval), a few measures are the
nearest documented counterpart rather than the identical estimator, and no
output here was compared value-by-value with a `tsna` run. The claim is that each analytical
step of the study is available as one Dynet verb on the study's own data, and
that the results are internally consistent — not that the two implementations
agree to floating-point tolerance.

# Session information

```{r session-info}
sessionInfo()
```
